---
title: "Data visualization"
author: "María Bueno Alvez & Max J Karlsson"
date: "01/10/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

#Placeholder

#Setup

##Load libraries

```{r setup, include=FALSE}

library(tidyverse)
library(pbapply)
library(readxl)

library(tidygraph)
library(ggraph)
library(ggrepel)
library(ggthemes)
library(ggrastr)
library(ggforce)
library(ggupset)
library(plotly)
library(patchwork)

library(uwot)

source("scripts/functions_utility.R")

select <- dplyr::select


```

## File structure
```{r}

dataset_metadata <- read_csv("run_settings/20210928_settings.csv")

data_paths <- read_csv("run_settings/20211012 all_datasets.csv")

file_structure <-
  create_folder_structure(dataset_id = dataset_metadata$dataset_id,
                          dataset_run_id = dataset_metadata$version,
                          main_folder = "results/Clustering_results",
                          folders = c("data" = "data", 
                                      "svg" = "svg",
                                      "UMAP" = "UMAP",
                                      "treemap" = "svg/treemap",
                                      "heatmap" = "svg/heatmap",
                                      "enrichment" = "enrichment",
                                      "clustering" = "clustering"))
```


## Load clustering data

```{r}

clustering_data <- 
  file_structure %>% 
  map(function(x) x$clustering) %>% 
  map(function(x) {
    file_ <- paste(x, "final_consensus.tsv", sep = "/")
    if(file.exists(file_)) {
      read_tsv(file_)
    } else {
      NULL
    }
  }) %>% 
  bind_rows(.id = "dataset_id") %>% 
  mutate(cluster = as.character(cluster)) 


```

## Load enrichment data

```{r}

enrichment_data <- 
  file_structure %>% 
  map(function(x) x$enrichment) %>% 
  map(function(x) {
    file_ <- paste(x, "enrichment_results.tsv", sep = "/")
    if(file.exists(file_)) {
      read_tsv(file_)
    } else {
      NULL
    }
  }) %>% 
  bind_rows(.id = "dataset_id") %>% 
  mutate(cluster = as.character(cluster)) 

```

## Load annotation

```{r}

cluster_annotation <-
  # list(tissue = read_csv("data/annotation/Tissue 20211013.csv"),
  #      blood = read_csv("data/annotation/Blood 20211013.csv"),
  #      celline = read_csv("data/annotation/Celline 20211013.csv"),
  #      brain = read_csv("data/annotation/Brain 20211013.csv")) %>% 
  list.files("data/annotation", full.names = T) %>% 
  {.[grepl("20211018", .)]} %>% 
  set_names(., gsub(".*\\d |\\.xlsx$", "", .)) %>% 
  map(. %>% read_excel()) %>% 
  bind_rows(.id = "dataset_id") %>% 
  rename(dataset_id = 1,
         cluster = 2,
         n_genes = 3,
         ann_specificity = 5, 
         ann_function = 6,
         reliability = 7,
         external_comment = 8,
         internal_comment = 9) %>% 
  mutate(cluster = as.character(cluster),
         full_annotation = paste(ann_specificity, 
                                 ann_function, 
                                 sep = " - "))

```

## Load UMAP data 

```{r}
umap_data <- 
  file_structure %>% 
  map(function(x) x$UMAP) %>% 
  map(function(x) {
    file_ <- paste(x, "UMAP.tsv", sep = "/")
    if(file.exists(file_)) {
      read_tsv(file_)
    } else {
      NULL
    }
  }) %>% 
  bind_rows(.id = "dataset_id") 

umap_poly_data <- 
  file_structure %>% 
  map(function(x) x$UMAP) %>% 
  map(function(x) {
    file_ <- paste(x, "UMAP_polygons.tsv", sep = "/")
    if(file.exists(file_)) {
      read_tsv(file_)
    } else {
      NULL
    }
  }) %>% 
  bind_rows(.id = "dataset_id") %>% 
  mutate(cluster = as.character(cluster)) 

```

## Load raw data

```{r}

raw_data <-
  readRDS("data/processed/combined_HPA_expression_data.RDS")

```

## Load other metadata


### Colors

```{r}

dataset_palette <- 
  c("tissue" = "#0083C4",  
    "brain" = "#FFDD00", 
    "blood" = "#CF161A", 
    "singlecell" = "#6AA692", 
    "celline" = "#97CF16")

dataset_palette2 <- 
  c("Tissue" = "#0083C4",  
    "Brain" = "#FFDD00", 
    "Blood" = "#CF161A", 
    "Single cell" = "#6AA692", 
    "Cell line" = "#97CF16")

all_HPA_colors <- 
  read_tsv("data/meta/HPA21_colors.txt", 
           col_names = c("organ_name", "tissue_name", "color")) 


organ_palette <- 
  all_HPA_colors %>% 
  select(organ_name, color) %>% 
  distinct() %>% 
  deframe()

tissue_palette <- 
  all_HPA_colors %>% 
  filter(organ_name != "Brain") %>% 
  gather(tissue_type, tissue, 1, 2) %>% 
  select(tissue, color) %>% 
  distinct() %>% 
  bind_rows(mutate(.,
                   tissue = str_to_sentence(tissue))) %>% 
  bind_rows(cluster_annotation %>% 
              select(ann_specificity) %>% 
              distinct() %>% 
              filter(grepl("^Low.*specificity$", ann_specificity)) %>% 
              rename(tissue = 1) %>% 
              mutate(color = "darkgray")) %>% 
  bind_rows(all_HPA_colors %>% 
              filter(organ_name == "Brain") %>% 
              select(tissue = organ_name, color) %>% 
              distinct())

# tissue_palette %>% group_by(tissue) %>% 
#   count %>% 
#   filter(n > 1)
#   
# all_HPA_colors %>% 
#   filter(tissue_name == "anterior cingulate gyrus, pregenual-dorsal") 
#   pull(color) %>% 
#   prismatic::color()
# 
# read_tsv("data/meta/HPA21_colors.txt", 
#            col_names = c("organ_name", "tissue_name", "color")) %>% 
#   filter(tissue_name == "amygdala") %>% 
#   pull(color) %>% 
#   
# 
# all_HPA_colors %>% 
#   group_by(tissue) %>% 
#   count %>% 
#   filter(n > 1) %>% 
#   View

cluster_color_pal <-
  tableau_color_pal(palette = "Classic Cyclic",
                    type = 'regular')(13) %>%
  colorRampPalette()


```


# Data overview

## Raw data

```{r}
raw_data %>% 
  map(. %>% 
        select(-1) %>% 
        ncol) %>% 
  unlist() %>% 
  enframe()
```

## Clustering

```{r}



```

## Annotation

```{r}

cluster_annotation %>% 
  select(ann_specificity1 = ann_specificity) %>% 
  distinct() %>% 
  expand_grid(ann_specificity2 = .$ann_specificity1) %>% 
  mutate(sim = 1 - stringdist::stringsim(ann_specificity1, ann_specificity2)) %>% 
  filter(ann_specificity1 < ann_specificity2) %>% 
  filter(sim < 0.3)%>% 
  arrange(sim)

cluster_annotation %>% 
  select(ann_specificity1 = ann_specificity) %>% 
  distinct() %>% 
  expand_grid(ann_specificity2 = .$ann_specificity1) %>% 
  mutate(sim = 1 - stringdist::stringsim(ann_specificity1, ann_specificity2)) %>% 
  filter(sim < 0.3) %>% 
  filter(ann_specificity1 != ann_specificity2) %>% 
  spread(ann_specificity2, sim, fill = 1) %>% 
  column_to_rownames("ann_specificity1") %>% 
  pheatmap::pheatmap()

cluster_annotation %>% 
  select(ann_function1 = ann_function ) %>% 
  distinct() %>% 
  expand_grid(ann_function2 = .$ann_function1) %>% 
  mutate(sim = 1 - stringdist::stringsim(ann_function1, ann_function2)) %>% 
  filter(ann_function1 < ann_function2) %>% 
  filter(sim < 0.3) %>% 
  arrange(sim)

cluster_annotation %>% 
  filter(grepl("&", full_annotation)) 

cluster_annotation %>% 
  filter(grepl(" and ", full_annotation))

cluster_annotation %>% 
  filter(grepl("-", ann_specificity)) %>% 
  select(ann_specificity) %>% 
  distinct() 

cluster_annotation %>% 
  filter(grepl("-", ann_function)) %>% 
  select(ann_function) %>% 
  distinct() 



##################################

annotation_status <-
  cluster_annotation %>% 
  select(dataset_id, cluster, n_genes, ann_specificity, ann_function) %>% 
  mutate(function_status = case_when(ann_function %in% c("Unknown function",
                                                         "Mixed function") ~
                                       ann_function,
                                     T ~ "Annotated function"),
         specificity_status = case_when(grepl("^Low.*specificity$", ann_specificity) ~ 
                                          "Low specificity",
                                        T ~ "Annotated specificity"))

plot_data <- annotation_status

p1 <- 
  plot_data %>% 
  group_by(dataset_id, function_status) %>% 
  count() %>% 
  ggplot(aes(n, dataset_id, fill = function_status, label = n)) +
  geom_col() + 
  geom_text(position = position_stack(vjust = 0.5),
            color = "white") +
  geom_text(data = . %>% 
              group_by(dataset_id) %>% 
              summarise(n = sum(n)),
            inherit.aes = F,
            aes(n, dataset_id, label = n),
            hjust = 0) +
  theme_void() +
  theme(axis.text.y = element_text(hjust = 1), 
        legend.position = "top",
        legend.direction = "vertical") +
  scale_x_continuous(expand = expansion(c(0, 0.2))) +
  scale_fill_manual(values = c("Annotated function" = "#647D96",
                               "Mixed function" = "gray",
                               "Unknown function" = "darkgray"),
                    name = "") +
  ggtitle("Function annotation")

p2 <- 
  plot_data %>% 
  group_by(dataset_id, specificity_status) %>% 
  count() %>% 
  ggplot(aes(n, dataset_id, fill = specificity_status, label = n)) +
  geom_col() + 
  geom_text(position = position_stack(vjust = 0.5),
            color = "white") +
  geom_text(data = . %>% 
              group_by(dataset_id) %>% 
              summarise(n = sum(n)),
            inherit.aes = F,
            aes(n, dataset_id, label = n),
            hjust = 0) +
  theme_void() +
  theme(axis.text.y = element_text(hjust = 1), 
        legend.position = "top", 
        legend.direction = "vertical") +
  scale_x_continuous(expand = expansion(c(0, 0.2))) +
  scale_fill_manual(values = c("Annotated specificity" = "#647D96",
                               "Low specificity" = "darkgray"),
                    name = "") +
  ggtitle("Specificity annotation")

p2 | p1

ggsave(savepath("Annotation status.pdf"), 
       width = 7, height = 4)
ggsave(savepath("Annotation status.png"), 
       width = 7, height = 4)

#####

p1 <- 
  plot_data %>% 
  group_by(dataset_id, function_status) %>% 
  summarise(n = sum(n_genes)) %>% 
  ggplot(aes(n, dataset_id, fill = function_status, label = n)) +
  geom_col() + 
  geom_text(position = position_stack(vjust = 0.5),
            color = "white",
            angle = -90) +
  geom_text(data = . %>% 
              group_by(dataset_id) %>% 
              summarise(n = sum(n)),
            inherit.aes = F,
            aes(n, dataset_id, label = n),
            hjust = 0) +
  theme_void() +
  theme(axis.text.y = element_text(hjust = 1), 
        legend.position = "top",
        legend.direction = "vertical") +
  scale_x_continuous(expand = expansion(c(0, 0.2))) +
  scale_fill_manual(values = c("Annotated function" = "#647D96",
                               "Mixed function" = "gray",
                               "Unknown function" = "darkgray"),
                    name = "") +
  ggtitle("Function annotation")

p2 <- 
  plot_data %>% 
  group_by(dataset_id, specificity_status) %>% 
  summarise(n = sum(n_genes)) %>% 
  ggplot(aes(n, dataset_id, fill = specificity_status, label = n)) +
  geom_col() + 
  geom_text(position = position_stack(vjust = 0.5),
            color = "white", 
            angle = -90) +
  geom_text(data = . %>% 
              group_by(dataset_id) %>% 
              summarise(n = sum(n)),
            inherit.aes = F,
            aes(n, dataset_id, label = n),
            hjust = 0) +
  theme_void() +
  theme(axis.text.y = element_text(hjust = 1), 
        legend.position = "top", 
        legend.direction = "vertical") +
  scale_x_continuous(expand = expansion(c(0, 0.2))) +
  scale_fill_manual(values = c("Annotated specificity" = "#647D96",
                               "Low specificity" = "darkgray"),
                    name = "") +
  ggtitle("Specificity annotation")

p2 | p1

ggsave(savepath("Annotation status genewise.pdf"), 
       width = 7, height = 4)
ggsave(savepath("Annotation status genewise.png"), 
       width = 7, height = 4)



plot_data %>% 
  group_by(function_status) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(frac = n / sum(n))

plot_data %>% 
  group_by(specificity_status) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(frac = n / sum(n))

plot_data %>% 
  group_by(function_status) %>% 
  summarise(n = sum(n_genes)) %>% 
  ungroup() %>% 
  mutate(frac = n / sum(n))

plot_data %>% 
  group_by(dataset_id, specificity_status) %>% 
  summarise(n = sum(n_genes)) %>% 
  group_by(dataset_id) %>% 
  # ungroup() %>% 
  mutate(frac = n / sum(n)) %>% 
  filter(specificity_status == "Annotated specificity")


#########################


clustering_data %>% 
  left_join(annotation_status) %>% 
  group_by(gene, specificity_status) %>% 
  summarise(datasets = list(sort(dataset_id))) %>% 
  
  # mutate(datasets)
  group_by(specificity_status, datasets) %>% 
  count() %>% 
  group_by(datasets) %>% 
  mutate(tot_n = sum(n)) %>% 
  ungroup() %>% 
  arrange(tot_n) %>% 
  filter(specificity_status == "Low specificity") %>% 
  group_by_all() %>% 
  mutate(n_dat = length(datasets[[1]])) %>% 
  ungroup() %>% 
  ggplot(aes(datasets, n, fill = specificity_status)) +
  geom_col() +
  scale_x_upset(n_intersections = 30) 

clustering_data %>% 
  left_join(annotation_status) %>% 
  select(dataset_id, gene, cluster, function_status, specificity_status) %>% 
  group_by(gene, specificity_status) %>% 
  count() %>% 
  group_by(specificity_status, n) %>% 
  count() %>%
  filter(specificity_status == "Low specificity") %>% 
  ggplot(aes(n, nn, fill = n, label = nn)) +
  geom_col() +
  geom_text(position = position_stack(), 
            vjust = 0) +
  scale_fill_gradient(low = "gray80", high = "gray20") +
  labs(x = "Number of datasets gene is in low specificity clustering") +
  theme_bw()

plot_data

```

## Enrichment data

```{r}

enrichment_data %>% 
  filter(test_type == "reactome") %>% 
  group_by(dataset_id, cluster) %>% 
  count %>% 
  arrange(-n)

plot_data <- 
  enrichment_data %>% 
  select(dataset_id, test_type, cluster) %>% 
  distinct() %>% 
  group_by(dataset_id, test_type) %>% 
  count() %>% 
  filter(!(grepl("GO", test_type) & grepl("original", test_type))) %>% 
  left_join(clustering_data %>% 
              group_by(dataset_id) %>% 
              summarise(n_cluster = n_distinct(cluster))) %>% 
  mutate(frac = n / n_cluster)


plot_data %>% 
  group_by(test_type) %>% 
  mutate(frac = frac / max(frac)) %>% 
  ggplot(aes(dataset_id, test_type, size = frac, fill = frac))+
  geom_point(shape = 21) +
  scale_fill_gradient(low = "white", 
                      high = "orangered") +
  theme_bw()

enrichment_summary_graph <-
  plot_data %>% 
  mutate(test_type = gsub("specificity_", "", test_type),
         test_type = case_when(test_type == "GO_BP_simplified" ~ "GO BP",
                               test_type == "GO_CC_simplified" ~ "GO CC",
                               test_type == "GO_MF_simplified" ~ "GO MF",
                               test_type == "panglao_cellmarkers" ~ "Panglao",
                               test_type == "protein_class" ~ "Protein class",
                               test_type == "reactome" ~ "Reactome",
                               test_type == "secretome_location" ~ "Secretome",
                               test_type == "subcellular_location" ~ "Subcellular location",
                               test_type == "trrust" ~ "TRRUST", 
                               test_type == "tissue" ~ "Tissue", 
                               test_type == "celline" ~ "Cell lines",
                               test_type == "blood" ~ "Blood", 
                               test_type == "singlecell" ~ "Single cell",
                               test_type == "brain" ~ "Brain", 
                               T ~ test_type),
         dataset_id = case_when(dataset_id  == "tissue" ~ "Tissue", 
                                dataset_id  == "celline" ~ "Cell line",
                                dataset_id  == "blood" ~ "Blood", 
                                dataset_id  == "singlecell" ~ "Single cell",
                                dataset_id  == "brain" ~ "Brain")) %>% 
  as_tbl_graph() 


enrichment_summary_graph %>% 
  
  create_layout(layout = 'stress') %>% 
  
  ggraph() + 
  geom_edge_link() +
  geom_node_point(size = 3) + 
  geom_node_text(aes(label = name)) +
  theme_void() + 
  scale_edge_color_gradient(low = "white", 
                            high = "orangered") 

node_type_coords <- 
  enrichment_summary_graph %>% 
  as_tibble() %>% 
  mutate(type = case_when(name %in% c("Blood",                
                                      "Brain",          
                                      "Cell line",
                                      "Single cell",
                                      "Tissue") ~ "Section",
                          name %in% c("Protein class",
                                      "Secretome", 
                                      "Subcellular location") ~ "HPA",
                          T ~ "External")) %>% 
  mutate(y = unclass(factor(type, c("External", "Section", "HPA")))) %>% 
  group_by(type) %>% 
  mutate(x = 1:length(name),
         x = x - mean(x), 
         x = ifelse(type == "HPA",
                    x * 2,
                    x))  


graph_data <- 
  enrichment_summary_graph %>% 
  left_join(node_type_coords) %>% 
  activate(edges) %>% 
  mutate(edge_type = case_when(node_type_coords$type[from] == "Section" &
                                 node_type_coords$type[to] == "Section" ~
                                 "Specificity",
                               node_type_coords$type[from] == "External" |
                                 node_type_coords$type[to] == "External" ~
                                 "External",
                               node_type_coords$type[from] == "HPA" |
                                 node_type_coords$type[to] == "HPA" ~
                                 "HPA")) %>% 
  create_layout(layout = 'manual',
                x = node_type_coords$x,
                y = node_type_coords$y) 

edge_data <- get_edges()(graph_data)

plots <- 
  lapply(unique(edge_data$edge_type),
         function(edge_type_) {
           edge_data_ <- 
             edge_data %>% 
             filter(edge_type == edge_type_)
           
           plot <- 
             graph_data %>% 
             ggraph() + 
             # geom_edge_diagonal(aes(width = frac,
             #                    color = frac)) +
             geom_edge_fan(data = edge_data_ %>%
                             filter(edge_type == "Specificity"),
                           aes(color = frac),
                           # arrow = arrow(length = unit(4, 'mm')),
                           strength = 2,
                           alpha = 0.7,
                           show.legend = F, ) +
             geom_edge_diagonal(data = edge_data_ %>% 
                                  filter(edge_type != "Specificity"),
                                aes(color = frac),
                                alpha = 0.7,
                                show.legend = F) +
             geom_node_point(data = . %>% 
                               filter(type == "Section"),
                             aes(color = name),
                             size = 16,
                             show.legend = F) + 
             geom_node_text(data = . %>% 
                              filter(type == "Section"),
                            aes(label = name)) +
             geom_node_text(data = . %>% 
                              filter(type == "HPA"),
                            aes(label = name),
                            vjust = 0) +
             geom_node_text(data = . %>% 
                              filter(type == "External"),
                            aes(label = name),
                            vjust = 1) +
             theme_void() + 
             scale_edge_color_gradient(low = "white", 
                                       high = "black", 
                                       limits = c(0, 1)) +
             scale_color_manual(values = dataset_palette2) +
             scale_x_continuous(expand = expansion(0.1))
           
           ggsave(savepath(paste0("Database integration network ", 
                                  edge_type_, ".pdf")),
                  plot = plot,
                  width = 8, height = 4)
           ggsave(savepath(paste0("Database integration network ", 
                                  edge_type_, ".png")),
                  plot = plot,
                  width = 8, height = 4)
           
           
         })





```



# UMAP


## UMAP polygons


### Create polygons

```{r}

umap_plot_data <- 
  umap_data %>%  
  left_join(clustering_data)

generate_cluster_hulls


hull_settings <- 
  expand_grid(dataset_id = unique(umap_plot_data$dataset_id),
              relative_bandwidth_i = c(150),
              poly_concavity = c(2),
              poly_smoothing = c(1.25),
              n = 1000) %>% 
  mutate(relative_bandwidth = 1 / relative_bandwidth_i,
         i = row_number())


hull_data <- 
  pblapply(hull_settings$i,
         function(i) {
           
           plot_data <- 
             umap_plot_data %>% 
             filter(dataset_id == hull_settings$dataset_id[i])
           
           generate_cluster_hulls(V1 = plot_data$UMAP_1_scaled,
                                  V2 = plot_data$UMAP_2_scaled, 
                                  element_id = plot_data$gene,
                                  cluster_membership = plot_data$cluster, 
                                  n = hull_settings$n[i],
                                  poly_concavity = hull_settings$poly_concavity[i],
                                  poly_smoothing = hull_settings$poly_smoothing[i],
                                  relative_bandwidth = hull_settings$relative_bandwidth[i]) 
           
         })


hulls <- 
  hull_data %>% 
  map(. %>% {.$hulls}) %>% 
  bind_rows(.id = "i") %>% 
  left_join(hull_settings %>% 
              mutate(i = as.character(i)))

hull_centers <- 
  hulls %>% 
  filter(landmass == 1, sub_cluster == 1) %>% 
  group_by(dataset_id, cluster, polygon_id) %>% 
  summarise(X = mean(X), 
            Y = mean(Y)) %>% 
  ungroup()

hull_centers_all <- 
  hulls %>% 
  group_by(dataset_id, cluster, polygon_id) %>% 
  summarise(X = mean(X), 
            Y = mean(Y)) %>% 
  ungroup()

hull_segments <-
  hull_centers_all %>% 
  left_join(hull_centers,
            by = c("dataset_id", "cluster"),
            suffix = c("", "end")) %>% 
  filter(polygon_id != polygon_idend)
  
hulls %>% 
  ggplot(aes(X, Y, group = polygon_id, fill = cluster)) +
  geom_hex(data = umap_data %>%
             filter(dataset_id %in% hull_settings$dataset_id),
           aes(UMAP_1_scaled, UMAP_2_scaled, fill = 1),
           fill = "gray70",
           bins = 200,
           inherit.aes = F) +
  geom_segment(data = hull_segments,
               aes(X, Y, 
                   xend = Xend, 
                   yend = Yend,
                   group = cluster),
               inherit.aes = F) +
  geom_polygon(show.legend = F,
               color = "black") +
  geom_point(data = hull_centers,
             aes(X, Y), 
             inherit.aes = F) +
  # facet_wrap( ~ dataset_id + 
  #               relative_bandwidth_i + 
  #               poly_concavity + 
  #               poly_smoothing, nrow = 2) +
  facet_wrap( ~ dataset_id) +
  coord_fixed() +
  theme_bw() +
  
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank())


  
  select(dataset_id, cluster, sub_cluster, landmass) %>% 
  filter(cluster == 3, dataset_id == "tissue") %>% 
  distinct()
  group_by(dataset_id, )

############
############

all_graph_umap_hulls <- 
  all_graph_umap_data_scaled %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  # slice(6) %>% 
  # slice(1, 4, 6) %>% 
  pull(dataset_id) %>%
  set_names(., .) %>% 
  pblapply(function(dataset_id_) {
    all_graph_umap_data_scaled %>% 
      filter(dataset_id == dataset_id_) %$%
      generate_cluster_hulls(UMAP_1,
                             UMAP_2, 
                             element_id = gene,
                             cluster_membership = factor(cluster), 
                             bandwidth_fct = 5,
                             n = 1000)
  })

all_graph_umap_hull_data <- 
  all_graph_umap_hulls %>% 
  map(. %$% 
        hulls) %>% 
  bind_rows(.id = "dataset_id")


rm(all_graph_umap_hulls)
gc()
# all_graph_umap_hull_data %>% 
#   filter(density_type == "primary secondary") %>% 
#   select(1, 2, 3, 4, 5) %>% 
#   distinct()

all_graph_umap_hull_data %>% 
  filter(density_type == "primary secondary") %>%
  ggplot(aes(X,Y, group = paste(cluster, sub_cluster, landmass), fill = cluster)) +
  geom_polygon(show.legend = F,
               color = "black") +
  theme_bw() +
  facet_grid(~dataset_id) +
  coord_fixed()
ggsave(run_id_savename("UMAP scaled hulls.pdf"),
       width = 8, height = 5)


# Save
# all_graph_umap_hull_data %>% 
#   filter(density_type == "primary secondary") %>% 
#   unite(polygon_id, cluster, sub_cluster, landmass, remove = F) %>% 
#   select(polygon_id, cluster, UMAP1 = X, UMAP2 = Y) %>% 
#   write_csv(run_id_savename("UMAP draft polygon coordinates.csv"))
#   
# 
# cluster_hulls$data %>% 
#   select(ensg_id = element_id, 
#          UMAP1 = V1, 
#          UMAP2 = V2,
#          cluster) %>% 
#   write_csv(run_id_savename("UMAP draft gene coordinates.csv"))

```



### plotly

```{r}
umap_poly_centers <- 
  umap_poly_data %>% 
  filter(sub_cluster == 1) %>% 
  left_join(cluster_annotation) %>% 
  group_by(dataset_id, cluster, full_annotation) %>% 
  summarise(X = mean(X),
            Y = mean(Y)) %>% 
  ungroup() 

umap_poly_data_colored <- 
  umap_poly_data %>%
  left_join(umap_poly_data %>% 
              select(dataset_id, cluster) %>% 
              distinct() %>% 
              group_by(dataset_id) %>% 
              mutate(cluster_color = colorRampPalette(ggthemes::tableau_color_pal(palette = "Classic Cyclic")(13))(n_distinct(cluster))))

umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point(data = umap_data %>% 
               filter(dataset_id == "tissue"),
             aes(UMAP_1_scaled, 
                 UMAP_2_scaled),
             color = "black",
             size = 0.1,
             inherit.aes = F) +
  geom_polygon(show.legend = F,
               color = "black",
               size = 0.1) +
  geom_text(data = umap_poly_centers %>% 
              filter(dataset_id == "tissue") ,
            aes(X, Y, label = full_annotation),
            size = 3,
            inherit.aes = F) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

umap_poly_data_colored %>% 
  filter(dataset_id == "blood") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point(data = umap_data %>% 
               filter(dataset_id == "blood"),
             aes(UMAP_1_scaled, 
                 UMAP_2_scaled),
             color = "black",
             size = 0.1,
             inherit.aes = F) +
  geom_polygon(show.legend = F,
               color = "black",
               size = 0.1) +
  geom_text(data = umap_poly_centers %>% 
              filter(dataset_id == "blood") ,
            aes(X, Y, label = full_annotation),
            size = 3,
            inherit.aes = F) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

plots <-
  unique(umap_poly_data_colored$dataset_id) %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    umap_poly_data_colored %>% 
      filter(dataset_id == dataset_id_) %>% 
      ggplot(aes(X, Y, fill = cluster_color, 
                 group = paste(cluster, sub_cluster))) +
      geom_hex(data = umap_data %>% 
                 filter(dataset_id == dataset_id_),
               aes(UMAP_1_scaled, 
                   UMAP_2_scaled,
                   fill = "1"),
               fill = "darkgray",
               bins = 1000,
               inherit.aes = F) +
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      geom_text(data = umap_poly_centers %>% 
                  filter(dataset_id == dataset_id_) ,
                aes(X, Y, label = full_annotation),
                size = 5,
                inherit.aes = F) +
      scale_fill_identity() +
      coord_fixed() +
      theme_void() +
      ggtitle(dataset_id_)
    
    
    
  })

for(dataset_id_ in names(plots)) {
  
  plots[[dataset_id_]] %>% 
    
    ggplotly() %>% 
    hide_legend() %>% 
    as_widget() %>% 
    
    htmlwidgets::saveWidget(savepath(paste0("interactive UMAP ",
                                            dataset_id_,
                                            ".html")))
}

```

### Regular poly UMAPs

```{r}


plots <- 
  umap_poly_centers %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  pull() %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    plot_data <- 
      umap_poly_data_colored %>% 
      filter(dataset_id == dataset_id_) 
    
    
    plot_data %>% 
      ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>% 
                        filter(dataset_id == dataset_id_),
                      aes(UMAP_1_scaled, 
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      coord_fixed() +
      theme_void() +
      ggtitle(str_to_sentence(dataset_id_)) +
      theme(plot.title = element_text(hjust = 0.5))
    
  })




for(dataset_id_ in names(plots)) {
  
  ggsave(savepath(paste0("Poly UMAP ",
                         dataset_id_,
                         ".pdf")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
  
  ggsave(savepath(paste0("Poly UMAP ",
                         dataset_id_,
                         ".png")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
}

plot <- wrap_plots(plots, nrow = 1)

ggsave(savepath("Poly UMAP combined.pdf"),
       plot = plot,
       width = 20, height = 8)

ggsave(savepath("Poly UMAP combined.png"),
       plot = plot,
       width = 20, height = 8)


```



### Summarized annotation UMAPs

```{r}

UMAP_plot_hclusts <- 
  umap_poly_centers %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  pull() %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    umap_poly_centers %>% 
      filter(dataset_id == dataset_id_) %>% 
      unite(id, dataset_id, cluster, full_annotation) %>% 
      column_to_rownames("id") %>% 
      dist() %>% 
      hclust(method = "average")
  })


plots <- 
  umap_poly_centers %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  pull() %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    plot_spec_ann <-
      cutree(UMAP_plot_hclusts[[dataset_id_]], h = 0.4) %>% 
      enframe("id", "UMAP_group") %>% 
      separate(id, into = c("dataset_id", "cluster", "full_annotation"), 
               sep = "_") %>% 
      left_join(cluster_annotation %>% 
                  select(dataset_id, cluster, ann_specificity, ann_function)) %>% 
      left_join(umap_poly_centers) 
    
    plot_spec_ann_centers <- 
      plot_spec_ann%>% 
      group_by(dataset_id, UMAP_group, ann_specificity) %>% 
      summarise(X = median(X), 
                Y = median(Y))
    
    
    plot_data <- 
      umap_poly_data_colored %>% 
      inner_join(plot_spec_ann %>% 
                   select(1, 2, ann_specificity)) %>% 
      filter(dataset_id == dataset_id_) 
    
    plot_pal <- 
      complete_palette(pal_terms = unique(plot_data$ann_specificity),
                       pal = deframe(tissue_palette), 
                       default_pal = cluster_color_pal)
    
    plot_pal <- 
      complete_palette_neighbor(pal_terms = unique(plot_data$ann_specificity),
                                pal = deframe(tissue_palette))
    
    plot_data %>% 
      ggplot(aes(X, Y, fill = ann_specificity, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>% 
                        filter(dataset_id == dataset_id_),
                      aes(UMAP_1_scaled, 
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      geom_polygon(show.legend = F,
                   color = "black",
                   size = 0.1) +
      geom_text(data = plot_spec_ann_centers %>% 
                  filter(dataset_id == dataset_id_) ,
                aes(X, Y, label = ann_specificity),
                size = 3,
                inherit.aes = F) +
      scale_fill_manual(values = plot_pal) +
      coord_fixed() +
      theme_void() +
      ggtitle(str_to_sentence(dataset_id_)) +
      theme(plot.title = element_text(hjust = 0.5))
    
  })



for(dataset_id_ in names(plots)) {
  
  ggsave(savepath(paste0("Specificity grouped ann UMAP ",
                         dataset_id_,
                         ".pdf")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
}

plot <- wrap_plots(plots, nrow = 1)

ggsave(savepath("Specificity grouped ann UMAP combined.pdf"),
       plot = plot,
       width = 20, height = 8)


```

## UMAP example

### Tissue Cluster 22
```{r}

example_cluster <- 
  cluster_annotation %>% 
  filter(dataset_id == "tissue") %>% 
  filter(reliability == "High") %>% 
  filter(cluster == 22)

example_cluster_genes <- 
  clustering_data %>% 
  filter(dataset_id == "tissue",
         cluster == example_cluster$cluster)

example_gene <- 
  "ENSG00000111834"
################

p1 <- 
  umap_data %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y)) +
  geom_point_rast(aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  coord_fixed() +
  theme_void() 

p2 <-
  umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>% 
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  geom_polygon(show.legend = F,
               color = NA,
               size = 0.1,
               alpha = 0.8) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

p3 <- 
  umap_poly_data_colored %>% 
  filter(cluster == example_cluster$cluster) %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>% 
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  geom_polygon(show.legend = F,
               color = NA,
               size = 0.1,
               alpha = 0.8) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

p4 <-
  umap_poly_data_colored %>% 
  # filter(cluster == example_cluster$cluster) %>% 
  filter(dataset_id == "tissue") %>% 
  {ggplot(., aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>%
                        filter(dataset_id == "tissue"),
                      aes(UMAP_1_scaled,
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      scale_x_continuous(limits = scales::expand_range(range(.$X[.$cluster == example_cluster$cluster]),
                                                       2)) +
      scale_y_continuous(limits = scales::expand_range(range(.$Y[.$cluster == example_cluster$cluster]),
                                                       2)) +
      coord_fixed() +
      theme_void() }


ggsave(savepath("UMAP example plot p1.pdf"),
       width = 8, height = 8,
       plot = p1)
ggsave(savepath("UMAP example plot p2.pdf"),
       width = 8, height = 8,
       plot = p2)
ggsave(savepath("UMAP example plot p3.pdf"),
       width = 8, height = 8,
       plot = p3)
ggsave(savepath("UMAP example plot p4.pdf"),
       width = 8, height = 8,
       plot = p4)

ggsave(savepath("UMAP example plot p1.png"),
       width = 8, height = 8,
       plot = p1)
ggsave(savepath("UMAP example plot p2.png"),
       width = 8, height = 8,
       plot = p2)
ggsave(savepath("UMAP example plot p3.png"),
       width = 8, height = 8,
       plot = p3)
ggsave(savepath("UMAP example plot p4.png"),
       width = 8, height = 8,
       plot = p4)


###############################################################

p5 <-
  umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>%
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled,
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  
  geom_polygon(data = . %>% 
                 filter(cluster != example_cluster$cluster),
               show.legend = F,
               fill = "gray",
               color = "black",
               size = 0.1,
               alpha = 0.8) +
  geom_polygon(data = . %>% 
                 filter(cluster == example_cluster$cluster),
               show.legend = F,
               color = "black",
               size = 0.1,
               alpha = 0.8) +
  geom_point(data = umap_data %>% 
               filter(dataset_id == "tissue") %>% 
               filter(gene == example_gene),
             aes(UMAP_1_scaled, 
                 UMAP_2_scaled),
             color = "red",
             size = 3,
             inherit.aes = F) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

ggsave(savepath("UMAP example plot p5.pdf"),
       width = 8, height = 8,
       plot = p5)

ggsave(savepath("UMAP example plot p5.png"),
       width = 8, height = 8,
       plot = p5)

p6 <-
  umap_poly_data_colored %>% 
  filter(cluster == example_cluster$cluster) %>%
  filter(dataset_id == "tissue") %>% 
  {ggplot(., aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>%
                        filter(gene %in% example_cluster_genes$gene) %>% 
                        filter(dataset_id == "tissue"),
                      aes(UMAP_1_scaled,
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      scale_x_continuous(limits = scales::expand_range(range(.$X[.$cluster == example_cluster$cluster]),
                                                       1)) +
      scale_y_continuous(limits = scales::expand_range(range(.$Y[.$cluster == example_cluster$cluster]),
                                                       1)) +
      coord_fixed() +
      theme_void() }

ggsave(savepath("UMAP example plot p6.pdf"),
       width = 8, height = 8,
       plot = p6)

ggsave(savepath("UMAP example plot p6.png"),
       width = 8, height = 8,
       plot = p6)

```

### Tissue cluster 54

```{r}

example_cluster <- 
  cluster_annotation %>% 
  filter(dataset_id == "tissue") %>% 
  # filter(reliability == "High") %>% 
  filter(cluster == 54)

example_cluster_genes <- 
  clustering_data %>% 
  filter(dataset_id == "tissue",
         cluster == example_cluster$cluster)

example_gene <- 
  "ENSG00000004864"
################

p1 <- 
  umap_data %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y)) +
  geom_point_rast(aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  coord_fixed() +
  theme_void() 

p2 <-
  umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>% 
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  geom_polygon(show.legend = F,
               color = NA,
               size = 0.1,
               alpha = 0.8) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

p3 <- 
  umap_poly_data_colored %>% 
  filter(cluster == example_cluster$cluster) %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>% 
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  geom_polygon(show.legend = F,
               color = NA,
               size = 0.1,
               alpha = 0.8) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

p4 <-
  umap_poly_data_colored %>% 
  # filter(cluster == example_cluster$cluster) %>% 
  filter(dataset_id == "tissue") %>% 
  {ggplot(., aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>%
                        filter(dataset_id == "tissue"),
                      aes(UMAP_1_scaled,
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      scale_x_continuous(limits = scales::expand_range(range(.$X[.$cluster == example_cluster$cluster]),
                                                       2)) +
      scale_y_continuous(limits = scales::expand_range(range(.$Y[.$cluster == example_cluster$cluster]),
                                                       2)) +
      coord_fixed() +
      theme_void() }


ggsave(savepath("UMAP example plot p1.pdf"),
       width = 8, height = 8,
       plot = p1)
ggsave(savepath("UMAP example plot p2.pdf"),
       width = 8, height = 8,
       plot = p2)
ggsave(savepath("UMAP example plot p3.pdf"),
       width = 8, height = 8,
       plot = p3)
ggsave(savepath("UMAP example plot p4.pdf"),
       width = 8, height = 8,
       plot = p4)

ggsave(savepath("UMAP example plot p1.png"),
       width = 8, height = 8,
       plot = p1)
ggsave(savepath("UMAP example plot p2.png"),
       width = 8, height = 8,
       plot = p2)
ggsave(savepath("UMAP example plot p3.png"),
       width = 8, height = 8,
       plot = p3)
ggsave(savepath("UMAP example plot p4.png"),
       width = 8, height = 8,
       plot = p4)


###############################################################

p5 <-
  umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>%
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled,
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  
  geom_polygon(data = . %>% 
                 filter(cluster != example_cluster$cluster),
               show.legend = F,
               fill = "gray",
               color = "black",
               size = 0.1,
               alpha = 0.8) +
  geom_polygon(data = . %>% 
                 filter(cluster == example_cluster$cluster),
               show.legend = F,
               color = "black",
               size = 0.1,
               alpha = 0.8) +
  geom_point(data = umap_data %>% 
               filter(dataset_id == "tissue") %>% 
               filter(gene == example_gene),
             aes(UMAP_1_scaled, 
                 UMAP_2_scaled),
             color = "red",
             size = 3,
             inherit.aes = F) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

ggsave(savepath("UMAP example plot p5.pdf"),
       width = 8, height = 8,
       plot = p5)

ggsave(savepath("UMAP example plot p5.png"),
       width = 8, height = 8,
       plot = p5)

p6 <-
  umap_poly_data_colored %>% 
  filter(cluster == example_cluster$cluster) %>%
  filter(dataset_id == "tissue") %>% 
  {ggplot(., aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>%
                        filter(gene %in% example_cluster_genes$gene) %>% 
                        filter(dataset_id == "tissue"),
                      aes(UMAP_1_scaled,
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      scale_x_continuous(limits = scales::expand_range(range(.$X[.$cluster == example_cluster$cluster]),
                                                       1)) +
      scale_y_continuous(limits = scales::expand_range(range(.$Y[.$cluster == example_cluster$cluster]),
                                                       1)) +
      coord_fixed() +
      theme_void() }

ggsave(savepath("UMAP example plot p6.pdf"),
       width = 8, height = 8,
       plot = p6)

ggsave(savepath("UMAP example plot p6.png"),
       width = 8, height = 8,
       plot = p6)

```



# Bubble heatmap

```{r}

plot_data <-
  enrichment_data %>%
  filter(!test_type %in% c("KEGG",
                           "GO_BP_original",
                           "GO_CC_original",
                           "GO_MF_original",
                           "GO_BP_simplified",
                           "GO_CC_simplified",
                           "GO_MF_simplified",
                           "reactome")) %>% 
  filter(grepl("specificity", test_type)) %>% 
  left_join(cluster_annotation %>% 
              select(dataset_id, cluster, full_annotation), sep = " ") %>% 
  unite(cluster, dataset_id, cluster, full_annotation) %>% 
  select(cluster, ID, odds_ratio) %>% 
  group_by_all() %>% 
  mutate(capped_odds_ratio = min(c(odds_ratio,
                                   50))) %>% 
  ungroup()

cluster_levels <- 
  clustering_data %>% 
  select(dataset_id, cluster) %>% 
  distinct() %>% 
  arrange(dataset_id, as.numeric(cluster)) %>% 
  left_join(cluster_annotation %>% 
              select(dataset_id, cluster, full_annotation), sep = " ") %>% 
  unite(cluster, dataset_id, cluster, full_annotation)



plot_size_range <- c(1, 4)

plot_legend_settings <- 
  tibble(odds_ratio_str = c(1, 10, 20, 30, 40, ">50"),
         y = 1:6,
         odds_ratio = c(1, 10, 20, 30, 40, 50)) 

plot_legend <-
  plot_legend_settings %>% 
  ggplot(aes(1, y, 
             size = odds_ratio,
             fill = odds_ratio,
             label = odds_ratio_str)) +
  geom_point(shape = 21,
             show.legend = F) +
  geom_text(size = 4, 
            hjust = 0,
            nudge_x = 0.1) +
  annotate("text", x = 1, y = 7, label = "Odds ratio",
           fontface = "bold", hjust = 0) +
  scale_x_continuous(limits = c(0, 2)) +
  scale_y_continuous(limits = c(0, 7)) +
  scale_fill_gradient(name = "Odds ratio",
                      low = "white", high = "orangered",
                      breaks = plot_legend_settings$odds_ratio,
                      limits = range(plot_legend_settings$odds_ratio)) + 
  scale_size_continuous(name = "Odds ratio", 
                        range = plot_size_range, 
                        breaks = plot_legend_settings$odds_ratio,
                        limits = range(plot_legend_settings$odds_ratio)) +
  theme_void()

plot_clustering <- 
  plot_data %>% 
  select(cluster, ID, odds_ratio) %>% 
  spread(ID, odds_ratio, fill = 0) %>% 
  column_to_rownames("cluster") %>%
  list(tissue = t(.), 
       cluster = .) %>%
  map(. %>% 
        dist(method = "binary") %>% 
        hclust(method = "ward.D2"))

plot_order <- 
  list(cluster = unique(c(with(plot_clustering$cluster,
                               labels[order]),
                          cluster_levels$cluster)),
       ID = with(plot_clustering$tissue,
                 labels[order]))

plot <-
  plot_data %>%
  ungroup() %>% 
  mutate(cluster = factor(cluster,
                          plot_order$cluster),
         ID = factor(ID,
                     plot_order$ID)) %>%
  ggplot(aes(ID, cluster, 
             size = capped_odds_ratio, 
             fill = capped_odds_ratio)) +
  geom_point(data = expand_grid(cluster = plot_order$cluster,
                                ID = plot_order$ID) %>%
               mutate(cluster = factor(cluster, unique(cluster)),
                      ID = factor(ID, unique(ID))),
             fill = NA,
             size = NA) +
  geom_point(shape = 21,
             show.legend = F) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0.5),
        legend.position = "right") +
  # coord_fixed() +
  scale_fill_gradient(name = "Odds ratio",
                      low = "white", high = "orangered",
                      breaks = plot_legend_settings$odds_ratio,
                      limits = range(plot_legend_settings$odds_ratio)) + 
  scale_size_continuous(name = "Odds ratio", 
                        range = plot_size_range, 
                        breaks = plot_legend_settings$odds_ratio,
                        limits = range(plot_legend_settings$odds_ratio)) +
  theme(axis.title = element_blank()) + 
  coord_fixed()

ggsave(savepath("Total Specificity Bubble heatmap.pdf"),
       plot = plot,
       width = 35, height = 35)


```


## Bubble network

```{r}
jaccard <- function(a, b) {
  intersection = length(intersect(a, b))
  union = length(a) + length(b) - intersection
  return (intersection/union)
}


# bubble_graph <- 
bubble_data <- 
  enrichment_data %>%
  filter(dataset_id != "brain") %>% 
  filter(test_type != "specificity_brain") %>% 
  
  # filter(test_type == "specificity_tissue") %>% 
  filter(grepl("specificity", test_type)) %>%
  filter(p.adjust < 0.01) %>% 
  left_join(cluster_annotation %>% 
              select(dataset_id, cluster, full_annotation), sep = " ") %>% 
  unite(cluster, dataset_id, cluster, full_annotation) %>% 
  select(cluster, ID, odds_ratio)  

bubble_dist <- 
  bubble_data %>% 
  spread(ID, odds_ratio, fill = 0) %>% 
  column_to_rownames("cluster") %>% 
  dist(method = "binary") %>%
  as.matrix() %>% 
  as_tibble(rownames = "cluster1") %>% 
  gather(cluster2, dist, -1) %>% 
  filter(cluster1 != cluster2)

bubble_dist %>% 
  ggplot(aes(dist)) +
  geom_density()

bubble_nn_data <- 
  bubble_dist %>%
  filter(dist < 0.75) %>% 
  group_by(cluster1) %>%
  top_n(5, -dist) %>%
  arrange(cluster1)


bubble_graph <- 
  bubble_nn_data %>% 
  as_tbl_graph() %>% 
  left_join(as_tibble(.) %>% 
              separate(name, into = c("dataset_id", "cluster", "full_annotation"),
                       sep = "_",
                       remove = F))
bubble_graph %>% 
  ggraph('fabric', sort.by = node_rank_fabric()) + 
  geom_node_range(colour = 'grey') + 
  geom_edge_span(end_shape = 'square') + 
  coord_fixed() +
  theme_void()


# umap_layout <- 
#   bubble_graph %>% 
#   activate(edges) %>% 
#   as_tibble() %>% 
#   spread(to, dist, fill = 1) %>% 
#   column_to_rownames("from") %>% 
#   umap(n_neighbors = 30) %>% 
#   as_tibble()

bubble_graph_layout <- 
  bubble_graph %>% 
  # create_layout(layout = 'manual', 
  #               x = umap_layout$V1,
  #               y = umap_layout$V2)
  create_layout(layout = 'fr')



bubble_graph_layout %>% 
  ggraph() + 
  geom_edge_link(color = "gray") +
  geom_node_point(aes(color = dataset_id),
                  size = 3) + 
  # geom_node_text(aes(label = full_annotation),
  #                size = 2) +
  
  theme_void() +
  scale_color_manual(values = dataset_palette) +
  coord_fixed()

ggsave(savepath("interactive specificity net.pdf"),
       width = 8, height = 8)
ggsave(savepath("interactive specificity net.png"),
       width = 8, height = 8)

plot <-
  bubble_graph_layout %>%
  as_tibble() %>% 
  ggplot(aes(x, y, label = full_annotation, color = dataset_id)) +
  geom_segment(data = get_edges()(bubble_graph_layout) %>% 
                 as_tibble(),
               aes(x = x,
                   y = y, 
                   xend = xend, 
                   yend = yend),
               size = 0.2,
               inherit.aes = F) +
  geom_point(size = 5) +
  geom_text(size = 3,
            color = "black") +
  theme_void() +
  scale_color_manual(values = dataset_palette) +
  coord_fixed()

plot %>% 
  ggplotly() %>% 
  as_widget() %>% 
  htmlwidgets::saveWidget(savepath("interactive specificity net.html"))

```


