---
title: "Data visualization"
author: "María Bueno Alvez & Max J Karlsson"
date: "01/10/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

#Placeholder

#Setup

##Load libraries

```{r setup, include=FALSE}

library(tidyverse)
library(pbapply)
library(readxl)
library(multidplyr)
library(clusterProfiler)

library(tidygraph)
library(ggraph)
library(ggrepel)
library(ggthemes)
library(ggrastr)
library(ggforce)
library(ggupset)
library(plotly)
library(patchwork)
library(ggalluvial)
library(uwot)
library(aricode)

source("scripts/functions_utility.R")
source("scripts/functions_annotation.R")
source("scripts/cluster rasterizer.R")
source("../HPA-visualization/scripts/HPA_visualization_functions.R")

select <- dplyr::select


```

## File structure
```{r}

dataset_metadata <- read_csv("run_settings/20210928_settings.csv") %>% 
  filter(dataset != "brain")


data_paths <- read_csv("run_settings/20211108 all_datasets.csv")

enrichment_settings <- read_csv("run_settings/20211013_enrichment_settings.csv")

file_structure <-
  create_folder_structure(dataset_id = dataset_metadata$dataset_id,
                          dataset_run_id = dataset_metadata$version,
                          main_folder = "results/Clustering_results",
                          folders = c("data" = "data", 
                                      "svg" = "svg",
                                      "UMAP" = "UMAP",
                                      "treemap" = "svg/treemap",
                                      "heatmap" = "svg/heatmap",
                                      "enrichment" = "enrichment",
                                      "clustering" = "clustering"))
```


## Load clustering data

```{r}

clustering_data <- 
  file_structure %>% 
  map(function(x) x$clustering) %>% 
  map(function(x) {
    file_ <- paste(x, "final_consensus.tsv", sep = "/")
    if(file.exists(file_)) {
      read_tsv(file_)
    } else {
      NULL
    }
  }) %>% 
  bind_rows(.id = "dataset_id") %>% 
  mutate(cluster = as.character(cluster)) 


```

## Load enrichment data

```{r}

enrichment_data <- 
  file_structure %>% 
  map(function(x) x$enrichment) %>% 
  map(function(x) {
    file_ <- paste(x, "enrichment_results.tsv", sep = "/")
    if(file.exists(file_)) {
      read_tsv(file_)
    } else {
      NULL
    }
  }) %>% 
  bind_rows(.id = "dataset_id") %>% 
  mutate(cluster = as.character(cluster)) 

```

## Load annotation

```{r}

cluster_annotation <-
  # list(tissue = read_csv("data/annotation/Tissue 20211013.csv"),
  #      blood = read_csv("data/annotation/Blood 20211013.csv"),
  #      celline = read_csv("data/annotation/Celline 20211013.csv"),
  #      brain = read_csv("data/annotation/Brain 20211013.csv")) %>% 
  list.files("data/annotation", full.names = T) %>% 
  {.[grepl("20211124", .)]} %>% 
  set_names(., gsub(".*\\d |\\.xlsx$", "", .)) %>% 
  map(. %>% read_excel()) %>% 
  bind_rows(.id = "dataset_id") %>%
  filter(dataset_id != "brain") %>% 
  rename(dataset_id = 1,
         cluster = 2,
         n_genes = 3,
         ann_specificity = 5, 
         ann_function = 6,
         reliability = 7,
         external_comment = 8,
         internal_comment = 9) %>% 
  mutate(cluster = as.character(cluster),
         full_annotation = paste(ann_specificity, 
                                 ann_function, 
                                 sep = " - "))

```


## Load UMAP data 

```{r}
umap_data <- 
  file_structure %>% 
  map(function(x) x$UMAP) %>% 
  map(function(x) {
    file_ <- paste(x, "UMAP.tsv", sep = "/")
    if(file.exists(file_)) {
      read_tsv(file_)
    } else {
      NULL
    }
  }) %>% 
  bind_rows(.id = "dataset_id") 

umap_poly_data <- 
  file_structure %>% 
  map(function(x) x$UMAP) %>% 
  map(function(x) {
    file_ <- paste(x, "UMAP_polygons.tsv", sep = "/")
    if(file.exists(file_)) {
      read_tsv(file_)
    } else {
      NULL
    }
  }) %>% 
  bind_rows(.id = "dataset_id") %>% 
  mutate(cluster = as.character(cluster)) 

umap_centers <- 
  file_structure %>% 
  map(function(x) x$UMAP) %>% 
  map(function(x) {
    file_ <- paste(x, "cluster_centers.tsv", sep = "/")
    if(file.exists(file_)) {
      read_tsv(file_)
    } else {
      NULL
    }
  }) %>% 
  bind_rows(.id = "dataset_id") %>% 
  mutate(cluster = as.character(cluster)) 

```

## Load raw data

```{r}

raw_data <-
  readRDS("data/processed/combined_HPA_expression_data.RDS")


consensus_data <- 
  raw_data[which(grepl("consensus$", names(raw_data)))] %>% 
  set_names(gsub("_consensus", "", names(.)))

```


## Load specificity classification & Tau

```{r}

gene_classification <- 
  enrichment_settings %>% 
  filter(grepl("specificity", id)) %>% 
  filter(!grepl("brain", id)) %>% 
  group_by(id) %>% 
  do({
    g_data <<- .
    
    get_specificity_db(enrichment_settings, g_data$id)
    
  }) %>% 
  ungroup() %>% 
  mutate(dataset_id = gsub("specificity_", "", id)) %>% 
  select(-id)

gene_classification_long <- 
  gene_classification %>% 
  mutate(enhanced_tissues = gsub(", ", ";", enhanced_tissues)) %>% 
  separate_rows(enhanced_tissues, sep = ",") %>% 
  mutate(enhanced_tissues = gsub(";", ", ", enhanced_tissues),
         enhanced_tissues = trimws(enhanced_tissues)) 

gene_consensus_tau <- 
  consensus_data %>% 
  map(. %>% 
        column_to_rownames("ensg_id") %>% 
        calculate_tau_score()) %>% 
  bind_rows(.id = "dataset_id") %>% 
  inner_join(gene_classification %>% 
               filter(specificity_category != "Not detected") %>% 
               select(dataset_id, 
                      gene = ensg_id))

write_csv(gene_consensus_tau,
          savepath("Gene Tau score.csv"))

gene_consensus_tau %>% 
  ggplot(aes(dataset_id, tau_score)) +
  geom_violin()
```



## Load other metadata


### Colors

```{r}

plot_datasets <- 
  c("tissue",
    "singlecell",
    "blood",
    "celline") 

# dataset_palette <- 
#   c("tissue" = "#0083C4",  
#     "brain" = "#FFDD00", 
#     "blood" = "#CF161A", 
#     "singlecell" = "#6AA692", 
#     "celline" = "#97CF16")
# 
# dataset_palette2 <- 
#   c("Tissue" = "#0083C4",  
#     "Brain" = "#FFDD00", 
#     "Blood" = "#CF161A", 
#     "Single cell" = "#6AA692", 
#     "Cell line" = "#97CF16")

dataset_palette <- 
  c(celline = "#ED446D",
    blood = "#FFA93A",
    tissue = "#20639A",
    singlecell = "#3BAEA1")

dataset_palette %>% 
  prismatic::color()
region_colors <- 
  read_tsv("data/colors/colors_regions.tsv")

consensus_colors <- 
  read_tsv("data/colors/colors_consensus.tsv")

region_palette <- 
  region_colors %>% 
  select(sample, color) %>% 
  deframe()

consensus_palette <- 
  consensus_colors %>% 
  select(sample, color) %>% 
  deframe()

# all_HPA_colors <- 
#   read_tsv("data/meta/HPA21_colors.txt", 
#            col_names = c("organ_name", "tissue_name", "color")) 

# organ_palette <- 
#   all_HPA_colors %>% 
#   select(organ_name, color) %>% 
#   distinct() %>% 
#   deframe()

# tissue_palette <- 
#   all_HPA_colors %>% 
#   filter(organ_name != "Brain") %>% 
#   gather(tissue_type, tissue, 1, 2) %>% 
#   select(tissue, color) %>% 
#   distinct() %>% 
#   bind_rows(mutate(.,
#                    tissue = str_to_sentence(tissue))) %>% 
#   bind_rows(cluster_annotation %>% 
#               select(ann_specificity) %>% 
#               distinct() %>% 
#               filter(grepl("^Low.*specificity$", ann_specificity)) %>% 
#               rename(tissue = 1) %>% 
#               mutate(color = "darkgray")) %>% 
#   bind_rows(all_HPA_colors %>% 
#               filter(organ_name == "Brain") %>% 
#               select(tissue = organ_name, color) %>% 
#               distinct())

# tissue_palette %>% group_by(tissue) %>% 
#   count %>% 
#   filter(n > 1)
#   
# all_HPA_colors %>% 
#   filter(tissue_name == "anterior cingulate gyrus, pregenual-dorsal") 
#   pull(color) %>% 
#   prismatic::color()
# 
# read_tsv("data/meta/HPA21_colors.txt", 
#            col_names = c("organ_name", "tissue_name", "color")) %>% 
#   filter(tissue_name == "amygdala") %>% 
#   pull(color) %>% 
#   
# 
# all_HPA_colors %>% 
#   group_by(tissue) %>% 
#   count %>% 
#   filter(n > 1) %>% 
#   View

cluster_color_pal <-
  tableau_color_pal(palette = "Classic Cyclic",
                    type = 'regular')(13) %>%
  colorRampPalette()


gene_category_pal <- 
  c("tissue enriched" = "#e41a1c",
    "group enriched" = "#FF9D00",
    "tissue enhanced" = "#984ea3",
    "low tissue specificity" = "grey40",
    
    "detected in all" = "#253494",
    "detected in many" = "#2c7fb8",
    "detected in some" = "#41b6c4",
    "detected in single" = "#a1dab4",
    
    "not detected" = "grey", 
    "not detected " = "grey", 
    
    "Tissue enriched" = "#e41a1c",
    "Group enriched" = "#FF9D00",
    "Tissue enhanced" = "#984ea3",
    "Low tissue specificity" = "grey40",
    
    "Detected in all" = "#253494",
    "Detected in many" = "#2c7fb8",
    "Detected in some" = "#41b6c4",
    "Detected in single" = "#a1dab4",
    
    "Not detected" = "grey", 
    "Not detected " = "grey")


spec_category_pal <- 
  gene_category_pal %>% 
  enframe() %>% 
  filter(name %in% gene_classification$specificity_category) %>% 
  deframe()

spec_category_levels <- 
  c('tissue enriched',
    'group enriched', 
    'tissue enhanced', 
    'low tissue specificity',  
    'not detected',
    
    'Tissue enriched',
    'Group enriched', 
    'Tissue enhanced', 
    'Low tissue specificity',  
    'Not detected')

dist_category_levels <- 
  c('detected in all', 
    'detected in many', 
    'detected in some', 
    'detected in single', 
    'not detected',
    
    'Detected in all', 
    'Detected in many', 
    'Detected in some', 
    'Detected in single', 
    'Not detected')

protein_loc_palette <- c("membrane" = "#6DB9C6",
                         "intracellular" = "#FCAC3B",
                         "secreted" = 	"#CE70A4",
                         "membrane/secreted" = "#CF5734")

protein_loc_palette %>% 
  prismatic::color()
```

### Info

```{r}

gene_info <- 
  read_tsv("data/meta/geneinfo_103.tsv")

gene_names <- 
  gene_info %>% 
  select(ensg_id, gene_name)


membr_secr_genes <- 
  read_tsv("data/meta/mem_sec_20090.txt")

```


# Data overview

## Raw data

```{r}
raw_data %>% 
  map(. %>% 
        select(-1) %>% 
        ncol) %>% 
  unlist() %>% 
  enframe()
```

## Clustering

```{r}

raw_data$tissue_consensus %>% 
  gather(sample, value, -1) %>% 
  filter(value < 0.1, 
         value > 0)

```

## Hierarchical clustering

```{r}


consensus_clust_cor_average <- 
  consensus_data %>% 
  map(. %>% 
        column_to_rownames("ensg_id") %>% 
        cor(method = "spearman", use = "pairwise.complete.obs") %>% 
        {1 - .} %>% 
        as.dist() %>% 
        hclust(method = "average"))


consensus_clust_cor_ward <- 
  consensus_data %>% 
  map(. %>% 
        column_to_rownames("ensg_id") %>% 
        cor(method = "spearman", use = "pairwise.complete.obs") %>% 
        {1 - .} %>% 
        as.dist() %>% 
        hclust(method = "ward.D2"))

consensus_clust_labelorder <- 
  list(cor_average = consensus_clust_cor_average,
       cor_ward = consensus_clust_cor_ward) %>% 
  map(. %>%
        map(. %>% 
              with(labels[order])))
```



## Annotation

```{r}

cluster_annotation %>% 
  select(ann_specificity1 = ann_specificity) %>% 
  distinct() %>% 
  expand_grid(ann_specificity2 = .$ann_specificity1) %>% 
  mutate(sim = 1 - stringdist::stringsim(ann_specificity1, ann_specificity2)) %>% 
  filter(ann_specificity1 < ann_specificity2) %>% 
  filter(sim < 0.3)%>% 
  arrange(sim)

cluster_annotation %>% 
  select(ann_specificity1 = ann_specificity) %>% 
  distinct() %>% 
  expand_grid(ann_specificity2 = .$ann_specificity1) %>% 
  mutate(sim = 1 - stringdist::stringsim(ann_specificity1, ann_specificity2)) %>% 
  filter(sim < 0.3) %>% 
  filter(ann_specificity1 != ann_specificity2) %>% 
  spread(ann_specificity2, sim, fill = 1) %>% 
  column_to_rownames("ann_specificity1") %>% 
  pheatmap::pheatmap()

cluster_annotation %>% 
  select(ann_function1 = ann_function ) %>% 
  distinct() %>% 
  expand_grid(ann_function2 = .$ann_function1) %>% 
  mutate(sim = 1 - stringdist::stringsim(ann_function1, ann_function2)) %>% 
  filter(ann_function1 < ann_function2) %>% 
  filter(sim < 0.3) %>% 
  arrange(sim)

cluster_annotation %>% 
  filter(grepl("&", full_annotation)) 

cluster_annotation %>% 
  filter(grepl(" and ", full_annotation))

cluster_annotation %>% 
  filter(grepl("-", ann_specificity)) %>% 
  select(ann_specificity) %>% 
  distinct() 

cluster_annotation %>% 
  filter(grepl("-", ann_function)) %>% 
  select(ann_function) %>% 
  distinct() 



##################################

annotation_status <-
  cluster_annotation %>% 
  select(dataset_id, cluster, n_genes, ann_specificity, ann_function) %>% 
  mutate(function_status = case_when(ann_function %in% c("Unknown function",
                                                         "Mixed function") ~
                                       ann_function,
                                     T ~ "Annotated function"),
         specificity_status = case_when(ann_specificity == "Non-specific" ~ 
                                          "Low specificity",
                                        T ~ "Annotated specificity"))

plot_data <- annotation_status

p1 <- 
  plot_data %>% 
  group_by(dataset_id, function_status) %>% 
  count() %>% 
  ggplot(aes(n, dataset_id, fill = function_status, label = n)) +
  geom_col() + 
  geom_text(position = position_stack(vjust = 0.5),
            color = "white") +
  geom_text(data = . %>% 
              group_by(dataset_id) %>% 
              summarise(n = sum(n)),
            inherit.aes = F,
            aes(n, dataset_id, label = n),
            hjust = 0) +
  theme_void() +
  theme(axis.text.y = element_text(hjust = 1), 
        legend.position = "top",
        legend.direction = "vertical") +
  scale_x_continuous(expand = expansion(c(0, 0.2))) +
  scale_fill_manual(values = c("Annotated function" = "#647D96",
                               "Mixed function" = "gray",
                               "Unknown function" = "darkgray"),
                    name = "") +
  ggtitle("Function annotation")

p2 <- 
  plot_data %>% 
  group_by(dataset_id, specificity_status) %>% 
  count() %>% 
  ggplot(aes(n, dataset_id, fill = specificity_status, label = n)) +
  geom_col() + 
  geom_text(position = position_stack(vjust = 0.5),
            color = "white") +
  geom_text(data = . %>% 
              group_by(dataset_id) %>% 
              summarise(n = sum(n)),
            inherit.aes = F,
            aes(n, dataset_id, label = n),
            hjust = 0) +
  theme_void() +
  theme(axis.text.y = element_text(hjust = 1), 
        legend.position = "top", 
        legend.direction = "vertical") +
  scale_x_continuous(expand = expansion(c(0, 0.2))) +
  scale_fill_manual(values = c("Annotated specificity" = "#647D96",
                               "Low specificity" = "darkgray"),
                    name = "") +
  ggtitle("Specificity annotation")

p2 | p1

ggsave(savepath("Annotation status.pdf"), 
       width = 7, height = 4)
ggsave(savepath("Annotation status.png"), 
       width = 7, height = 4)

#####

p1 <- 
  plot_data %>% 
  group_by(dataset_id, function_status) %>% 
  summarise(n = sum(n_genes)) %>% 
  ggplot(aes(n, dataset_id, fill = function_status, label = n)) +
  geom_col() + 
  geom_text(position = position_stack(vjust = 0.5),
            color = "white",
            angle = -90) +
  geom_text(data = . %>% 
              group_by(dataset_id) %>% 
              summarise(n = sum(n)),
            inherit.aes = F,
            aes(n, dataset_id, label = n),
            hjust = 0) +
  theme_void() +
  theme(axis.text.y = element_text(hjust = 1), 
        legend.position = "top",
        legend.direction = "vertical") +
  scale_x_continuous(expand = expansion(c(0, 0.2))) +
  scale_fill_manual(values = c("Annotated function" = "#647D96",
                               "Mixed function" = "gray",
                               "Unknown function" = "darkgray"),
                    name = "") +
  ggtitle("Function annotation")

p2 <- 
  plot_data %>% 
  group_by(dataset_id, specificity_status) %>% 
  summarise(n = sum(n_genes)) %>% 
  ggplot(aes(n, dataset_id, fill = specificity_status, label = n)) +
  geom_col() + 
  geom_text(position = position_stack(vjust = 0.5),
            color = "white", 
            angle = -90) +
  geom_text(data = . %>% 
              group_by(dataset_id) %>% 
              summarise(n = sum(n)),
            inherit.aes = F,
            aes(n, dataset_id, label = n),
            hjust = 0) +
  theme_void() +
  theme(axis.text.y = element_text(hjust = 1), 
        legend.position = "top", 
        legend.direction = "vertical") +
  scale_x_continuous(expand = expansion(c(0, 0.2))) +
  scale_fill_manual(values = c("Annotated specificity" = "#647D96",
                               "Low specificity" = "darkgray"),
                    name = "") +
  ggtitle("Specificity annotation")

p2 | p1

ggsave(savepath("Annotation status genewise.pdf"), 
       width = 7, height = 4)
ggsave(savepath("Annotation status genewise.png"), 
       width = 7, height = 4)



plot_data %>% 
  group_by(function_status) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(frac = n / sum(n))

plot_data %>% 
  group_by(specificity_status) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(frac = n / sum(n))

plot_data %>% 
  group_by(function_status) %>% 
  summarise(n = sum(n_genes)) %>% 
  ungroup() %>% 
  mutate(frac = n / sum(n))

plot_data %>% 
  group_by(dataset_id, specificity_status) %>% 
  summarise(n = sum(n_genes)) %>% 
  group_by(dataset_id) %>% 
  # ungroup() %>% 
  mutate(frac = n / sum(n)) %>% 
  filter(specificity_status == "Annotated specificity")


#########################


clustering_data %>% 
  left_join(annotation_status) %>% 
  group_by(gene, specificity_status) %>% 
  summarise(datasets = list(sort(dataset_id))) %>% 
  
  # mutate(datasets)
  group_by(specificity_status, datasets) %>% 
  count() %>% 
  group_by(datasets) %>% 
  mutate(tot_n = sum(n)) %>% 
  ungroup() %>% 
  arrange(tot_n) %>% 
  filter(specificity_status == "Low specificity") %>% 
  group_by_all() %>% 
  mutate(n_dat = length(datasets[[1]])) %>% 
  ungroup() %>% 
  ggplot(aes(datasets, n, fill = specificity_status)) +
  geom_col() +
  scale_x_upset(n_intersections = 30) 

clustering_data %>% 
  left_join(annotation_status) %>% 
  select(dataset_id, gene, cluster, function_status, specificity_status) %>% 
  group_by(gene, specificity_status) %>% 
  count() %>% 
  group_by(specificity_status, n) %>% 
  count() %>%
  filter(specificity_status == "Low specificity") %>% 
  ggplot(aes(n, nn, fill = n, label = nn)) +
  geom_col() +
  geom_text(position = position_stack(), 
            vjust = 0) +
  scale_fill_gradient(low = "gray80", high = "gray20") +
  labs(x = "Number of datasets gene is in low specificity clustering") +
  theme_bw()

plot_data

```

## Enrichment data

```{r}

enrichment_data %>% 
  filter(test_type == "reactome") %>% 
  group_by(dataset_id, cluster) %>% 
  count %>% 
  arrange(-n)

plot_data <- 
  enrichment_data %>% 
  select(dataset_id, test_type, cluster) %>% 
  distinct() %>% 
  group_by(dataset_id, test_type) %>% 
  count() %>% 
  filter(!(grepl("GO", test_type) & grepl("original", test_type))) %>% 
  left_join(clustering_data %>% 
              group_by(dataset_id) %>% 
              summarise(n_cluster = n_distinct(cluster))) %>% 
  mutate(frac = n / n_cluster)


plot_data %>% 
  group_by(test_type) %>% 
  mutate(frac = frac / max(frac)) %>% 
  ggplot(aes(dataset_id, test_type, size = frac, fill = frac))+
  geom_point(shape = 21) +
  scale_fill_gradient(low = "white", 
                      high = "orangered") +
  theme_bw()

enrichment_summary_graph <-
  plot_data %>% 
  mutate(test_type = gsub("specificity_", "", test_type),
         test_type = case_when(test_type == "GO_BP_simplified" ~ "GO BP",
                               test_type == "GO_CC_simplified" ~ "GO CC",
                               test_type == "GO_MF_simplified" ~ "GO MF",
                               test_type == "panglao_cellmarkers" ~ "Panglao",
                               test_type == "protein_class" ~ "Protein class",
                               test_type == "reactome" ~ "Reactome",
                               test_type == "secretome_location" ~ "Secretome",
                               test_type == "subcellular_location" ~ "Subcellular location",
                               test_type == "trrust" ~ "TRRUST", 
                               test_type == "tissue" ~ "Tissue", 
                               test_type == "celline" ~ "Cell lines",
                               test_type == "blood" ~ "Blood", 
                               test_type == "singlecell" ~ "Single cell",
                               test_type == "brain" ~ "Brain", 
                               T ~ test_type),
         dataset_id = case_when(dataset_id  == "tissue" ~ "Tissue", 
                                dataset_id  == "celline" ~ "Cell line",
                                dataset_id  == "blood" ~ "Blood", 
                                dataset_id  == "singlecell" ~ "Single cell",
                                dataset_id  == "brain" ~ "Brain")) %>% 
  as_tbl_graph() 


enrichment_summary_graph %>% 
  
  create_layout(layout = 'stress') %>% 
  
  ggraph() + 
  geom_edge_link() +
  geom_node_point(size = 3) + 
  geom_node_text(aes(label = name)) +
  theme_void() + 
  scale_edge_color_gradient(low = "white", 
                            high = "orangered") 

node_type_coords <- 
  enrichment_summary_graph %>% 
  as_tibble() %>% 
  mutate(type = case_when(name %in% c("Blood",                
                                      "Brain",          
                                      "Cell line",
                                      "Single cell",
                                      "Tissue") ~ "Section",
                          name %in% c("Protein class",
                                      "Secretome", 
                                      "Subcellular location") ~ "HPA",
                          T ~ "External")) %>% 
  mutate(y = unclass(factor(type, c("External", "Section", "HPA")))) %>% 
  group_by(type) %>% 
  mutate(x = 1:length(name),
         x = x - mean(x), 
         x = ifelse(type == "HPA",
                    x * 2,
                    x))  


graph_data <- 
  enrichment_summary_graph %>% 
  left_join(node_type_coords) %>% 
  activate(edges) %>% 
  mutate(edge_type = case_when(node_type_coords$type[from] == "Section" &
                                 node_type_coords$type[to] == "Section" ~
                                 "Specificity",
                               node_type_coords$type[from] == "External" |
                                 node_type_coords$type[to] == "External" ~
                                 "External",
                               node_type_coords$type[from] == "HPA" |
                                 node_type_coords$type[to] == "HPA" ~
                                 "HPA")) %>% 
  create_layout(layout = 'manual',
                x = node_type_coords$x,
                y = node_type_coords$y) 

edge_data <- get_edges()(graph_data)

plots <- 
  lapply(unique(edge_data$edge_type),
         function(edge_type_) {
           edge_data_ <- 
             edge_data %>% 
             filter(edge_type == edge_type_)
           
           plot <- 
             graph_data %>% 
             ggraph() + 
             # geom_edge_diagonal(aes(width = frac,
             #                    color = frac)) +
             geom_edge_fan(data = edge_data_ %>%
                             filter(edge_type == "Specificity"),
                           aes(color = frac),
                           # arrow = arrow(length = unit(4, 'mm')),
                           strength = 2,
                           alpha = 0.7,
                           show.legend = F, ) +
             geom_edge_diagonal(data = edge_data_ %>% 
                                  filter(edge_type != "Specificity"),
                                aes(color = frac),
                                alpha = 0.7,
                                show.legend = F) +
             geom_node_point(data = . %>% 
                               filter(type == "Section"),
                             aes(color = name),
                             size = 16,
                             show.legend = F) + 
             geom_node_text(data = . %>% 
                              filter(type == "Section"),
                            aes(label = name)) +
             geom_node_text(data = . %>% 
                              filter(type == "HPA"),
                            aes(label = name),
                            vjust = 0) +
             geom_node_text(data = . %>% 
                              filter(type == "External"),
                            aes(label = name),
                            vjust = 1) +
             theme_void() + 
             scale_edge_color_gradient(low = "white", 
                                       high = "black", 
                                       limits = c(0, 1)) +
             scale_color_manual(values = dataset_palette2) +
             scale_x_continuous(expand = expansion(0.1))
           
           ggsave(savepath(paste0("Database integration network ", 
                                  edge_type_, ".pdf")),
                  plot = plot,
                  width = 8, height = 4)
           ggsave(savepath(paste0("Database integration network ", 
                                  edge_type_, ".png")),
                  plot = plot,
                  width = 8, height = 4)
           
           
         })





```

## Expression beeswarm

```{r}
plot_data <- 
  raw_data$singlecell_consensus %>% 
  gather(cell_type, ntpm, -1)

plot_labels <- 
  plot_data %>% 
  group_by(cell_type) %>% 
  top_n(3, ntpm) %>% 
  left_join(gene_names)

plot_data %>% 
  mutate(cell_type = factor(cell_type,
                            consensus_clust_labelorder$cor_average$singlecell)) %>% 
  ggplot(aes(cell_type, 
             log10(ntpm),
             fill = cell_type)) +
  geom_violin(scale = "width",
              draw_quantiles = 0.5, 
              show.legend = F) +
  geom_text(data = plot_labels,
            aes(label = gene_name),
            size = 1,
            hjust = 0) +
  scale_fill_manual(values = consensus_palette) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1,
                                   vjust = 0.5))
ggsave(savepath("Singlecell max exp bee.pdf"),
       width = 10, height = 6)

# 
# plot_data <- 
#   raw_data$singlecell_consensus %>% 
#   gather(cell_type, ntpm, -1) %>% 
#   left_join(gene_consensus_tau %>% 
#               filter(dataset_id == "singlecell") %>% 
#               select(-dataset_id),
#             by = c("ensg_id" = "gene")) %>% 
#   inner_join(gene_classification_long %>% 
#                filter(dataset_id == "singlecell",
#                       !is.na(enhanced_tissues)),
#              by = c("ensg_id", 
#                     "cell_type" = "enhanced_tissues"))

# plot_labels <- 
#   plot_data %>% 
#   group_by(cell_type) %>% 
#   top_n(3, ntpm) %>% 
#   left_join(gene_names)
# 
# plot_data %>% 
#   mutate(cell_type = factor(cell_type,
#                             consensus_clust_labelorder$cor_average$singlecell)) %>% 
#   ggplot(aes(cell_type, 
#              tau_score,
#              fill = cell_type)) +
#   geom_violin(scale = "width",
#               draw_quantiles = 0.5, 
#               show.legend = F) +
#   geom_text(data = plot_labels,
#             aes(label = gene_name),
#             size = 1,
#             hjust = 0) +
#   scale_fill_manual(values = consensus_palette) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1,
#                                    vjust = 0.5))
# ggsave(savepath("Singlecell max exp bee.pdf"),
#        width = 10, height = 6)
plot_data_sum <- 
  plot_data %>% 
  left_join(membr_secr_genes) %>% 
  group_by(cell_type, class) %>% 
  summarise(ntpm = sum(ntpm)) %>% 
  mutate(frac = ntpm / sum(ntpm)) %>% 
  ungroup() %>% 
  mutate(cell_type = factor(cell_type, 
                            consensus_clust_labelorder$cor_average$singlecell))

plot_data_sum %>% 
  write_tsv("data/processed/sc exp sum.tsv")

plot_data_sum %>% 
  ggplot(aes(frac, cell_type, fill = class)) +
  geom_col() +
  scale_fill_manual(name = "Protein class", 
                    values = protein_loc_palette) +
  scale_x_continuous(expand = expansion(0)) +
  labs(x = "Fraction of expression") +
  theme_bw() +
  theme(axis.title.y = element_blank(),
        legend.position = "bottom")

ggsave(savepath("sc fraction secreted bar.pdf"),
       width = 5, height = 10)


  
plot_data_sum %>% 
  select(-ntpm) %>% 
  spread(class, frac) %>% 
  ggplot(aes(membrane, secreted, color = cell_type)) +
  geom_abline(linetype = "dotted",
              color = "gray") +
  geom_point(show.legend = F) +
  geom_text(data = . %>% 
              filter(secreted > 0.2 |
                       membrane > 0.5 |
                       intracellular > 0.85),
            aes(label = cell_type),
            show.legend = F, 
            size = 3,
            vjust = 0,
            hjust = -0.05) +
  
  scale_x_continuous(limits = 0:1, 
                     expand = expansion(0)) + 
  scale_y_continuous(limits = 0:1,
                     expand = expansion(0)) +
  scale_color_manual(values = consensus_palette) +
  coord_fixed() +
  labs(title = "Fraction of expression",
       x = "Membrane proteins",
       y = "Secreted proteins") +
  theme_bw()

ggsave(savepath("sc fraction secreted vs membrane.pdf"),
       width = 5, height = 5)
```


## Correlation datasets 


### All datasets 

```{r}


plot_data <- 
  raw_data[names(raw_data) %in%
             c(#"blood_sample",
               "blood_consensus",
               "tissue_region",
               # "tissue_consensus",
               "celline_consensus",
               # "singlecell_sample",
               "singlecell_consensus")] 


plot_names <- 
  plot_data %>% 
  map(. %>% 
        select(-1) %>% 
        names() %>% 
        enframe("i", "sample")
      ) %>% 
  bind_rows(.id = "dataset_id") %>% 
  select(-i)

plot_combos <- 
  plot_names %>%
  unite(id1, dataset_id, sample, sep = ";") %>% 
  expand_grid(id2 = .$id1) %>% 
  separate(id1, into = c("ds1", "sample1"), sep = ";") %>% 
  separate(id2, into = c("ds2", "sample2"), sep = ";") %>% 
  filter(ds1 != ds2)

plot_cor <- 
  plot_data %>% 
  map(. %>% 
        column_to_rownames("ensg_id")) %>% 
  bind_cols() %>% 
  # head(10000) %>% 
  cor(method = "spearman",
      use = "pairwise.complete.obs")


# all vs all top 2

plot_combo_data <- 
  plot_combos %>% 
  group_by_all() %>% 
  mutate(cor = plot_cor[sample1, sample2]) %>% 
  ungroup() %>%  
  filter(cor > 0.6) %>% 
  group_by(sample1) %>% 
  top_n(2, cor) %>% 
  ungroup() %>% 
  group_by_all() %>% 
  mutate(id = paste(sort(c(sample1, sample2)), collapse = "_")) %>% 
  group_by(id) %>% 
  slice(1) %>% 
  ungroup()

set.seed(212)
plot_cor_graph <- 
  plot_combo_data %>%
  select(sample1, sample2, cor) %>% 
  as_tbl_graph(directed = F) %>% 
  left_join(plot_names %>% 
              select(name = sample, 
                     dataset = dataset_id) %>% 
              distinct(),
            by = "name") %>% 
  mutate(cluster = group_louvain() %>% 
           as.character()) 


set.seed(2)
plot_cor_graph %>% 
  create_layout(layout = 'fr') %>% 
  ggraph() + 
  ggalt::geom_encircle(aes(x, y, 
                           fill = cluster),
                       color = NA, 
                       alpha = 0.2,
                       s_shape=0.5,
                       expand = 0.01,
                       show.legend = F) +
  geom_edge_link(color = "gray", 
                 width = 0.1) +
  geom_node_point(aes(color = name,
                      shape = dataset),
                  show.legend = T) +
  geom_node_text(aes(label = name),
                 size = 1.5,
                 vjust = -0.5) +
  # scale_size_continuous(range = c(0.5, 2)) +
  scale_edge_color_gradient(low = "white", high = "gray",
                            limits = 0:1) +
  scale_color_manual(values = region_palette,
                     guide = "none") +
  scale_fill_discrete(guide = "none") +
  theme_void() + 
  coord_fixed()

ggsave(savepath("dataset correlation network all vs all top2.pdf"),
       width = 6, height = 6)

## Single cell vs all top3


plot_combo_data <- 
  plot_combos %>%
  filter(ds1 == "singlecell_consensus") %>% 
  group_by_all() %>% 
  mutate(cor = plot_cor[sample1, sample2]) %>% 
  ungroup() %>%  
  filter(cor > 0.6) %>% 
  group_by(sample1) %>% 
  top_n(3, cor) %>% 
  ungroup() %>% 
  group_by_all() %>% 
  mutate(id = paste(sort(c(sample1, sample2)), collapse = "_")) %>% 
  group_by(id) %>% 
  slice(1) %>% 
  ungroup()

set.seed(212)
plot_cor_graph <- 
  plot_combo_data %>%
  select(sample1, sample2, cor) %>% 
  as_tbl_graph(directed = F) %>% 
  left_join(plot_names %>% 
              select(name = sample, 
                     dataset = dataset_id) %>% 
              distinct(),
            by = "name") %>% 
  mutate(cluster = group_louvain() %>% 
           as.character()) 


set.seed(2)
plot_cor_graph %>% 
  create_layout(layout = 'fr') %>% 
  ggraph() + 
  ggalt::geom_encircle(aes(x, y, 
                           fill = cluster),
                       color = NA, 
                       alpha = 0.2,
                       s_shape=0.5,
                       expand = 0.01,
                       show.legend = F) +
  geom_edge_link(color = "gray", 
                 width = 0.1) +
  geom_node_point(aes(color = name,
                      shape = dataset),
                  show.legend = T) +
  geom_node_text(aes(label = name),
                 size = 1.5,
                 vjust = -0.5) +
  # scale_size_continuous(range = c(0.5, 2)) +
  scale_edge_color_gradient(low = "white", high = "gray",
                            limits = 0:1) +
  scale_color_manual(values = region_palette,
                     guide = "none") +
  scale_fill_discrete(guide = "none") +
  theme_void() + 
  coord_fixed()

ggsave(savepath("dataset correlation network sc vs all top3.pdf"),
       width = 6, height = 6)


## all vs all - not celline - top2


plot_combo_data <- 
  plot_combos %>%
  filter(ds1 != "celline_consensus") %>% 
  filter(ds2 != "celline_consensus") %>% 
  group_by_all() %>% 
  mutate(cor = plot_cor[sample1, sample2]) %>% 
  ungroup() %>%  
  filter(cor > 0.6) %>% 
  group_by(sample1) %>% 
  top_n(2, cor) %>% 
  ungroup() %>% 
  group_by_all() %>% 
  mutate(id = paste(sort(c(sample1, sample2)), collapse = "_")) %>% 
  group_by(id) %>% 
  slice(1) %>% 
  ungroup()

plot_combo_data %>% 
  filter(sample1 == "amygdala")
  filter(sample2 == "Smooth muscle cells")
plot_cor["amygdala",] %>% 
  enframe() %>% 
  arrange(-value) %>% 
  View

set.seed(212)
plot_cor_graph <- 
  plot_combo_data %>%
  select(sample1, sample2, cor) %>% 
  as_tbl_graph(directed = F) %>% 
  left_join(plot_names %>% 
              select(name = sample, 
                     dataset = dataset_id) %>% 
              distinct(),
            by = "name") %>% 
  mutate(cluster = group_louvain() %>% 
           as.character()) 


set.seed(34)
plot_cor_graph %>% 
  create_layout(layout = 'fr') %>% 
  ggraph() + 
  ggalt::geom_encircle(aes(x, y, 
                           fill = cluster),
                       color = NA, 
                       alpha = 0.2,
                       s_shape=0.5,
                       expand = 0.01,
                       show.legend = F) +
  geom_edge_link(color = "gray", 
                 width = 0.1) +
  geom_node_point(aes(color = name,
                      shape = dataset),
                  show.legend = T) +
  geom_node_text(aes(label = name),
                 size = 1.5,
                 vjust = -0.5) +
  # scale_size_continuous(range = c(0.5, 2)) +
  scale_edge_color_gradient(low = "white", high = "gray",
                            limits = 0:1) +
  scale_color_manual(values = region_palette,
                     guide = "none") +
  scale_fill_discrete(guide = "none") +
  theme_void() + 
  coord_fixed()

ggsave(savepath("dataset correlation network all vs all top2 nocelline.pdf"),
       width = 6, height = 6)




```

### Correlation SC - tissue

```{r}

# singlecell_cluster_data %>% 
#   # filter(cell_type == "Smooth muscle cells") %>% 
#   left_join(raw_data$tissue_region) %>% 
#   group_by(sample) %>% 
#   summarise(cor = cor(ntpm, `amygdala`, method = "spearman")) %>% 
#   arrange(-cor)
# 
singlecell_cluster_data <- 
  raw_data$singlecell_sample %>% 
  gather(sample, ntpm, -1) %>% 
  left_join(select(., sample) %>% 
              distinct() %>% 
              separate(sample, into = c("tissue", "cell_type", "cluster"), 
                       sep = "_", 
                       remove = F))

tissue_region_data <- 
  raw_data$tissue_region %>% 
  gather(tissue_name, ntpm, -1) 


sc_tissue_mapping <- 
  singlecell_cluster_data %>% 
  select(tissue) %>% 
  distinct() %>% 
  mutate(tissue2 = tolower(tissue)) %>% 
  left_join(raw_data$tissue_region %>% 
              names() %>% 
              enframe() %>% 
              select(-name) %>% 
              mutate(tissue_name = value),
            by = c("tissue2" = "value")) %>% 
  mutate(tissue_name = case_when(tissue2 == "adipose" ~ "adipose tissue",
                                 tissue2 == "brain" ~ "cerebral cortex",
                                 tissue2 == "bronchus" ~ "lung",
                                 tissue2 == "endometrium" ~ "endometrium 1",
                                 tissue2 == "eye" ~ "retina",
                                 tissue2 == "skin" ~ "skin 1",
                                 tissue2 == "stomach" ~ "stomach 1",
                                 T ~ tissue_name)) 
  
sc_tissue_joined_data <- 
  singlecell_cluster_data %>% 
  left_join(sc_tissue_mapping) %>% 
  left_join(tissue_region_data,
            by = c("ensg_id", "tissue_name"),
            suffix = c("_sc", "_tissue")) %>% 
  select(ensg_id, sample, tissue, tissue_name, cell_type, cluster, ntpm_sc, ntpm_tissue)
  
plot_cor <- 
  sc_tissue_joined_data %>% 
  group_by(tissue, tissue_name, cell_type) %>% 
  summarise(cor = cor(ntpm_sc, ntpm_tissue, method = "spearman")) %>% 
  ungroup() %>% 
  mutate(cor_label = round(cor, 3))

# sc_tissue_joined_data %>% 
#   # filter(tissue == "Testis") %>% 
#   ggplot(aes(ntpm_sc, ntpm_tissue)) +
#   geom_hex(aes(fill = stat(log10(count))),
#            bins = 30) + 
#   geom_text(data = plot_cor,
#             aes(1, 1, label = cor_label)) +
#   facet_wrap(~tissue_name + cell_type) +
#   scale_x_log10() +
#   scale_y_log10() +
#   scale_fill_viridis() +
#   theme_bw()
# ggsave(savepath("sc vs tissue cor.pdf"),
#        width = 16, height = 24)


plot_cor %>% 
  ggplot(aes(cor, tissue_name, label = cell_type)) +
  geom_text()

rm(singlecell_cluster_data, tissue_region_data)



temp_sc_data <- 
  raw_data$singlecell_consensus %>% 
  gather(cell_type, ntpm, -1)

temp_tissue_data <-
  raw_data$tissue_region %>% 
  gather(tissue, ntpm, -1)

sc_tissue_cor <- 
  expand_grid(cell_type = unique(temp_sc_data$cell_type),
              tissue = unique(temp_tissue_data$tissue)) %>% 
  left_join(temp_sc_data) %>% 
  left_join(temp_tissue_data,
            by = c("tissue", "ensg_id"),
            suffix = c("_sc", "_tissue")) %>% 
  group_by(cell_type, tissue) %>% 
  summarise(cor = cor(ntpm_sc, ntpm_tissue, method = "spearman",
                      use = "pairwise.complete.obs")) %>% 
  ungroup()

sc_tissue_cor %>% 
  spread(tissue, cor) %>% 
  column_to_rownames("cell_type") %>% 
  

  pheatmap(clustering_method = "ward.D2",
           # clustering_distance_rows = 
           
           color = viridis::inferno(100, direction = -1),
           border_color = F,
           annotation_col = tibble(name = names(region_palette)) %>% 
             mutate(tissue = name) %>% 
             column_to_rownames("name"), 
           annotation_row = tibble(name = names(consensus_palette)) %>% 
             mutate("cell type" = name) %>% 
             column_to_rownames("name"), 
           annotation_colors = list(tissue = region_palette,
                                    `cell type`  = consensus_palette),
           cutree_rows = 8, 
           cutree_cols = 8,
           annotation_legend = F,
           filename = savepath("tissue sc cor heatmap.pdf"),
           width = 10, height = 11)

dev.off()
dev.off()

plot_data <-
  sc_tissue_cor %>% 
  filter(cor > 0.6) %>% 
  group_by(cell_type) %>% 
  top_n(3, cor) %>% 
  ungroup() %>% 
  group_by_all() %>% 
  mutate(id = paste(sort(c(cell_type, tissue)), collapse = "_")) %>% 
  group_by(id) %>% 
  slice(1) %>% 
  ungroup()

set.seed(212)
sc_tissue_cor_graph <- 
  plot_data %>%
  as_tbl_graph(directed = F) %>% 
  left_join(plot_data %>% 
              select(1:2) %>% 
              distinct() %>% 
              gather(type, name) %>% 
              distinct(),
            by = "name") %>% 
  mutate(cluster = group_louvain() %>% 
           as.character()) 


set.seed(2)
sc_tissue_cor_graph %>% 
  create_layout(layout = 'fr') %>% 
  ggraph() + 
  ggalt::geom_encircle(aes(x, y, 
                           fill = cluster),
                       color = NA, 
                       alpha = 0.2,
                       s_shape=0.5,
                       expand = 0.01,
                       show.legend = F) +
  geom_edge_link(color = "gray", 
                 width = 0.1) +
  geom_node_point(aes(color = name,
                      shape = type),
                  show.legend = T) +
  geom_node_text(aes(label = name),
                 size = 1.5,
                 vjust = -0.5) +
  # scale_size_continuous(range = c(0.5, 2)) +
  scale_edge_color_gradient(low = "white", high = "gray",
                            limits = 0:1) +
  scale_color_manual(values = region_palette,
                     guide = "none") +
  scale_fill_discrete(guide = "none") +
  theme_void() + 
  coord_fixed()

ggsave(savepath("sc tissue correlation network.pdf"),
       width = 6, height = 6)





```


# UMAP

```{r}

# 
# read_tsv("../../data/HPA/HPA21_E103_files/single cell/sc_cell_data.tsv") %>% 
#   filter(assay_id %in% 1:2) %>% 
#   ggplot(aes(x, y, color = factor(cluster_id))) +
#   geom_point_rast(show.legend = F, 
#                   size = 0.1) +
#   facet_wrap(~assay_id) +
#   coord_fixed() +
#   theme_void()

```


## UMAP polygons


### Create polygons

```{r}
# 
# umap_plot_data <- 
#   umap_data %>%  
#   left_join(clustering_data)
# 
# 
# 
# hull_settings <- 
#   expand_grid(dataset_id = unique(umap_plot_data$dataset_id),
#               relative_bandwidth_i = c(150),
#               poly_concavity = c(2),
#               poly_smoothing = c(1.25),
#               n = 1000) %>% 
#   mutate(relative_bandwidth = 1 / relative_bandwidth_i,
#          i = row_number())
# 
# 
# hull_data <- 
#   pblapply(hull_settings$i,
#          function(i) {
#            
#            plot_data <- 
#              umap_plot_data %>% 
#              filter(dataset_id == hull_settings$dataset_id[i])
#            
#            generate_cluster_hulls(V1 = plot_data$UMAP_1_scaled,
#                                   V2 = plot_data$UMAP_2_scaled, 
#                                   element_id = plot_data$gene,
#                                   cluster_membership = plot_data$cluster, 
#                                   n = hull_settings$n[i],
#                                   poly_concavity = hull_settings$poly_concavity[i],
#                                   poly_smoothing = hull_settings$poly_smoothing[i],
#                                   relative_bandwidth = hull_settings$relative_bandwidth[i]) 
#            
#          })
# 
# 
# hulls <- 
#   hull_data %>% 
#   map(. %>% {.$hulls}) %>% 
#   bind_rows(.id = "i") %>% 
#   left_join(hull_settings %>% 
#               mutate(i = as.character(i)))
# 
# hull_centers <- 
#   hull_data %>% 
#   map(. %>% {.$center_density}) %>% 
#   bind_rows(.id = "i") %>% 
#   left_join(hull_settings %>% 
#               mutate(i = as.character(i)))
#   # hulls %>% 
#   # filter(landmass == 1, sub_cluster == 1) %>% 
#   # group_by(dataset_id, cluster, polygon_id) %>% 
#   # summarise(X = mean(X), 
#   #           Y = mean(Y)) %>% 
#   # ungroup()
# # 
# # hull_centers_all <- 
# #   hulls %>% 
# #   group_by(dataset_id, cluster, polygon_id) %>% 
# #   summarise(X = mean(X), 
# #             Y = mean(Y)) %>% 
# #   ungroup()
# # 
# # hull_segments <-
# #   hull_centers_all %>% 
# #   left_join(hull_centers,
# #             by = c("dataset_id", "cluster"),
# #             suffix = c("", "end")) %>% 
# #   filter(polygon_id != polygon_idend)
# #   
# hulls %>% 
#   # filter( X < -0,
#   #         Y > 0) %>% 
#   ggplot(aes(X, Y, group = polygon_id, fill = cluster)) +
#   # geom_hex(data = umap_data %>%
#   #            filter(dataset_id %in% hull_settings$dataset_id),
#   #          aes(UMAP_1_scaled, UMAP_2_scaled, fill = 1),
#   #          fill = "gray70",
#   #          bins = 200,
#   #          inherit.aes = F) +
#   # geom_segment(data = hull_segments,
#   #              aes(X, Y, 
#   #                  xend = Xend, 
#   #                  yend = Yend,
#   #                  group = cluster),
#   #              inherit.aes = F) +
#   geom_polygon(show.legend = F,
#                color = "black") +
#   geom_point(data = hull_centers,
#              aes(x, y),
#              inherit.aes = F) +
#   # facet_wrap( ~ dataset_id + 
#   #               relative_bandwidth_i + 
#   #               poly_concavity + 
#   #               poly_smoothing, nrow = 2) +
#   facet_wrap( ~ dataset_id) +
#   coord_fixed() +
#   theme_bw() +
#   
#   theme(panel.background = element_rect(fill = "white"),
#         panel.grid = element_blank())
# 
# 
#   
#   select(dataset_id, cluster, sub_cluster, landmass) %>% 
#   filter(cluster == 3, dataset_id == "tissue") %>% 
#   distinct()
#   group_by(dataset_id, )
# 
# 
#   
#   
############
############

```



### plotly

```{r}
umap_poly_centers <- 
  umap_centers %>% 
  left_join(cluster_annotation) 

umap_poly_data_colored <- 
  umap_poly_data %>%
  left_join(umap_poly_data %>% 
              select(dataset_id, cluster) %>% 
              distinct() %>% 
              group_by(dataset_id) %>% 
              mutate(cluster_color = colorRampPalette(ggthemes::tableau_color_pal(palette = "Classic Cyclic")(13))(n_distinct(cluster))))

umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point(data = umap_data %>% 
               filter(dataset_id == "tissue"),
             aes(UMAP_1_scaled, 
                 UMAP_2_scaled),
             color = "black",
             size = 0.1,
             inherit.aes = F) +
  geom_polygon(show.legend = F,
               color = "black",
               size = 0.1) +
  geom_text(data = umap_poly_centers %>% 
              filter(dataset_id == "tissue") ,
            aes(x, y, label = full_annotation),
            size = 3,
            inherit.aes = F) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

umap_poly_data_colored %>% 
  filter(dataset_id == "blood") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point(data = umap_data %>% 
               filter(dataset_id == "blood"),
             aes(UMAP_1_scaled, 
                 UMAP_2_scaled),
             color = "black",
             size = 0.1,
             inherit.aes = F) +
  geom_polygon(show.legend = F,
               color = "black",
               size = 0.1) +
  geom_text(data = umap_poly_centers %>% 
              filter(dataset_id == "blood") ,
            aes(x, y, label = full_annotation),
            size = 3,
            inherit.aes = F) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

plots <-
  unique(umap_poly_data_colored$dataset_id) %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    umap_poly_data_colored %>% 
      filter(dataset_id == dataset_id_) %>% 
      ggplot(aes(X, Y, fill = cluster_color, 
                 group = paste(cluster, sub_cluster))) +
      geom_hex(data = umap_data %>% 
                 filter(dataset_id == dataset_id_),
               aes(UMAP_1_scaled, 
                   UMAP_2_scaled,
                   fill = "1"),
               fill = "darkgray",
               bins = 1000,
               inherit.aes = F) +
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      geom_text(data = umap_poly_centers %>% 
                  filter(dataset_id == dataset_id_) ,
                aes(x, y, label = full_annotation),
                size = 5,
                inherit.aes = F) +
      scale_fill_identity() +
      coord_fixed() +
      theme_void() +
      ggtitle(dataset_id_)
    
    
    
  })

for(dataset_id_ in names(plots)) {
  
  plots[[dataset_id_]] %>% 
    
    ggplotly() %>% 
    hide_legend() %>% 
    as_widget() %>% 
    
    htmlwidgets::saveWidget(savepath(paste0("interactive UMAP ",
                                            dataset_id_,
                                            ".html")))
}

```

### Regular poly UMAPs

```{r}


plots <- 
  umap_poly_centers %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  pull() %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    plot_data <- 
      umap_poly_data_colored %>% 
      filter(dataset_id == dataset_id_) 
    
    
    plot_data %>% 
      ggplot(aes(X, Y, fill = cluster_color, group = polygon_id)) +
      geom_point_rast(data = umap_data %>% 
                        filter(dataset_id == dataset_id_),
                      aes(UMAP_1_scaled, 
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      coord_fixed() +
      theme_void() 
    # ggtitle(str_to_sentence(dataset_id_)) +
    # theme(plot.title = element_text(hjust = 0.5))
    
  })




for(dataset_id_ in names(plots)) {
  
  ggsave(savepath(paste0("Poly UMAP ",
                         dataset_id_,
                         ".pdf")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
  
  ggsave(savepath(paste0("Poly UMAP ",
                         dataset_id_,
                         ".png")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
}

plot <- wrap_plots(plots, nrow = 1)

ggsave(savepath("Poly UMAP combined.pdf"),
       plot = plot,
       width = 20, height = 8)

ggsave(savepath("Poly UMAP combined.png"),
       plot = plot,
       width = 20, height = 8)


#####

plots <- 
  umap_poly_centers %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  pull() %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    plot_data <- 
      umap_poly_data_colored %>% 
      filter(dataset_id == dataset_id_) 
    
    
    plot_data %>% 
      ggplot(aes(X, Y, fill = cluster_color, group = polygon_id)) +
      geom_point_rast(data = umap_data %>% 
                        filter(dataset_id == dataset_id_),
                      aes(UMAP_1_scaled, 
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      geom_polygon(show.legend = F,
                   fill = "gray",
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      coord_fixed() +
      theme_void() 
    
  })




for(dataset_id_ in names(plots)) {
  
  ggsave(savepath(paste0("Poly UMAP gray ",
                         dataset_id_,
                         ".pdf")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
  
  ggsave(savepath(paste0("Poly UMAP gray ",
                         dataset_id_,
                         ".png")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
}

```

### Investigate polygons

```{r}

umap_poly_data %>% 
  filter(cluster == 26, 
         dataset_id == "celline") %>% 
  ggplot(aes(X, Y, group = polygon_id)) +
  geom_polygon(color = "black") +
  facet_wrap(~dataset_id) +
  coord_fixed()

```

### Annotated polygons test

```{r}


plots <- 
  umap_poly_centers %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  pull() %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    umap_poly_data_colored %>% 
      filter(dataset_id == dataset_id_) %>% 
      ggplot(aes(X, Y, fill = cluster_color, group = polygon_id)) +
      geom_point_rast(data = umap_data %>% 
                        filter(dataset_id == dataset_id_),
                      aes(UMAP_1_scaled, 
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      geom_point(show.legend = F,
                 color = NA) +
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      geom_text_repel(data = umap_poly_centers %>% 
                        filter(dataset_id == dataset_id_),
                      aes(x, y, label = full_annotation),
                      inherit.aes = F, 
                      box.padding = 0.5,
                      size = 3) +
      scale_fill_identity() +
      coord_fixed() +
      theme_void() 
    
  })




for(dataset_id_ in names(plots)) {
  
  ggsave(savepath(paste0("Poly UMAP labeled ",
                         dataset_id_,
                         ".pdf")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
  
  ggsave(savepath(paste0("Poly UMAP labeled ",
                         dataset_id_,
                         ".png")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
}




```



### Summarized annotation UMAPs

```{r}

UMAP_plot_hclusts <- 
  umap_poly_centers %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  pull() %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    umap_poly_centers %>% 
      filter(dataset_id == dataset_id_) %>% 
      select(dataset_id, cluster, x, y, full_annotation) %>% 
      unite(id, dataset_id, cluster, full_annotation) %>% 
      column_to_rownames("id") %>% 
      dist() %>% 
      hclust(method = "average")
  })


plots <- 
  umap_poly_centers %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  pull() %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    plot_spec_ann <-
      cutree(UMAP_plot_hclusts[[dataset_id_]], h = 0.4) %>% 
      enframe("id", "UMAP_group") %>% 
      separate(id, into = c("dataset_id", "cluster", "full_annotation"), 
               sep = "_") %>% 
      left_join(cluster_annotation %>% 
                  select(dataset_id, cluster, ann_specificity, ann_function)) %>% 
      left_join(umap_poly_centers) 
    
    plot_spec_ann_centers <- 
      plot_spec_ann %>% 
      group_by(dataset_id, UMAP_group, ann_specificity) %>% 
      summarise(x = median(x), 
                y = median(y))
    
    
    plot_data <- 
      umap_poly_data_colored %>% 
      inner_join(plot_spec_ann %>% 
                   select(1, 2, ann_specificity)) %>% 
      filter(dataset_id == dataset_id_) 
    
    # plot_pal <- 
    #   complete_palette(pal_terms = unique(plot_data$ann_specificity),
    #                    pal = consensus_palette, 
    #                    default_pal = cluster_color_pal)
    
    plot_pal <- 
      complete_palette_neighbor(pal_terms = unique(plot_data$ann_specificity),
                                pal = region_palette,
                                force_colors = c("Non-specific" = "gray50"))
    
    plot_data %>% 
      ggplot(aes(x, y, fill = ann_specificity, group = paste(cluster, sub_cluster))) +
      
      geom_point_rast(data = umap_data %>% 
                        filter(dataset_id == dataset_id_),
                      aes(UMAP_1_scaled, 
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      geom_polygon(aes(X, Y), 
                   show.legend = F,
                   color = "black",
                   size = 0.1, 
                   alpha = 0.8) +
      geom_text(data = plot_spec_ann_centers %>% 
                  filter(dataset_id == dataset_id_) ,
                aes(x, y, label = ann_specificity),
                size = 3,
                inherit.aes = F) +
      scale_fill_manual(values = plot_pal) +
      coord_fixed() +
      theme_void() +
      ggtitle(str_to_sentence(dataset_id_)) +
      theme(plot.title = element_text(hjust = 0.5))
    
  })



for(dataset_id_ in names(plots)) {
  
  ggsave(savepath(paste0("Specificity grouped ann UMAP ",
                         dataset_id_,
                         ".pdf")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
}

plot <- wrap_plots(plots, nrow = 1)

ggsave(savepath("Specificity grouped ann UMAP combined.pdf"),
       plot = plot,
       width = 20, height = 8)



######


plots <- 
  umap_poly_centers %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  pull() %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    plot_spec_ann <-
      cutree(UMAP_plot_hclusts[[dataset_id_]], h = 0.4) %>% 
      enframe("id", "UMAP_group") %>% 
      separate(id, into = c("dataset_id", "cluster", "full_annotation"), 
               sep = "_") %>% 
      left_join(cluster_annotation %>% 
                  select(dataset_id, cluster, ann_specificity, ann_function)) %>% 
      left_join(umap_poly_centers) 
    
    plot_spec_ann_centers <- 
      plot_spec_ann %>% 
      group_by(dataset_id, UMAP_group, ann_specificity) %>% 
      summarise(x = median(x), 
                y = median(y))
    
    
    plot_data <- 
      umap_poly_data_colored %>% 
      inner_join(plot_spec_ann %>% 
                   select(1, 2, ann_specificity)) %>% 
      filter(dataset_id == dataset_id_) 
    
    # plot_pal <- 
    #   complete_palette(pal_terms = unique(plot_data$ann_specificity),
    #                    pal = deframe(tissue_palette), 
    #                    default_pal = cluster_color_pal)
    # 
    plot_pal <- 
      complete_palette_neighbor(pal_terms = unique(plot_data$ann_specificity),
                                pal = region_palette,
                                force_colors = c("Non-specific" = "gray50"))
    
    plot_data %>% 
      ggplot(aes(x, y, fill = ann_specificity, group = paste(cluster, sub_cluster))) +
      
      geom_point_rast(data = umap_data %>% 
                        filter(dataset_id == dataset_id_),
                      aes(UMAP_1_scaled, 
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      geom_polygon(aes(X, Y), 
                   show.legend = F,
                   color = "black",
                   size = 0.1, 
                   alpha = 0.8) +
      # geom_text(data = plot_spec_ann_centers %>% 
      #             filter(dataset_id == dataset_id_) ,
      #           aes(x, y, label = ann_specificity),
      #           size = 3,
      #           inherit.aes = F) +
      scale_fill_manual(values = plot_pal) +
      coord_fixed() +
      theme_void()
    
  })



for(dataset_id_ in names(plots)) {
  
  ggsave(savepath(paste0("Specificity grouped ann UMAP notext ",
                         dataset_id_,
                         ".pdf")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
  
  ggsave(savepath(paste0("Specificity grouped ann UMAP notext ",
                         dataset_id_,
                         ".png")),
         plot = plots[[dataset_id_]],
         width = 8, height = 8)
}
```

## UMAP example

### Tissue Cluster 22
```{r}

example_cluster <- 
  cluster_annotation %>% 
  filter(dataset_id == "tissue") %>% 
  filter(reliability == "High") %>% 
  filter(cluster == 22)

example_cluster_genes <- 
  clustering_data %>% 
  filter(dataset_id == "tissue",
         cluster == example_cluster$cluster)

example_gene <- 
  "ENSG00000111834"
################

p1 <- 
  umap_data %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y)) +
  geom_point_rast(aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  coord_fixed() +
  theme_void() 

p2 <-
  umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>% 
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  geom_polygon(show.legend = F,
               color = NA,
               size = 0.1,
               alpha = 0.8) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

p3 <- 
  umap_poly_data_colored %>% 
  filter(cluster == example_cluster$cluster) %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>% 
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  geom_polygon(show.legend = F,
               color = NA,
               size = 0.1,
               alpha = 0.8) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

p4 <-
  umap_poly_data_colored %>% 
  # filter(cluster == example_cluster$cluster) %>% 
  filter(dataset_id == "tissue") %>% 
  {ggplot(., aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>%
                        filter(dataset_id == "tissue"),
                      aes(UMAP_1_scaled,
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      scale_x_continuous(limits = scales::expand_range(range(.$X[.$cluster == example_cluster$cluster]),
                                                       2)) +
      scale_y_continuous(limits = scales::expand_range(range(.$Y[.$cluster == example_cluster$cluster]),
                                                       2)) +
      coord_fixed() +
      theme_void() }


ggsave(savepath("UMAP example plot p1.pdf"),
       width = 8, height = 8,
       plot = p1)
ggsave(savepath("UMAP example plot p2.pdf"),
       width = 8, height = 8,
       plot = p2)
ggsave(savepath("UMAP example plot p3.pdf"),
       width = 8, height = 8,
       plot = p3)
ggsave(savepath("UMAP example plot p4.pdf"),
       width = 8, height = 8,
       plot = p4)

ggsave(savepath("UMAP example plot p1.png"),
       width = 8, height = 8,
       plot = p1)
ggsave(savepath("UMAP example plot p2.png"),
       width = 8, height = 8,
       plot = p2)
ggsave(savepath("UMAP example plot p3.png"),
       width = 8, height = 8,
       plot = p3)
ggsave(savepath("UMAP example plot p4.png"),
       width = 8, height = 8,
       plot = p4)


###############################################################

p5 <-
  umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>%
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled,
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  
  geom_polygon(data = . %>% 
                 filter(cluster != example_cluster$cluster),
               show.legend = F,
               fill = "gray",
               color = "black",
               size = 0.1,
               alpha = 0.8) +
  geom_polygon(data = . %>% 
                 filter(cluster == example_cluster$cluster),
               show.legend = F,
               color = "black",
               size = 0.1,
               alpha = 0.8) +
  geom_point(data = umap_data %>% 
               filter(dataset_id == "tissue") %>% 
               filter(gene == example_gene),
             aes(UMAP_1_scaled, 
                 UMAP_2_scaled),
             color = "red",
             size = 3,
             inherit.aes = F) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

ggsave(savepath("UMAP example plot p5.pdf"),
       width = 8, height = 8,
       plot = p5)

ggsave(savepath("UMAP example plot p5.png"),
       width = 8, height = 8,
       plot = p5)

p6 <-
  umap_poly_data_colored %>% 
  filter(cluster == example_cluster$cluster) %>%
  filter(dataset_id == "tissue") %>% 
  {ggplot(., aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>%
                        filter(gene %in% example_cluster_genes$gene) %>% 
                        filter(dataset_id == "tissue"),
                      aes(UMAP_1_scaled,
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      scale_x_continuous(limits = scales::expand_range(range(.$X[.$cluster == example_cluster$cluster]),
                                                       1)) +
      scale_y_continuous(limits = scales::expand_range(range(.$Y[.$cluster == example_cluster$cluster]),
                                                       1)) +
      coord_fixed() +
      theme_void() }

ggsave(savepath("UMAP example plot p6.pdf"),
       width = 8, height = 8,
       plot = p6)

ggsave(savepath("UMAP example plot p6.png"),
       width = 8, height = 8,
       plot = p6)

```

### Tissue cluster 54

```{r}

example_cluster <- 
  cluster_annotation %>% 
  filter(dataset_id == "tissue") %>% 
  # filter(reliability == "High") %>% 
  filter(cluster == 54)

example_cluster_genes <- 
  clustering_data %>% 
  filter(dataset_id == "tissue",
         cluster == example_cluster$cluster)

example_gene <- 
  "ENSG00000004864"
################

p1 <- 
  umap_data %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y)) +
  geom_point_rast(aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  coord_fixed() +
  theme_void() 

p2 <-
  umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>% 
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  geom_polygon(show.legend = F,
               color = NA,
               size = 0.1,
               alpha = 0.8) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

p3 <- 
  umap_poly_data_colored %>% 
  filter(cluster == example_cluster$cluster) %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>% 
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled, 
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  geom_polygon(show.legend = F,
               color = NA,
               size = 0.1,
               alpha = 0.8) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

p4 <-
  umap_poly_data_colored %>% 
  # filter(cluster == example_cluster$cluster) %>% 
  filter(dataset_id == "tissue") %>% 
  {ggplot(., aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>%
                        filter(dataset_id == "tissue"),
                      aes(UMAP_1_scaled,
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      scale_x_continuous(limits = scales::expand_range(range(.$X[.$cluster == example_cluster$cluster]),
                                                       2)) +
      scale_y_continuous(limits = scales::expand_range(range(.$Y[.$cluster == example_cluster$cluster]),
                                                       2)) +
      coord_fixed() +
      theme_void() }


ggsave(savepath("UMAP example plot p1.pdf"),
       width = 8, height = 8,
       plot = p1)
ggsave(savepath("UMAP example plot p2.pdf"),
       width = 8, height = 8,
       plot = p2)
ggsave(savepath("UMAP example plot p3.pdf"),
       width = 8, height = 8,
       plot = p3)
ggsave(savepath("UMAP example plot p4.pdf"),
       width = 8, height = 8,
       plot = p4)

ggsave(savepath("UMAP example plot p1.png"),
       width = 8, height = 8,
       plot = p1)
ggsave(savepath("UMAP example plot p2.png"),
       width = 8, height = 8,
       plot = p2)
ggsave(savepath("UMAP example plot p3.png"),
       width = 8, height = 8,
       plot = p3)
ggsave(savepath("UMAP example plot p4.png"),
       width = 8, height = 8,
       plot = p4)


###############################################################

p5 <-
  umap_poly_data_colored %>% 
  filter(dataset_id == "tissue") %>% 
  ggplot(aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
  geom_point_rast(data = umap_data %>%
                    filter(dataset_id == "tissue"),
                  aes(UMAP_1_scaled,
                      UMAP_2_scaled),
                  color = "darkgray",
                  size = 0.1,
                  inherit.aes = F) +
  
  
  geom_polygon(data = . %>% 
                 filter(cluster != example_cluster$cluster),
               show.legend = F,
               fill = "gray",
               color = "black",
               size = 0.1,
               alpha = 0.8) +
  geom_polygon(data = . %>% 
                 filter(cluster == example_cluster$cluster),
               show.legend = F,
               color = "black",
               size = 0.1,
               alpha = 0.8) +
  geom_point(data = umap_data %>% 
               filter(dataset_id == "tissue") %>% 
               filter(gene == example_gene),
             aes(UMAP_1_scaled, 
                 UMAP_2_scaled),
             color = "red",
             size = 3,
             inherit.aes = F) +
  scale_fill_identity() +
  coord_fixed() +
  theme_void() 

ggsave(savepath("UMAP example plot p5.pdf"),
       width = 8, height = 8,
       plot = p5)

ggsave(savepath("UMAP example plot p5.png"),
       width = 8, height = 8,
       plot = p5)

p6 <-
  umap_poly_data_colored %>% 
  filter(cluster == example_cluster$cluster) %>%
  filter(dataset_id == "tissue") %>% 
  {ggplot(., aes(X, Y, fill = cluster_color, group = paste(cluster, sub_cluster))) +
      geom_point_rast(data = umap_data %>%
                        filter(gene %in% example_cluster_genes$gene) %>% 
                        filter(dataset_id == "tissue"),
                      aes(UMAP_1_scaled,
                          UMAP_2_scaled),
                      color = "darkgray",
                      size = 0.1,
                      inherit.aes = F) +
      
      geom_polygon(show.legend = F,
                   color = NA,
                   size = 0.1,
                   alpha = 0.8) +
      scale_fill_identity() +
      scale_x_continuous(limits = scales::expand_range(range(.$X[.$cluster == example_cluster$cluster]),
                                                       1)) +
      scale_y_continuous(limits = scales::expand_range(range(.$Y[.$cluster == example_cluster$cluster]),
                                                       1)) +
      coord_fixed() +
      theme_void() }

ggsave(savepath("UMAP example plot p6.pdf"),
       width = 8, height = 8,
       plot = p6)

ggsave(savepath("UMAP example plot p6.png"),
       width = 8, height = 8,
       plot = p6)

```

# Sample UMAPs

```{r}

sample_UMAPs <- 
  raw_data[names(raw_data) %in%
             c("blood_sample",
               "blood_consensus",
               "tissue_region",
               "tissue_consensus",
               "celline_consensus",
               "singlecell_sample",
               "singlecell_consensus")] %>% 
  map(. %>% 
        column_to_rownames("ensg_id") %>% 
        remove_genes(rm_NA_genes = T,
                     rm_not_expressed = T, 
                     not_expressed_lim = 5,
                     rm_no_variance = T) %>% 
        scale_data(logp1_scale = T, 
                   zscore_scale = T) %>% 
        do_pca() %>% 
        get_pca_scores(use_R2cum_PCselection = T,
                       R2cum_lim = 0.8) %>% 
        do_umap(n_neighbors = floor(sqrt(dim(.)[1]))) %>%
        # do_umap(n_neighbors = dim(.)[2]) %>% 
        as_tibble(rownames = "sample"))

# sample_UMAPs %>% 
#   bind_rows(.id = "dataset_id") %>% 
#   ggplot(aes(UMAP1, UMAP2, color = sample)) +
#   geom_point(show.legend = F) + 
#   facet_wrap(~dataset_id, scales = "free") +
#   scale_color_manual(values = region_palette) +
#   theme_bw()

sample_UMAPs$singlecell_sample %>% 
  separate(sample, into = c("tissue", "cell_type", "cluster"), sep = "_") %>% 
  ggplot(aes(UMAP1, UMAP2, color = cell_type)) +
  # geom_path(show.legend = F) + 
  geom_point(show.legend = F) +
  geom_text_repel(data = . %>% 
                    group_by(cell_type) %>% 
                    summarise(UMAP1 = mean(UMAP1),
                              UMAP2 = mean(UMAP2)),
                  aes(label = cell_type),
                  size = 2,
                  show.legend = F) +
  
  scale_color_manual(values = region_palette) +
  theme_bw() +
  coord_fixed()
ggsave(savepath("singlecell sample UMAP.pdf"),
       width = 8, height = 10)

##############################################

plots <- 
  lapply(names(sample_UMAPs), 
         function(ds_name) {
           
           if(ds_name == "singlecell_sample") {
             plot_data <- 
               sample_UMAPs[[ds_name]] %>% 
               separate(sample, into = c("tissue", "sample", "cluster"), sep = "_")
               
           } else if(ds_name == "blood_sample") {
             plot_data <- 
               sample_UMAPs[[ds_name]] %>% 
               separate(sample, into = c("sample", "id1", "id2"), sep = "_")
               
           } else {
             plot_data <- 
               sample_UMAPs[[ds_name]]
           }
           
           plot_data%>% 
             ggplot(aes(UMAP1, UMAP2, color = sample)) +
             # geom_path(show.legend = F) + 
             geom_point(show.legend = F, 
                        size = 1) +
             geom_text_repel(data = . %>% 
                               group_by(sample) %>% 
                               summarise(UMAP1 = mean(UMAP1),
                                         UMAP2 = mean(UMAP2)),
                             aes(label = sample),
                             size = 2,
                             show.legend = F) +
             
             scale_color_manual(values = region_palette) +
             theme_bw() +
             coord_fixed() +
             theme(axis.text = element_blank(),
                   axis.ticks = element_blank(),
                   panel.grid = element_blank()) +
             ggtitle(ds_name)
           
         })

pdf(savepath("Sample UMAPs.pdf"),
    width = 6, height = 6)
plots
dev.off()


plots <- 
  lapply(names(sample_UMAPs), 
         function(ds_name) {
           
           if(ds_name == "singlecell_sample") {
             plot_data <- 
               sample_UMAPs[[ds_name]] %>% 
               separate(sample, into = c("tissue", "sample", "cluster"), sep = "_")
               
           } else if(ds_name == "blood_sample") {
             plot_data <- 
               sample_UMAPs[[ds_name]] %>% 
               separate(sample, into = c("sample", "id1", "id2"), sep = "_")
               
           } else {
             plot_data <- 
               sample_UMAPs[[ds_name]]
           }
           
           plot_data%>% 
             ggplot(aes(UMAP1, UMAP2, color = sample)) +
             # geom_path(show.legend = F) + 
             geom_point(show.legend = F, 
                        size = 1) +
             geom_text_repel(data = . %>% 
                               group_by(sample) %>% 
                               summarise(UMAP1 = mean(UMAP1),
                                         UMAP2 = mean(UMAP2)),
                             aes(label = sample),
                             size = 2,
                             # box.padding = unit(2, "mm"),
                             show.legend = F) +
             
             scale_color_manual(values = region_palette) +
             theme_bw() +
             coord_fixed() +
             theme(axis.text = element_blank(),
                   axis.ticks = element_blank(),
                   axis.title = element_blank(),
                   panel.grid = element_blank(),
                   plot.background = element_blank(),
                   panel.border = element_blank())
           
         })

pdf(savepath("Sample UMAPs simplified.pdf"),
    width = 6, height = 6)
plots
dev.off()

```

### Combined
```{r}

plot_data <- 
  raw_data[names(raw_data) %in%
             c(#"blood_sample",
               "blood_consensus",
               "tissue_region",
               # "tissue_consensus",
               "celline_consensus",
               # "singlecell_sample",
               "singlecell_consensus")] 



comb_UMAP_processing <- 
  function(plot_data) {
    plot_data_comb <- 
      plot_data %>% 
      map(. %>% 
            column_to_rownames("ensg_id") %>% 
            remove_genes(rm_NA_genes = T,
                         rm_not_expressed = T, 
                         not_expressed_lim = 1,
                         rm_no_variance = T)) 
    
    plot_names <- 
      plot_data %>% 
      map(. %>% 
            select(-1) %>% 
            names() %>% 
            enframe("i", "sample")) %>% 
      bind_rows(.id = "dataset_id")
    
    plot_common_genes <- 
      plot_data_comb %>% 
      map(. %>% 
            rownames()) %>% 
      unlist() %>% 
      table() %>% 
      {names(.)[. == length(plot_data)]}
    
    plot_data_comb <- 
      plot_data_comb %>% 
      map(. %>% 
            {.[plot_common_genes,]}) %>% 
      bind_cols() 
    
    wide_data <- 
      plot_data_comb
    
    all(plot_names$sample == names(wide_data))
    
    batch <- 
      plot_names$dataset_id
    
    combined_PCA <-
      plot_data_comb %>% 
      scale_data(logp1_scale = T) %>%
      do_limma(batch = batch) %>% 
      scale_data(zscore_scale = T) %>%
      # %>% 
      extract_variable_features(log1p_data = F) %>%
      
      do_pca() %>% 
      get_pca_scores(use_R2cum_PCselection = T,
                     R2cum_lim = 0.8) 
  }

  
combined_PCA <- 
  comb_UMAP_processing(plot_data)

combined_UMAP <-
  combined_PCA %>% 
  do_umap(n_neighbors = 30) %>% 
  as_tibble(rownames = "sample")

combined_UMAP %>% 
  left_join(plot_names) %>% 
  ggplot(aes(UMAP1, UMAP2, color = sample)) +
  geom_point(data = combined_UMAP,
             color = "gray",
             size = 0.5) +
  geom_point(show.legend = F) + 
  facet_wrap(~dataset_id) +
  scale_color_manual(values = region_palette) +
  theme_bw() +
  coord_fixed()
ggsave(savepath("Combined sample UMAP.pdf"),
       width = 8, height = 10)

combined_PCA %>% 
  as_tibble(rownames = "sample") %>% 
  left_join(plot_names) %>% 
  ggplot(aes(PC3, PC4, color = sample)) +
  geom_point(data = combined_PCA %>% 
               as_tibble(rownames = "sample"),
             color = "gray",
             size = 0.5) +
  geom_point(show.legend = F) + 
  facet_wrap(~dataset_id) +
  scale_color_manual(values = region_palette) +
  theme_bw() +
  coord_fixed()
ggsave(savepath("Combined sample PCA.pdf"),
       width = 8, height = 10)


plot <- 
  combined_UMAP %>% 
  left_join(plot_names) %>% 
  ggplot(aes(UMAP1, UMAP2, color = sample,
             label = sample,
             shape = dataset_id)) +
  geom_point(show.legend = T) + 
  geom_text(size = 3, 
            show.legend = F,
            vjust = 0) +
  scale_color_manual(values = region_palette,
                     guide = "none") +
  theme_bw() +
  coord_fixed()



plot %>% 
  ggplotly() %>% 
  as_widget() %>% 
  htmlwidgets::saveWidget(savepath("interactive combined sample UMAP.html"))

plot <- 
  combined_UMAP %>% 
  left_join(plot_names) %>% 
  ggplot(aes(UMAP1, UMAP2, color = sample,
             label = sample, shape = dataset_id)) +
  geom_point(show.legend = T) + 
  scale_color_manual(values = region_palette,
                     guide = "none") +
  theme_bw() +
  coord_fixed()

ggsave(savepath("interactive combined sample UMAP.pdf"),
       plot = plot,
       width = 8, height = 8)




```

## Pairwise

```{r}


plot_data <- 
  raw_data[names(raw_data) %in%
             c("blood_consensus",
               "tissue_region",
               "celline_consensus")] 


combined_PCA_pw <-
  plot_data %>% 
  map(. %>%
        list(raw_data$singlecell_consensus) %>% 
        set_names(c("data", "singlecell")) %>% 
        comb_UMAP_processing())

combined_UMAP_pw <-
  combined_PCA_pw %>%
  map(. %>% 
        do_umap(n_neighbors = 30) %>% 
        as_tibble(rownames = "sample"))

plot_names <- 
  plot_data %>% 
  map(. %>% 
        names()) %>% 
  unlist()

combined_UMAP_pw %>% 
  bind_rows(.id = "dataset_id") %>% 
  mutate(type = ifelse(sample %in% plot_names,
                       "sample",
                       "singlecell")) %>% 
  ggplot(aes(UMAP1, UMAP2, color = sample,
             label = sample, shape = type)) +
  geom_point(show.legend = T) + 
  geom_text_repel(data = . %>% 
                    filter(runif(nrow(.)) > 0.75),
                  size = 1.5) +
  scale_color_manual(values = region_palette,
                     guide = "none") +
  facet_wrap(~dataset_id) +
  coord_fixed() +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())


ggsave(savepath("Pairwise combined sample UMAPs.pdf"),
       width = 10, height = 5)


```


# Bubble heatmap

```{r}

plot_data <-
  enrichment_data %>%
  filter(!test_type %in% c("KEGG",
                           "GO_BP_original",
                           "GO_CC_original",
                           "GO_MF_original",
                           "GO_BP_simplified",
                           "GO_CC_simplified",
                           "GO_MF_simplified",
                           "reactome")) %>% 
  filter(grepl("specificity", test_type)) %>% 
  left_join(cluster_annotation %>% 
              select(dataset_id, cluster, full_annotation), sep = " ") %>% 
  unite(cluster, dataset_id, cluster, full_annotation) %>% 
  select(cluster, ID, odds_ratio) %>% 
  group_by_all() %>% 
  mutate(capped_odds_ratio = min(c(odds_ratio,
                                   50))) %>% 
  ungroup()

cluster_levels <- 
  clustering_data %>% 
  select(dataset_id, cluster) %>% 
  distinct() %>% 
  arrange(dataset_id, as.numeric(cluster)) %>% 
  left_join(cluster_annotation %>% 
              select(dataset_id, cluster, full_annotation), sep = " ") %>% 
  unite(cluster, dataset_id, cluster, full_annotation)



plot_size_range <- c(1, 4)

plot_legend_settings <- 
  tibble(odds_ratio_str = c(1, 10, 20, 30, 40, ">50"),
         y = 1:6,
         odds_ratio = c(1, 10, 20, 30, 40, 50)) 

plot_legend <-
  plot_legend_settings %>% 
  ggplot(aes(1, y, 
             size = odds_ratio,
             fill = odds_ratio,
             label = odds_ratio_str)) +
  geom_point(shape = 21,
             show.legend = F) +
  geom_text(size = 4, 
            hjust = 0,
            nudge_x = 0.1) +
  annotate("text", x = 1, y = 7, label = "Odds ratio",
           fontface = "bold", hjust = 0) +
  scale_x_continuous(limits = c(0, 2)) +
  scale_y_continuous(limits = c(0, 7)) +
  scale_fill_gradient(name = "Odds ratio",
                      low = "white", high = "orangered",
                      breaks = plot_legend_settings$odds_ratio,
                      limits = range(plot_legend_settings$odds_ratio)) + 
  scale_size_continuous(name = "Odds ratio", 
                        range = plot_size_range, 
                        breaks = plot_legend_settings$odds_ratio,
                        limits = range(plot_legend_settings$odds_ratio)) +
  theme_void()

plot_clustering <- 
  plot_data %>% 
  select(cluster, ID, odds_ratio) %>% 
  spread(ID, odds_ratio, fill = 0) %>% 
  column_to_rownames("cluster") %>%
  list(tissue = t(.), 
       cluster = .) %>%
  map(. %>% 
        dist(method = "binary") %>% 
        hclust(method = "ward.D2"))

plot_order <- 
  list(cluster = unique(c(with(plot_clustering$cluster,
                               labels[order]),
                          cluster_levels$cluster)),
       ID = with(plot_clustering$tissue,
                 labels[order]))

plot <-
  plot_data %>%
  ungroup() %>% 
  mutate(cluster = factor(cluster,
                          plot_order$cluster),
         ID = factor(ID,
                     plot_order$ID)) %>%
  ggplot(aes(ID, cluster, 
             size = capped_odds_ratio, 
             fill = capped_odds_ratio)) +
  geom_point(data = expand_grid(cluster = plot_order$cluster,
                                ID = plot_order$ID) %>%
               mutate(cluster = factor(cluster, unique(cluster)),
                      ID = factor(ID, unique(ID))),
             fill = NA,
             size = NA) +
  geom_point(shape = 21,
             show.legend = F) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0.5),
        legend.position = "right") +
  # coord_fixed() +
  scale_fill_gradient(name = "Odds ratio",
                      low = "white", high = "orangered",
                      breaks = plot_legend_settings$odds_ratio,
                      limits = range(plot_legend_settings$odds_ratio)) + 
  scale_size_continuous(name = "Odds ratio", 
                        range = plot_size_range, 
                        breaks = plot_legend_settings$odds_ratio,
                        limits = range(plot_legend_settings$odds_ratio)) +
  theme(axis.title = element_blank()) + 
  coord_fixed()

ggsave(savepath("Total Specificity Bubble heatmap.pdf"),
       plot = plot,
       width = 35, height = 35)


```


## Bubble network

```{r}
jaccard <- function(a, b) {
  intersection = length(intersect(a, b))
  union = length(a) + length(b) - intersection
  return (intersection/union)
}


# bubble_graph <- 
bubble_data <- 
  enrichment_data %>%
  filter(dataset_id != "brain") %>% 
  filter(test_type != "specificity_brain") %>% 
  
  # filter(test_type == "specificity_tissue") %>% 
  filter(grepl("specificity", test_type)) %>%
  filter(p.adjust < 0.01) %>% 
  left_join(cluster_annotation %>% 
              select(dataset_id, cluster, full_annotation), sep = " ") %>% 
  unite(cluster, dataset_id, cluster, full_annotation) %>% 
  select(cluster, ID, odds_ratio)  

bubble_dist <- 
  bubble_data %>% 
  spread(ID, odds_ratio, fill = 0) %>% 
  column_to_rownames("cluster") %>% 
  dist(method = "binary") %>%
  as.matrix() %>% 
  as_tibble(rownames = "cluster1") %>% 
  gather(cluster2, dist, -1) %>% 
  filter(cluster1 != cluster2)

bubble_dist %>% 
  ggplot(aes(dist)) +
  geom_density()

bubble_nn_data <- 
  bubble_dist %>%
  filter(dist < 0.75) %>% 
  group_by(cluster1) %>%
  top_n(5, -dist) %>%
  arrange(cluster1)


bubble_graph <- 
  bubble_nn_data %>% 
  as_tbl_graph() %>% 
  left_join(as_tibble(.) %>% 
              separate(name, into = c("dataset_id", "cluster", "full_annotation"),
                       sep = "_",
                       remove = F))
bubble_graph %>% 
  ggraph('fabric', sort.by = node_rank_fabric()) + 
  geom_node_range(colour = 'grey') + 
  geom_edge_span(end_shape = 'square') + 
  coord_fixed() +
  theme_void()


# umap_layout <- 
#   bubble_graph %>% 
#   activate(edges) %>% 
#   as_tibble() %>% 
#   spread(to, dist, fill = 1) %>% 
#   column_to_rownames("from") %>% 
#   umap(n_neighbors = 30) %>% 
#   as_tibble()

bubble_graph_layout <- 
  bubble_graph %>% 
  # create_layout(layout = 'manual', 
  #               x = umap_layout$V1,
  #               y = umap_layout$V2)
  create_layout(layout = 'fr')



bubble_graph_layout %>% 
  ggraph() + 
  geom_edge_link(color = "gray") +
  geom_node_point(aes(color = dataset_id),
                  size = 3) + 
  # geom_node_text(aes(label = full_annotation),
  #                size = 2) +
  
  theme_void() +
  scale_color_manual(values = dataset_palette) +
  coord_fixed()

ggsave(savepath("interactive specificity net.pdf"),
       width = 8, height = 8)
ggsave(savepath("interactive specificity net.png"),
       width = 8, height = 8)

plot <-
  bubble_graph_layout %>%
  as_tibble() %>% 
  ggplot(aes(x, y, label = full_annotation, color = dataset_id)) +
  geom_segment(data = get_edges()(bubble_graph_layout) %>% 
                 as_tibble(),
               aes(x = x,
                   y = y, 
                   xend = xend, 
                   yend = yend),
               size = 0.2,
               inherit.aes = F) +
  geom_point(size = 5) +
  geom_text(size = 3,
            color = "black") +
  theme_void() +
  scale_color_manual(values = dataset_palette) +
  coord_fixed()

plot %>% 
  ggplotly() %>% 
  as_widget() %>% 
  htmlwidgets::saveWidget(savepath("interactive specificity net.html"))

```


# Classification

## specificity

```{r}

gene_classification

```


## Tau vs specificity
```{r}
p1 <-
  gene_classification %>% 
  left_join(gene_consensus_tau,
            by = c("dataset_id", "ensg_id" = "gene")) %>% 
  mutate(specificity_category = factor(specificity_category, 
                                       spec_category_levels)) %>% 
  arrange(desc(dataset_id)) %>% 
  mutate(dataset_id = factor(dataset_id, unique(dataset_id))) %>% 
  ggplot(aes(tau_score, specificity_category, fill = specificity_category)) +
  geom_violin(scale = "width", 
              show.legend = F) +
  facet_wrap(~dataset_id, ncol = 1, strip.position = "left") +
  scale_fill_manual(values = spec_category_pal) +
  scale_x_continuous(limits = c(0, 1),
                     breaks = c(0, 0.5, 1),
                     expand = expansion(0)) +
  # coord_fixed(1/5) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.line.x = element_line(),
        panel.border = element_blank()) +
  labs(x = "Tau Score")

p2 <- 
  gene_classification %>% 
  group_by(dataset_id, specificity_category) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(specificity_category = factor(specificity_category, 
                                       spec_category_levels)) %>% 
  arrange(desc(dataset_id)) %>% 
  mutate(dataset_id = factor(dataset_id, unique(dataset_id))) %>% 
  
  ggplot(aes(n, specificity_category, fill = specificity_category,
             label = n)) +
  geom_col() +
  geom_text(hjust = 0,
            vjust = 0.5) +
  facet_wrap(~dataset_id, ncol = 1, strip.position = "left") +
  scale_fill_manual(values = spec_category_pal) +
  scale_x_continuous(limits = c(0, 20000),
                     breaks = c(0, 10000, 20000),
                     expand = expansion(0)) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.line.x = element_line(),
        panel.border = element_blank(), 
        strip.background = element_blank(),
        strip.text = element_blank()) +
  labs(x = "Number of genes")

p1 + p2
ggsave(savepath("Specificity bar tau.pdf"),
       width = 6, height = 10)


p1 <-
  gene_classification %>%
  filter(dataset_id == "singlecell") %>% 
  left_join(gene_consensus_tau,
            by = c("dataset_id", "ensg_id" = "gene")) %>% 
  mutate(specificity_category = factor(specificity_category, 
                                       spec_category_levels)) %>% 
  arrange(desc(dataset_id)) %>% 
  mutate(dataset_id = factor(dataset_id, unique(dataset_id))) %>% 
  ggplot(aes(tau_score, specificity_category, fill = specificity_category)) +
  geom_violin(scale = "width", 
              show.legend = F) +
  facet_wrap(~dataset_id, ncol = 1, strip.position = "left") +
  scale_fill_manual(values = spec_category_pal) +
  scale_x_continuous(limits = c(0, 1),
                     breaks = c(0, 0.5, 1),
                     expand = expansion(0)) +
  # coord_fixed(1/5) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.line.x = element_line(),
        panel.border = element_blank()) +
  labs(x = "Tau Score")

p2 <- 
  gene_classification %>% 
  filter(dataset_id == "singlecell") %>% 
  group_by(dataset_id, specificity_category) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(specificity_category = factor(specificity_category, 
                                       spec_category_levels)) %>% 
  arrange(desc(dataset_id)) %>% 
  mutate(dataset_id = factor(dataset_id, unique(dataset_id))) %>% 
  
  ggplot(aes(n, specificity_category, fill = specificity_category,
             label = n)) +
  geom_col() +
  geom_text(hjust = 0,
            vjust = 0.5) +
  facet_wrap(~dataset_id, ncol = 1, strip.position = "left") +
  scale_fill_manual(values = spec_category_pal) +
  scale_x_continuous(limits = c(0, 20000),
                     breaks = c(0, 10000, 20000),
                     expand = expansion(0)) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.line.x = element_line(),
        panel.border = element_blank(), 
        strip.background = element_blank(),
        strip.text = element_blank()) +
  labs(x = "Number of genes")

p1 + p2
ggsave(savepath("Specificity bar tau sc only.pdf"),
       width = 6, height = 3)
ggsave(savepath("Specificity bar tau sc only.png"),
       width = 6, height = 3)


```

## Tau vs Expression

```{r}
consensus_maxexp <- 
  consensus_data %>% 
  map(. %>% column_to_rownames("ensg_id") %>% 
        apply(MARGIN = 1, 
              max) %>% 
        enframe("gene", "max_exp")) %>% 
  bind_rows(.id = "dataset_id")


gene_consensus_tau %>% 
  left_join(consensus_maxexp) %>% 
  left_join(gene_classification,
            by = c("gene" = "ensg_id", "dataset_id")) %>% 
  ggplot(aes(tau_score, log10(max_exp))) +
  geom_hex(aes(alpha = stat(count)),
           fill = "orangered",
           bins = 50) +
  scale_alpha_continuous(range = c(0, 1)) +
  scale_fill_viridis() +
  facet_grid(specificity_category ~ dataset_id) +
  # facet_wrap(~dataset_id) +
  theme_bw()

```


## Classification per tissue

```{r}

consensus_data_cols <- 
  consensus_data %>% 
  map(. %>% 
        select(-1) %>% 
        names() %>% 
        enframe()) %>% 
  bind_rows(.id = "dataset_id") 

###############


plot_data <- 
  gene_classification_long %>% 
  left_join(gene_consensus_tau,
            by = c("dataset_id", "ensg_id" = "gene")) %>% 
  filter(!is.na(enhanced_tissues)) %>% 
  mutate(dataset_id = factor(dataset_id, plot_datasets))

plot_data_sum <- 
  plot_data %>% 
  group_by(dataset_id, 
           specificity_category,
           enhanced_tissues) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(specificity_category = factor(specificity_category,
                                       rev(spec_category_levels)))

left_plots <- 
  plot_datasets %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    plot_data_clust <- 
      consensus_clust_cor_average[[dataset_id_]] %>% 
      ggdendro::dendro_data() %>% 
      {left_join(.$segments, 
                 .$labels, 
                 by = c("x" = "x", "yend" = "y")) }
    
    
    plot_data_clust %>%
      ggplot() +
      geom_segment(aes(x=y, y=x, xend=yend, yend=xend, group = label))+
      geom_rect(data = . %>% 
                  filter(!is.na(label)),
                aes(xmin=0, ymin=x + 0.5, 
                    xmax=-0.02, ymax=xend - 0.5, 
                    fill = label),
                color = "black",
                show.legend = F) +
      # scale_color_manual(values = tissue_pal)+
      scale_fill_manual(values = consensus_palette)+
      scale_x_reverse(expand = expansion(0), position = "top")+
      scale_y_continuous(expand = expansion(0)) +
      xlab("1 - Spearman's rho") +
      
      theme(axis.text.y = element_blank(), 
            axis.title.y = element_blank(), 
            axis.ticks.y = element_blank(),
            plot.margin = unit(c(1,1,1,1), units = "mm"), 
            panel.background = element_blank()) 
  })

right_plots <- 
  plot_datasets %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    
    plot_data_dataset <- 
      plot_data_sum %>% 
      filter(dataset_id == dataset_id_) %>% 
      mutate(enhanced_tissues = factor(enhanced_tissues, 
                                       consensus_clust_labelorder$cor_average[[dataset_id_]]))
    
    right_plot <-
      plot_data_dataset %>%
      ggplot(aes(n, enhanced_tissues, fill = specificity_category)) +
      geom_col(width = 0.8, size = 0.1, 
               show.legend = F) +
      theme_bw() +
      scale_fill_manual(values = gene_category_pal, name = "Specificity") +
      # coord_flip() +
      xlab("Number of genes") +
      
      scale_x_continuous(position = "top", expand = expansion(c(0, 0.1))) + 
      scale_y_discrete(expand = expansion()) +
      
      theme(axis.text.y = element_text(hjust = 0.5), 
            legend.position = c(0.7, 0.5),
            axis.title.y = element_blank(),
            panel.border = element_blank())
    
    
  })

plots <- 
  plot_datasets %>% 
  set_names(., .) %>% 
  lapply(function(dataset_id_) {
    
    left_plots[[dataset_id_]] + 
      right_plots[dataset_id_]
    
  })

for(dataset_ in names(plots)) {
  ggsave(savepath(paste0("Specificity barplot", 
                         dataset_,
                         ".pdf")),
         plot = plots[[dataset_]],
         width = 7, height= 10)
  
}

# Combined plot

plot_datasets <-
  names(consensus_clust_cor_average) %>% 
  sort()



plot_order <- 
  consensus_clust_labelorder$cor_average[plot_datasets] %>% 
  unlist()

plot_data_all <- 
  plot_data_sum %>% 
  ungroup() %>% 
  mutate(enhanced_tissues = factor(enhanced_tissues,
                                   plot_order))

plot_data_clust <- 
  consensus_clust_cor_average %>% 
  map(. %>% 
        ggdendro::dendro_data() %>% 
        {left_join(.$segments, 
                   .$labels, 
                   by = c("x" = "x", 
                          "yend" = "y"))}) %>% 
  bind_rows(.id = "dataset_id") %>% 
  as_tibble() %>% 
  mutate(dataset_id = factor(dataset_id, rev(plot_datasets)))

left_plot <-
  plot_data_clust %>%
  ggplot() +
  geom_segment(aes(x=y, y=x, xend=yend, yend=xend, group = label))+
  geom_rect(data = . %>% 
              filter(!is.na(label)),
            aes(xmin=0, ymin=x + 0.5, 
                xmax=-0.02, ymax=xend - 0.5, 
                fill = label),
            color = "black",
            show.legend = F) +
  # scale_color_manual(values = tissue_pal)+
  scale_fill_manual(values = consensus_palette)+
  scale_x_reverse(expand = expansion(0), position = "top",
                  limits = c(0.4, -0.02),
                  breaks = c(0, 0.2, 0.4))+
  scale_y_continuous(expand = expansion(0)) +
  xlab("1 - Spearman's rho") +
  facet_grid(dataset_id ~ ., 
             space = "free",
             scales = "free_y", switch = "y") +
  # theme_bw() +
  theme(axis.text.y = element_blank(), 
        axis.title.y = element_blank(), 
        axis.ticks.y = element_blank(),
        plot.margin = unit(c(1,1,1,1), units = "mm"), 
        panel.background = element_blank()) 

right_plot <-
  plot_data_all %>%
  ggplot(aes(n, enhanced_tissues, fill = specificity_category)) +
  geom_col(width = 0.8, size = 0.1, 
           show.legend = F) +
  theme_bw() +
  scale_fill_manual(values = gene_category_pal, name = "Specificity") +
  # coord_flip() +
  xlab("Number of genes") +
  
  scale_x_continuous(position = "top", expand = expansion(c(0, 0.1))) + 
  scale_y_discrete(expand = expansion(0)) +
  facet_grid(dataset_id ~ ., 
             space = "free",
             scales = "free_y") +
  theme(axis.text.y = element_text(), 
        legend.position = c(0.7, 0.5),
        axis.title.y = element_blank(),
        panel.border = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank())



left_plot | right_plot

ggsave(savepath("Combined classification barplot.pdf"),
       width = 7, height = 24)

plot_meta <- 
  plot_data_all %>% 
  group_by(dataset_id) %>% 
  summarise(n = n_distinct(enhanced_tissues)) %>% 
  deframe()

left_plots2 <- 
  names(left_plots) %>% 
  set_names(., .) %>% 
  lapply(function(x) left_plots[[x]] + 
           scale_y_continuous(expand = expansion(0, 
                                                 c(max(plot_meta) -
                                                     plot_meta[[x]], 
                                                   0))))
right_plots2 <- 
  names(right_plots) %>% 
  set_names(., .) %>% 
  lapply(function(x) right_plots[[x]] + 
           scale_y_discrete(expand = expansion(0, 
                                               c(max(plot_meta) -
                                                   plot_meta[[x]], 
                                                 0))))
plots2 <- 
  names(right_plots) %>% 
  lapply(function(dataset_id_) {
    
    left_plots2[[dataset_id_]] + 
      right_plots2[dataset_id_]
  })

wrap_plots(plots2, nrow = 1,)

ggsave(savepath("Combined classification barplot row.pdf"),
       width = 20, height = 10)



######  


plot_data_sum %>%
  group_by(enhanced_tissues) %>% 
  mutate(total_n = sum(n)) %>%
  
  ungroup() %>% 
  arrange(total_n) %>% 
  mutate(enhanced_tissues = factor(enhanced_tissues, unique(enhanced_tissues))) %>% 
  ggplot(aes(enhanced_tissues, n, fill = specificity_category)) +
  geom_col(width = 0.8, size = 0.1, 
           show.legend = F) +
  # simple_theme +
  scale_fill_manual(values = gene_category_pal, name = "Specificity") +
  # coord_flip() +
  ylab("Number of genes") +
  
  facet_grid(~ dataset_id, 
             space = "free", 
             scales = "free_x") +
  scale_y_continuous(expand = expansion(c(0, 0.1))) + 
  scale_x_discrete(expand = expansion()) +
  theme_bw() +
  theme(legend.position = c(0.7, 0.5),
        panel.border = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = -90, vjust = 0.5, hjust = 0))
ggsave(savepath("Classification barplot sorted.pdf"),
       width = 24, height = 4)
ggsave(savepath("Classification barplot sorted.png"),
       width = 24, height = 4)


####################

plot_data_tau <- 
  plot_data %>% 
  filter(specificity_category %in% c("Tissue enriched",
                                     "Group enriched")) %>%
  mutate(enhanced_tissues = factor(enhanced_tissues, 
                                   plot_order),
         dataset_id = factor(dataset_id,
                             rev(plot_datasets))) %>% 
  group_by(dataset_id, enhanced_tissues) %>% 
  summarise(quantile = c(0.25, 0.5, 0.75),
            quantil_value = quantile(tau_score, quantile)) %>% 
  ungroup()

plot_data_tau %>% 
  ggplot(aes(enhanced_tissues, tau_score, fill = enhanced_tissues)) +
  geom_crossbar(data = plot_data_tau %>% 
                  spread(quantile, quantil_value),
                aes(ymin = `0.25`,
                    y = `0.5`,
                    ymax = `0.75`),
                show.legend = F) +
  # geom_boxplot(show.legend = F, 
  #              outlier.colour = NA, 
  #              coef = 0) +
  
  ylab("Tau IQR") +
  facet_grid(~ dataset_id, 
             space = "free",
             scales = "free_x") +
  scale_y_continuous(limits = c(0.7, 1),
                     expand = expansion(0)) + 
  scale_x_discrete(expand = expansion(0), position = "top") +
  scale_fill_manual(values = consensus_palette) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, hjust = 1), 
        strip.placement = "outside",
        legend.position = c(0.7, 0.5),
        axis.title.x = element_blank(),
        panel.border = element_blank())

ggsave(savepath("Tau crossbar.pdf"),
       width = 24, height = 6)


plot_data_tau %>% 
  spread(quantile, quantil_value) %>% 
  arrange(-`0.5`) %>% 
  mutate(enhanced_tissues = factor(enhanced_tissues, enhanced_tissues)) %>% 
  ggplot(aes(enhanced_tissues, 
             ymin = `0.25`,
             y = `0.5`,
             ymax = `0.75`, 
             fill = enhanced_tissues)) +
  geom_crossbar(data = ,
                aes(),
                show.legend = F) +
  ylab("Tau IQR") +
  facet_grid(~ dataset_id, 
             space = "free",
             scales = "free_x") +
  scale_y_continuous(limits = c(0.7, 1),
                     expand = expansion(0)) + 
  scale_x_discrete(expand = expansion(0), position = "top") +
  scale_fill_manual(values = consensus_palette) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, hjust = 1), 
        strip.placement = "outside",
        legend.position = c(0.7, 0.5),
        axis.title.x = element_blank(),
        panel.border = element_blank())

ggsave(savepath("Tau crossbar ordered.pdf"),
       width = 24, height = 6)
```


## Compare tau between datasets

```{r}

gene_classification %>%
  
  select(ensg_id, dataset_id1 = dataset_id, 
         specificity_category1 = specificity_category) %>% 
  filter(specificity_category1 != "Not detected") %>% 
  distinct() %>% 
  expand_grid(dataset_id2 = unique(.$dataset_id1)) %>% 
  left_join(select(., ensg_id, dataset_id2 = dataset_id1,
                   specificity_category2 = specificity_category1) %>% 
              distinct()) %>% 
  filter(specificity_category2 != "Not detected") %>% 
  filter(dataset_id1 != dataset_id2) %>% 
  left_join(gene_consensus_tau,
            by = c("dataset_id1" = "dataset_id", "ensg_id" = "gene")) %>% 
  left_join(gene_consensus_tau,
            by = c("dataset_id2" = "dataset_id", "ensg_id" = "gene"),
            suffix = c("1", "2")) %>% 
  
  mutate(dataset_id1 = factor(dataset_id1, rev(plot_datasets)),
         dataset_id2 = factor(dataset_id2, rev(plot_datasets))) %>% 
  
  ggplot(aes(tau_score1, tau_score2)) +
  # geom_point_rast() +
  geom_hex(aes(fill = stat(log10(count))),
           bins = 25) +
  geom_text(data = . %>% 
              group_by(dataset_id1, dataset_id2) %>% 
              summarise(cor = cor(tau_score1, tau_score2, method = "spearman")),
            aes(0, 0, label = round(cor, 2), 
                size = cor),
            hjust = 0,
            vjust = 0) +
  scale_fill_viridis() +
  scale_size_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 1),
                     expand = expansion(0),
                     breaks = c(0, 0.5, 1)) +
  scale_y_continuous(limits = c(0, 1),
                     expand = expansion(0),
                     breaks = c(0, 0.5, 1)) +
  facet_grid(dataset_id2 ~ dataset_id1) +
  coord_fixed() +
  theme_bw() + 
  theme(panel.spacing = unit(1, "mm")) +
  ggtitle("Spearman correlation of tau for expressed genes")
ggsave(savepath("Tau correlation.pdf"),
       width = 8, height = 6)

```

## Compare classification between datasets

```{r}

plot_data_raw <- 
  gene_classification %>%
  
  select(ensg_id, dataset_id1 = dataset_id, 
         specificity_category1 = specificity_category) %>% 
  distinct() %>% 
  expand_grid(dataset_id2 = unique(.$dataset_id1)) %>% 
  left_join(select(., ensg_id, dataset_id2 = dataset_id1,
                   specificity_category2 = specificity_category1) %>% 
              distinct()) %>% 
  filter(dataset_id1 != dataset_id2)

plot_data <- 
  plot_data_raw %>% 
  # left_join(gene_consensus_tau,
  #           by = c("dataset_id1" = "dataset_id", "ensg_id" = "gene")) %>% 
  # left_join(gene_consensus_tau,
  #           by = c("dataset_id2" = "dataset_id", "ensg_id" = "gene"),
  #           suffix = c("1", "2")) %>% 
  
  
  group_by(dataset_id1, specificity_category1,
           dataset_id2, specificity_category2) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(dataset_id1 = factor(dataset_id1, rev(plot_datasets)),
         dataset_id2 = factor(dataset_id2, rev(plot_datasets)),
         specificity_category1 = factor(specificity_category1,
                                        spec_category_levels),
         specificity_category2 = factor(specificity_category2,
                                        spec_category_levels)) 


###
data %>% 
  select(var1 = var1, 
         var2 = var2, 
         cat1 = cat1, 
         cat2 = cat2) %>% 
  ungroup() %>%
  mutate(row_n = row_number()) %>%
  gather(cat_type, cat, -var1, -var2, -row_n) %>%
  mutate(cat = factor(cat, levels = cat_levels),
         cat_type = case_when(cat_type == "cat1" ~ cat_names[1],
                              cat_type == "cat2" ~ cat_names[2]),
         cat_type = factor(cat_type, 
                           levels = cat_names))
###

# plot_data_raw %>% 
#   # mutate(specificity_category1 = fct_collapse(specificity_category1, 
#   #                                             elevated = c("Tissue enriched",
#   #                                                          "Group enriched"),
#   #                                             not_elevated = c("Low tissue specificity",
#   #                                                              "Tissue enhanced",
#   #                                                              "Not detected")),
#   #        specificity_category2 = fct_collapse(specificity_category2, 
#   #                                             elevated = c("Tissue enriched",
#   #                                                          "Group enriched"),
#   #                                             not_elevated = c("Low tissue specificity",
#   #                                                              "Tissue enhanced",
#   #                                                              "Not detected"))) %>% 
#   group_by(ds1 = dataset_id1,
#            ds2 = dataset_id2) %>% 
#   
#   summarise(ari = ARI(specificity_category1, specificity_category2))

plot_data  %>%
  mutate(i = row_number(),
         ds1 = dataset_id1,
         ds2 = dataset_id2) %>% 
  pivot_longer(cols = contains("dataset") | contains("spec"),
               names_to = c(".value", "set"), 
               names_pattern = "(.*)(.$)") %>% 
  ggplot(aes(x = set, stratum = specificity_category, 
             alluvium = n,
             y = n,
             fill = specificity_category)) +
  scale_x_discrete(expand = c(.1, .1), position = "top") +
  geom_flow(show.legend = F) +
  geom_stratum(show.legend = F, color = NA) + 
  scale_fill_manual(values = gene_category_pal) + 
  facet_grid(ds2 ~ ds1, 
             scales = "free_x") +
  
  theme(axis.text.x = element_text(size = 18, face = "bold"),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(), 
        panel.background = element_blank(), 
        axis.title = element_blank())

ggsave(savepath("Spec allivial.pdf"),
       width = 6, height = 7)

```

###ORA 

```{r}


ORA_data <- 
  gene_classification_long %>% 
  filter(specificity_category %in% c("Tissue enhanced", 
                                     "Group enhanced"))

ORA_settings <- 
  expand_grid(dataset_id = plot_datasets,
              db_dataset_id = plot_datasets) %>% 
  filter(dataset_id != db_dataset_id)

universe <- 
  gene_classification_long$ensg_id %>% 
  unique

spec_ORA_res <- 
  ORA_settings %>% 
  group_by_all() %>% 
  do({
    
    settings <<- .
    
    gene_associations <- 
      ORA_data %>% 
      filter(dataset_id == settings$dataset_id) %>% 
      select(partition = enhanced_tissues, 
             gene = ensg_id)
    database <- 
      ORA_data %>% 
      filter(dataset_id == settings$db_dataset_id) %>% 
      select(term_id = enhanced_tissues, 
             gene = ensg_id)
    
    
    
    perform_ORA(gene_associations, 
                database,
                universe,
                n_clusters = 5)
  }) %>% 
  ungroup()

plot_data <- 
  spec_ORA_res %>% 
  mutate(GeneFrac = as.numeric(gsub("\\/.*", "", GeneRatio)) /
           as.numeric(gsub(".*\\/", "", GeneRatio)),
         BgFrac = as.numeric(gsub("\\/.*", "", BgRatio)) /
           as.numeric(gsub(".*\\/", "", BgRatio)),
         odds_ratio = GeneFrac / BgFrac)   %>% 
  filter(odds_ratio > 2)

plot_order <- 
  plot_data %>% 
  select(partition, ID, odds_ratio) %>% 
  spread(ID, odds_ratio, fill = 0) %>% 
  column_to_rownames("partition") %>% 
  cluster_wide_data(distance_method = "binary",
                    clustering_method = "ward.D2", 
                    cluster_rows = T,
                    cluster_cols = T) %>% 
  {list(partition = rownames(.),
        ID = colnames(.))}

plot_data_ordered <- 
  plot_data %>% 
  mutate(partition = factor(partition, plot_order$partition),
         ID = factor(ID, plot_order$ID),
         dataset_id = factor(dataset_id, rev(plot_datasets)),
         db_dataset_id = factor(db_dataset_id, rev(plot_datasets))) %>% 
  arrange(odds_ratio) %>% 
  filter(dataset_id == "singlecell")


# p1 <- 
plot_data_ordered %>% 
  ggplot(aes(ID, partition, size = odds_ratio,
             color = odds_ratio)) +
  geom_point() +
  geom_tile(data = . %>% 
              select(partition) %>% 
              distinct(),
            aes(0, partition, fill = partition),
            show.legend = F,
            inherit.aes = F,
            color = "black")  +
  geom_tile(data = . %>% 
              select(ID, db_dataset_id) %>% 
              distinct(),
            aes(ID, 0, fill = ID),
            show.legend = F,
            inherit.aes = F,
            color = "black")  +
  facet_grid(~db_dataset_id,
             space = "free",
             scales = "free_x") +
  scale_color_gradient(high = "orangered", 
                      low = "lightgray") +
  scale_fill_manual(values = consensus_palette) +
  scale_size_continuous(range = c(0.5,4)) +
  labs(y = "Single cell type",
       x = "Tissue/Cell specificity",
       title = "Statistical significant overlap in specificity classification",
       subtitle = "Single cell type VS") +
  # coord_fixed() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, 
                                   vjust = 0.5))

ggsave(savepath("Spec-spec ORA bubble singlecell.pdf"),
       width = 16, height = 12)


plot_data_ordered_all <- 
  plot_data %>% 
  mutate(partition = factor(partition, plot_order$partition),
         ID = factor(ID, plot_order$ID),
         dataset_id = factor(dataset_id, rev(plot_datasets)),
         db_dataset_id = factor(db_dataset_id, rev(plot_datasets))) %>% 
  arrange(odds_ratio) 


# p1 <- 
plot_data_ordered_all %>% 
  # select(dataset_id, db_dataset_id, partition) %>%
  # distinct() %>%
  # filter(dataset_id == "tissue") %>%
  # complete(partition, nesting(db_dataset_id, dataset_id)) %>% View
  # filter(complete.cases(.))
  # group_by(dataset_id)
  ggplot(aes(ID, partition, size = odds_ratio,
             color = odds_ratio)) +
  geom_point() +
  geom_tile(data = . %>% 
              select(dataset_id, partition) %>% 
              distinct(),
            aes(0, partition, fill = partition),
            show.legend = F,
            inherit.aes = F,
            color = "black")  +
  geom_tile(data = . %>% 
              select(db_dataset_id, ID) %>% 
              distinct(),
            aes(ID, 0, fill = ID),
            show.legend = F,
            inherit.aes = F,
            color = "black")  +
  facet_grid(dataset_id ~ db_dataset_id,
             space = "free",
             scales = "free") +
  scale_color_gradient(high = "orangered", 
                      low = "lightgray") +
  scale_fill_manual(values = consensus_palette) +
  scale_size_continuous(range = c(0.5,4)) +
  labs(y = "Tissue/Cell type",
       x = "Tissue/Cell specificity",
       title = "Statistical significant overlap in specificity classification",
       subtitle = "All VS All") +
  # coord_fixed() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, 
                                   vjust = 0.5))

ggsave(savepath("Spec-spec ORA bubble all.pdf"),
       width = 19, height = 19)
# #####
# 
# p1 <- 
#   plot_data_ordered %>% 
#   ggplot(aes(ID, partition, size = odds_ratio,
#              fill = odds_ratio)) +
#   geom_point(shape = 21) +
#   facet_grid(~db_dataset_id,
#              space = "free",
#              scales = "free_x") +
#   scale_fill_gradient(high = "orangered", 
#                       low = "white") +
#   # coord_fixed() +
#   theme_bw() +
#   theme(axis.text = element_blank(),
#         axis.ticks = element_blank(),
#         axis.title = element_blank())
# 
# 
# p1_ybar <-
#   plot_data_ordered %>% 
#   select(partition) %>% 
#   ggplot(aes("1", partition, fill = partition)) +
#   geom_tile(show.legend = F,
#             color = "black") +
#   scale_fill_manual(values = consensus_palette) +
#   scale_x_discrete(expand = expansion(0)) +
#   coord_fixed() +
#   theme(axis.text.x = element_blank(),
#         axis.title.x = element_blank(), 
#         axis.ticks.x = element_blank())
# 
# p1_xbar <-
#   plot_data_ordered %>% 
#   select(ID) %>% 
#   ggplot(aes(ID, "1", fill = ID)) +
#   geom_tile(show.legend = F,
#             color = "black") +
#   scale_fill_manual(values = consensus_palette) +
#   scale_x_discrete(expand = expansion(0)) +
#   coord_fixed() +
#   theme(axis.text.y = element_blank(),
#         axis.title.y = element_blank(), 
#         axis.ticks.y = element_blank(),
#         axis.text.x = element_text(angle = 90, hjust = 1, 
#                                    vjust = 0.5))
# 
# p1_blank <- 
#   ggplot() + 
#   geom_blank() +
#   theme_void()
# 
# (p1_ybar | p1) / (p1_blank | p1_xbar)

```

