---
title: "Cluster annotation"
author: "Max J Karlsson & María Bueno Alvez"
date: "23/07/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Run Checklist

## 1. Make sure the dataset metadata is the correct file

## 2. Check that gene dictionaries and databases are the right version

##---
##---

# Setup

## Load libraries
```{r setup, include=FALSE}

library(tidyverse)
library(tidytext)
library(multidplyr)
library(vroom)
library(magrittr)
library(broom)
library(ggpmisc)
library(rrvgo)
library(treemapify)
library(pbapply)
library(clusterProfiler)
library(patchwork)
library(ggrastr)
source("scripts/functions_utility.R")
source("scripts/generate_gene_clusters.R")

select <- dplyr::select
slice <- dplyr::slice
rename <- dplyr::rename
simplify <- clusterProfiler::simplify




```

## File structure
```{r}

dataset_metadata <- read_csv("run_settings/20210901_settings.csv")

file_structure <-
  create_folder_structure(dataset_id = dataset_metadata$dataset_id,
                          dataset_run_id = dataset_metadata$version,
                          main_folder = "results/Clustering results",
                          folders = c("data" = "data", 
                                      "svg" = "svg",
                                      "treemap" = "svg/treemap",
                                      "heatmap" = "svg/heatmap",
                                      "enrichment" = "enrichment",
                                      "clustering" = "clustering"))
```


## Load clustering data

```{r}

clustering_data <- 
  file_structure %>% 
  map(function(x) x$clustering) %>% 
  map(function(x) read_csv(paste(x, "final_consensus.csv", sep = "/"))) %>% 
  bind_rows(.id = "dataset_id") %>% 
  mutate(cluster = as.character(cluster))


```

## Load gene mapping

```{r}
entrez_dict <- 
  read_tsv("data/meta/ensembl103_entrez.txt") %>% 
  select(ensg_id = 1, 
         entrez = 2) %>% 
  filter(!is.na(entrez))

gene_info <- 
  read_tsv("data/meta/geninfo_92.tsv")

```


## Load databases

```{r}

enrichment_settings <- 
  read_csv("run_settings/20210907_enrichment_settings.csv")


enrichment_settings_versions <- 
  enrichment_settings %>% 
  mutate(version = case_when(is.na(version) & 
                               grepl("^GO", id) ~ paste("org.Hs.eg.db",
                                                        packageVersion("org.Hs.eg.db")),
                             is.na(version) & 
                               id == "KEGG" ~ gsub("-", "", Sys.Date()),
                             !is.na(version) ~ version))

database_formatting_functions <- 
  list(secretome_location = . %>% 
         select(ensg_id = Ensembl,
                term = `Secretome location`) %>% 
         filter(!is.na(term)) %>% 
         mutate(term_id = term),
       
       specificity_blood = . %>% 
         select(ensg_id = Ensembl,
                term = `RNA blood cell specific NX`) %>% 
         separate_rows(term, sep = ";") %>% 
         mutate(term = gsub(":.*", "", term)) %>% 
         filter(!is.na(term)) %>% 
         filter(ensg_id %in% clustering_data$gene) %>% 
         mutate(term_id = term),
       
       specificity_brain = . %>% 
         select(ensg_id = Ensembl,
                term = `RNA brain regional specific NX`) %>% 
         separate_rows(term, sep = ";") %>% 
         mutate(term = gsub(":.*", "", term)) %>% 
         filter(!is.na(term)) %>% 
         filter(ensg_id %in% clustering_data$gene) %>% 
         mutate(term_id = term),
       
       specificity_singlecell = . %>% 
         select(ensg_id = Ensembl,
                term = `RNA single cell type specific NX`) %>% 
         separate_rows(term, sep = ";") %>% 
         mutate(term = gsub(":.*", "", term)) %>% 
         filter(!is.na(term)) %>% 
         filter(ensg_id %in% clustering_data$gene) %>% 
         mutate(term_id = term),
       
       specificity_tissue = . %>% 
         select(ensg_id = Ensembl,
                term = `RNA tissue specific NX`) %>% 
         separate_rows(term, sep = ";") %>% 
         mutate(term = gsub(":.*", "", term)) %>% 
         filter(!is.na(term)) %>% 
         filter(ensg_id %in% clustering_data$gene) %>% 
         mutate(term_id = term),
       
       subcellular_location = . %>% 
         filter(Gene %in% clustering_data$gene) %>% 
         select(ensg_id = Gene, 
                Enhanced, 
                Supported, 
                Approved) %>% 
         gather(reliability, location, -1) %>% 
         filter(!is.na(location)) %>% 
         separate_rows(location, sep = ";") %>% 
         select(ensg_id, term = location) %>% 
         mutate(term_id = term),
       
       panglao_cellmarkers = . %>% 
         filter(grepl("Hs", species), 
                `gene type` %in% c("protein-coding gene",
                                   "protein coding gene")) %>% 
         select(gene_name = 2, 
                term = 3) %>% 
         distinct() %>% 
         mutate(term_id = term),
       
       protein_class = . %>% select(ensg_id = Ensembl,
                                    protein_class = `Protein class`) %>% 
         separate_rows(protein_class, sep = ", ") %>% 
         rename(term = protein_class) %>% 
         mutate(term_id = term),
       
       reactome = . %>% 
         filter(species == "Homo sapiens") %>% 
         select(ensg_id, term = description, term_id = id) %>% 
         distinct() %>% 
         filter(ensg_id %in% clustering_data$gene), 
       
       trrust = . %>% 
         select(gene_name, term) %>% 
         mutate(term_id = term))



gene_formatting_functions <- 
  list(`gene names` = . %>% 
         left_join(gene_info %>% 
                     select(ensg_id, gene_name),
                   by = c("gene_name")) %>% 
         select(ensg_id, term, term_id))
  




database_terms <- 
  enrichment_settings %>% 
  filter(!is.na(file)) %>% 
  group_by(id) %>% 
  do({
    g_data <<- .
    
    raw_db <- 
      read_tsv(paste0("annotation_databases/", g_data$file))
    
    db_format_function <- 
      database_formatting_functions[[g_data$id]]
    
    out_db <- 
      raw_db %>% 
      db_format_function()
    
    if(g_data$gene_format != "ensembl") {
      gene_format_function <- gene_formatting_functions[[g_data$gene_format]]
      
      out_db <- 
        out_db %>%
        gene_format_function()
        
    }
    
    stopifnot(names(out_db) == c("ensg_id", "term", "term_id")) 
    
    out_db
    
  }) %>% 
  ungroup()



```


#Enrichment analyses


##GO

```{r}



worker_cluster <- new_cluster(n = 5)
cluster_library(worker_cluster, c("dplyr",
                                  "tidyverse"))
cluster_copy(worker_cluster, c("compareCluster", 
                               "simplify", 
                               "min"))

cluster_membership <-
  clustering_data %>% 
  expand_grid(ontology = c("BP", "CC", "MF")) 

GO_enrichment_res <- 
  cluster_membership %>% 
  group_by(ontology) %>% 
  do({
    ontology_data <<- .
    
    ontology <- unique(ontology_data$ontology)
    cluster_copy(worker_cluster, c("ontology"))
    
    ontology_data %>% 
      group_by(dataset_id) %>% 
      partition(worker_cluster) %>% 
      do({
        dataset_data <- .
        dataset_res <- 
          compareCluster(gene ~ cluster,
                         fun = "enrichGO",
                         data = dataset_data,
                         keyType = "ENSEMBL",
                         OrgDb = "org.Hs.eg.db",
                         minGSSize = 10,
                         maxGSSize = 500,
                         universe = unique(dataset_data$gene),
                         ont = ontology) 
        
        bind_rows(simplified = dataset_res %>% 
                    simplify(cutoff = 0.7, by = "p.adjust", select_fun = min) %>% 
                    as_tibble(),
                  original = dataset_res %>% 
                    as_tibble(),
                  .id = "simplified")
      }) %>% 
      ungroup() %>% 
      collect()
      
    
    
  }) %>% 
  ungroup()


rm(worker_cluster)




```


##KEGG

```{r}

worker_cluster <- new_cluster(n = 5)
cluster_library(worker_cluster, c("dplyr",
                                  "tidyverse"))
cluster_copy(worker_cluster, c("compareCluster"))

clustering_data_entrez <- 
  clustering_data %>%
  left_join(entrez_dict,
            by = c("gene" = "ensg_id")) 

KEGG_enrichment_res <- 
  clustering_data_entrez %>% 
  
  group_by(dataset_id) %>% 
  partition(worker_cluster) %>% 
  do({
    dataset_data <- .
    compareCluster(entrez ~ cluster,
                   fun = "enrichKEGG",
                   data = dataset_data,
                   universe = unique(dataset_data$entrez)) %>% 
      as_tibble()
  }) %>% 
  ungroup() %>% 
  collect()
      
    


rm(worker_cluster)


```

## Other databases

```{r}

worker_cluster <- new_cluster(n = 5)
cluster_library(worker_cluster, c("dplyr",
                           "tidyverse"))
cluster_copy(worker_cluster, c("enricher"))

database_enrichment_res <- 
  database_terms %>% 
  group_by(id) %>% 
  do({
    # For each database
    database <<- .
    
    term2gene <- 
      database %>% 
      select(term, ensg_id)
    
    cluster_copy(worker_cluster, c("term2gene"))
    
    # For 
    clustering_data %>% 
      group_by(dataset_id) %>% 
      partition(worker_cluster) %>% 
      do({
        g_clustering_data <<- .
        
        g_clustering_data %>% 
          group_by(cluster) %>% 
          
          do({
            g_data <- .
            
            pull(g_data, gene) %>% 
              enricher(maxGSSize = Inf, 
                       universe = unique(g_clustering_data$gene),
                       TERM2GENE = term2gene) %>% 
              as_tibble()
          }) %>% 
          ungroup()
        
      }) %>% 
      ungroup() %>% 
      collect()
    
  }) %>% 
  ungroup()
  
rm(worker_cluster)

```


##Save results

```{r}


all_res <-
  list(GO_enrichment_res %>% 
         mutate(test_type = paste0("GO_", ontology, "_", simplified)) %>% 
         select(-ontology, -Cluster, -simplified),
       KEGG_enrichment_res %>% 
         mutate(test_type = "KEGG") %>% 
         select(-Cluster),
       database_enrichment_res %>% 
         rename(test_type = id)) %>% 
    bind_rows() %>% 
  mutate(GeneFrac = as.numeric(gsub("\\/.*", "", GeneRatio)) /
           as.numeric(gsub(".*\\/", "", GeneRatio)),
         BgFrac = as.numeric(gsub("\\/.*", "", BgRatio)) /
           as.numeric(gsub(".*\\/", "", BgRatio)),
         odds_ratio = GeneFrac / BgFrac)   %>% 
  left_join(enrichment_settings_versions %>% 
              select(test_type = id,
                     test_description = description, 
                     database_version = version)) %>% 

  select(dataset_id, test_type, test_description, database_version, 
         cluster, ID, Description,
         GeneRatio, GeneFrac, BgRatio, BgFrac, odds_ratio, everything()) 




all_res %>% 
  group_by(dataset_id) %>% 
  do({
    g_data <<- .
    
    enrichment_savefile <- 
      paste(file_structure[[unique(g_data$dataset_id)]]$enrichment,
            "enrichment_results.tsv",
            sep = "/")
    
    write_tsv(g_data,
              enrichment_savefile)
    
  })






```

# Generate plots

## Treemaps 

```{r}

ontology_pal <- 
  c("MF" = "#FF6F00",
    "CC" = "#C71B00",
    "BP" = "#018EA0")

# Generate similarity matrix for all significant GO terms:
GO_sim <-
  pblapply(c("BP" = "GO_BP_original",
             "CC" = "GO_CC_original", 
             "MF" = "GO_MF_original"),
           function(ont_) {
             GO_terms <-
               all_res %>% 
               filter(test_type == ont_) %>% 
               pull(ID) %>% 
               unique()
             
             calculateSimMatrix(GO_terms, 
                                orgdb = "org.Hs.eg.db",
                                ont = str_extract(ont_, "BP|CC|MF"),
                                method = "Wang")
           })


# Generate plot legend: 


plot_legend <-
  enframe(ontology_pal,
          "Ontology", 
          "Color") %>% 
  ggplot(aes(1, 1, fill = Ontology)) +
  geom_tile() +
  scale_fill_manual(values = ontology_pal)
plot_legend <- 
  ggpubr::get_legend(plot_legend) %>% 
  ggpubr::as_ggplot()

file_structure %>% 
  map(function(x) x$treemap) %>% 
  lapply(function(treemap_path) {
    ggsave(paste(treemap_path, "treemap legend.svg",
                 sep = "/"),
           plot = plot_legend,
       width = 1, height = 1)
  })
  
# Generate plot data for treemaps


worker_cluster <- new_cluster(n = 6)
cluster_library(worker_cluster, c("dplyr",
                                  "tidyverse"))
cluster_copy(worker_cluster, c("GO_sim", 
                               "reduceSimMatrix"))

plot_data <- 
  all_res %>%
  filter(test_type %in% c("GO_BP_original",
                          "GO_CC_original", 
                          "GO_MF_original")) %>% 
  mutate(GeneFrac = as.numeric(gsub("\\/.*", "", GeneRatio)) /
           as.numeric(gsub(".*\\/", "", GeneRatio)),
         ontology = str_extract(test_type, "BP|CC|MF")) %>% 
  group_by(ontology, test_type, dataset_id, cluster) %>% 
  filter(n_distinct(ID) >= 2) %>% 
  group_by(ontology, test_type, dataset_id, cluster) %>% 
  partition(worker_cluster) %>% 
  do({
    g_data <<- .
    
    sim_mat <- 
      GO_sim[[unique(g_data$ontology)]][g_data$ID,
                                        g_data$ID]
    
    sim_mat %>%
      reduceSimMatrix(setNames(-log10(g_data$p.adjust), g_data$ID),
                      threshold=0.7,
                      orgdb="org.Hs.eg.db") %>%
      as_tibble() %>%
      rename(term_cluster = cluster)
    
    
  }) %>% 
  
  ungroup() %>% 
  collect()

rm(worker_cluster)

plot_data %>% 
  group_by(dataset_id, cluster) %>% 
  do({
    g_data <<- .
    
    treemap_savefile <- 
      paste0(file_structure[[unique(g_data$dataset_id)]]$treemap,
             "/",
             "treemap ", 
             unique(g_data$cluster),
             ".svg")
    
    
    
  })


plot_settings <-
  plot_data %>% 
  select(dataset_id, cluster) %>% 
  arrange(dataset_id, as.numeric(cluster)) %>% 
  distinct() %>% 
  mutate(i = row_number(), 
         treemap_folder = file_structure[dataset_id] %>% 
           map(function(x) x$treemap) %>% 
           unlist(),
         treemap_savefile = paste0(treemap_folder,
                                   "/",
                                   "treemap ", 
                                   unique(cluster),
                                   ".svg"))

plots <- 
  pblapply(plot_settings$i,
           function(i) {
             plot_data %>% 
               filter(dataset_id == plot_settings$dataset_id[i],
                      cluster == plot_settings$cluster[i]) %>% 
               left_join(enframe(ontology_pal,
                                 "ontology", 
                                 "color"),
                         by = "ontology") %>% 
               group_by(parent) %>% 
               mutate(nn = n_distinct(go),
                      color2 = spread_colors(unique(color), n_distinct(go), colorrange = 0.2)) %>% 
               ggplot(aes(area = score, subgroup = parentTerm)) +
               
               geom_treemap(aes(fill = color2), 
                            color = "black",
                            show.legend = T) +
               geom_treemap_subgroup_border(color = "black") +
               
               geom_treemap_text(aes(label = term),
                                 colour = "black", 
                                 place = "centre",
                                 alpha = 0.4,
                                 grow = TRUE) +
               
               geom_treemap_subgroup_text(place = "centre", 
                                          grow = T, 
                                          reflow = T,
                                          alpha = 1, 
                                          colour = "white", 
                                          fontface = "bold",
                                          min.size = 0) +
               
               scale_fill_identity() +
               theme_void() 
               # ggtitle(paste(plot_settings$dataset_id[i],
               #               "cluster",
               #               plot_settings$cluster[i]))
           })


for(i in plot_settings$i) {
  setting <- 
    plot_settings[i, ]
  ggsave(setting$treemap_savefile,
         plots[[i]],
         width = 6, height = 4)
}



```

## Heatmaps

```{r}
cluster_membconf_data <- 
  clustering_data %>% 
  # filter(dataset_id == "tissue") %>% 
  select(dataset_id, ensg_id = gene, cluster) %>% 
  mutate(cluster_membership_confidence = rnorm(nrow(.), 0.95, 0.05),
         cluster_membership_confidence = ifelse(cluster_membership_confidence > 1, 
                                                1, 
                                                round(cluster_membership_confidence, 2)))

heatmap_palette <- 
  colorRampPalette(c("white", "orangered"))(100)

heatmap_data <-
  file_structure %>% 
  map(function(x) x$data) %>% 
  map(function(x) vroom(paste(x, 
                              "data.tsv",
                              sep = "/"))) %>% 
  bind_rows(.id = "dataset_id")

heatmap_clustering_data <- 
  heatmap_data %>% 
  left_join(clustering_data %>% 
              select(dataset_id, gene, cluster),
            by = c("dataset_id",
                   "ensg_id" = "gene")) %>% 
  group_by(dataset_id, ensg_id) %>% 
  mutate(tmm_scaled = tmm / max(tmm),
         tmm_zscore = (tmm - mean(tmm, na.rm = T)) / sd(tmm, na.rm = T)) %>% 
  ungroup()


# worker_cluster <- new_cluster(n = 5)
# cluster_library(worker_cluster, c("dplyr",
#                            "tidyverse"))
# cluster_copy(worker_cluster, c("file_structure",
#                                "cluster_membconf_data"))


heatmap_clustering_data %>% 
  # filter(cluster %in% 1:10) %>%
  group_by(dataset_id, cluster) %>% 
  # partition(worker_cluster) %>%
  do({
    g_cluster_data <<- .
    g_dataset_id <- unique(g_cluster_data$dataset_id)
    
    
    plot_savename_zscore <- 
      paste0(file_structure[[g_dataset_id]]$heatmap, "/",
             "heatmap zscore ",
             unique(g_cluster_data$cluster), 
             ".svg")
    
    plot_savename_scaled <- 
      paste0(file_structure[[g_dataset_id]]$heatmap, "/",
             "heatmap scaled ",
             unique(g_cluster_data$cluster), 
             ".svg")
    
    
    g_cluster_data_wide <- 
      g_cluster_data %>% 
      select(ensg_id, sample_id, tmm_zscore) %>% 
      spread(sample_id, tmm_zscore) %>% 
      column_to_rownames("ensg_id") 
    
    gene_hclust <- 
      g_cluster_data_wide %>% 
      dist() %>% 
      hclust(method = "ward.D2")
    
    sample_hclust <- 
      g_cluster_data_wide %>%
      t() %>% 
      dist() %>% 
      hclust(method = "ward.D2")
    
    plot_data <- 
      g_cluster_data %>% 
      mutate(ensg_id = factor(ensg_id, 
                              rev(gene_hclust$labels[gene_hclust$order])),
             sample_id = factor(sample_id, 
                                sample_hclust$labels[sample_hclust$order])) 
    
    plot_confidence <- 
      cluster_membconf_data %>%
      filter(dataset_id == g_dataset_id) %>% 
      filter(ensg_id %in% gene_hclust$labels) %>% 
      mutate(ensg_id = factor(ensg_id, 
                              rev(gene_hclust$labels[gene_hclust$order]))) %>% 
      ggplot(aes("Confidence", ensg_id, fill = cluster_membership_confidence)) + 
      geom_tile_rast() +
      theme_void() +
      scale_fill_gradient(low = "#D1EEEA", high = "#2A5674",
                          limits = c(0.5, 1),
                          breaks = c(0.5, 0.75, 1),
                          name = "Confidence") +
      theme(axis.text.x = element_text(angle = -90, face = "bold"), 
            legend.position = "left")
    
    plot_heat_zscore <- 
      plot_data %>% 
      ggplot(aes(sample_id, ensg_id, fill = tmm_zscore)) +
      geom_tile_rast() +
      scale_fill_gradient2(low = "#4575B4", high = "#D73027", mid = "#FFFFFF",
                           name = "Z-score") +
      theme_void() +
      theme(axis.text.x = element_text(angle = -90,
                                       hjust = 0, 
                                       size = 6),
            legend.position = "top")
    
    
    plot_heat_scaled <- 
      plot_data %>% 
      ggplot(aes(sample_id, ensg_id, fill = tmm_scaled)) +
      geom_tile_rast() +
      scale_fill_gradient2(low = "#4575B4", high = "#D73027", mid = "#FFFFFF",
                           name = "Relative Expression",
                           breaks = c(0, 0.5, 1),
                           limits = c(0, 1)) +
      theme_void() +
      theme(axis.text.x = element_text(angle = -90,
                                       hjust = 0,
                                       size = 6),
            legend.position = "top")
      
    column_width <- (1/53)
    plot_n <- length(sample_hclust$labels)
    plot_scaling_factor <- column_width * plot_n
    
    plot_widths <- 
      c(3, plot_n) %>% 
      {. * column_width} %>% 
      {. / max(.)}
    
    plot_width <- 
      1 + 4 * plot_scaling_factor
    
    p1 <- 
      plot_confidence + plot_heat_zscore +
      plot_layout(widths = plot_widths, nrow = 1)
    p2 <- 
      plot_confidence + plot_heat_scaled +
      plot_layout(widths = plot_widths, nrow = 1)
    
    ggsave(plot_savename_zscore, plot = p1,  
           width = plot_width, 
           height = 5)
    ggsave(plot_savename_scaled, plot = p2,
           width = plot_width, 
           height = 5)
    
    tibble()
  }) %>% 
  ungroup() %>% 
  collect()

rm(worker_cluster)
```

```{r}
# g_cluster_data <<- .
# g_dataset_id <- unique(g_cluster_data$dataset_id)


plot_savename <- 
  paste0("heatmap original ",
         unique(g_cluster_data$cluster), 
         ".svg")

plot_savename_raster <- 
  paste0("heatmap raster ",
         unique(g_cluster_data$cluster), 
         ".svg")


g_cluster_data_wide <- 
  g_cluster_data %>% 
  select(ensg_id, sample_id, tmm_zscore) %>% 
  spread(sample_id, tmm_zscore) %>% 
  column_to_rownames("ensg_id") 

gene_hclust <- 
  g_cluster_data_wide %>% 
  dist() %>% 
  hclust(method = "ward.D2")

sample_hclust <- 
  g_cluster_data_wide %>%
  t() %>% 
  dist() %>% 
  hclust(method = "ward.D2")

plot_data <- 
  g_cluster_data %>% 
  mutate(ensg_id = factor(ensg_id, 
                          rev(gene_hclust$labels[gene_hclust$order])),
         sample_id = factor(sample_id, 
                            sample_hclust$labels[sample_hclust$order])) 

plot_confidence <- 
  cluster_membconf_data %>%
  filter(dataset_id == g_dataset_id) %>% 
  filter(ensg_id %in% gene_hclust$labels) %>% 
  mutate(ensg_id = factor(ensg_id, 
                          rev(gene_hclust$labels[gene_hclust$order]))) %>% 
  ggplot(aes("Confidence", ensg_id, fill = cluster_membership_confidence)) + 
  geom_tile_rast() +
  theme_void() +
  scale_fill_gradient(low = "#D1EEEA", high = "#2A5674",
                      limits = c(0.5, 1),
                      breaks = c(0.5, 0.75, 1),
                      name = "Confidence") +
  theme(axis.text.x = element_text(angle = -90, face = "bold"), 
        legend.position = "left")

plot_heat_rast <- 
  plot_data %>% 
  ggplot(aes(sample_id, ensg_id, fill = tmm_zscore)) +
  geom_tile_rast() +
  scale_fill_gradient2(low = "#4575B4", high = "#D73027", mid = "#FFFFFF",
                       name = "Z-score") +
  theme_void() +
  theme(axis.text.x = element_text(angle = -90,
                                   hjust = 0, 
                                   size = 6),
        legend.position = "top")

plot_heat <- 
  plot_data %>% 
  ggplot(aes(sample_id, ensg_id, fill = tmm_zscore)) +
  geom_tile() +
  scale_fill_gradient2(low = "#4575B4", high = "#D73027", mid = "#FFFFFF",
                       name = "Z-score") +
  theme_void() +
  theme(axis.text.x = element_text(angle = -90,
                                   hjust = 0, 
                                   size = 6),
        legend.position = "top")



column_width <- (1/53)
plot_n <- length(sample_hclust$labels)
plot_scaling_factor <- column_width * plot_n

plot_widths <- 
  c(3, plot_n) %>% 
  {. * column_width} %>% 
  {. / max(.)}

plot_width <- 
  1 + 4 * plot_scaling_factor

plot_rast <- 
  plot_confidence + plot_heat_rast +
  plot_layout(widths = plot_widths, nrow = 1)
plot_ <- 
  plot_confidence + plot_heat +
  plot_layout(widths = plot_widths, nrow = 1)

ggsave(plot_savename_raster, plot = plot_rast,  
       width = plot_width, 
       height = 5)
ggsave(plot_savename, plot = plot_,
       width = plot_width, 
       height = 5)

tibble()
```


# Old

```{r}

fischer_res_GO <- 
  clustering_data %>% 
  filter(dataset_id == "tissue") %>% 
  group_by(dataset_id) %>% 
  do({
    
    cluster_membership <- 
      set_names(.$cluster,
                .$gene)
    
    GO_terms_moderated %>%
      group_by(GO_domain) %>% 
      do({
          term_membership <-
            distinct(.) %$% 
            split(ensg_id, GO)
          
          gene_fischer_test(cluster_membership, term_membership)
      })
    
  }) %>% 
  ungroup()

fischer_res_GO %>% 
  select(dataset_id, GO_domain) %>% 
  distinct()


fischer_res_GO %>% 
  left_join(GO_terms_dict,
            by = c("term" = "GO")) %>% 
  select(1,2,3,4, term_name, everything()) #%>% 
  # View
```
## Fischer

```{r}
# 
# spec_class %>% filter(type == "tissue" | type == "singlecell") %>% 
#   filter(elevated %in% c("Early spermatids", "intestine")) %>% spread(type, elevated) %>% 
#   filter(complete.cases(.)) %>% 
#   left_join(gene_info %>% 
#               select(ensg_id, gene_name, gene_description)) %>% 
#   left_join(clustering_data %>% 
#               filter(dataset_id == "tissue"),
#             by = c("ensg_id" = "gene")) %>% 
#   View

gene_fischer_test <- 
  function(cluster_membership, term_membership, alternative = "greater") {
    n_genes <- 
      n_distinct(names(cluster_membership))
    
    term_membership_moderated <-  
      term_membership %>% 
      enframe("term", "gene") %>%
      unnest(cols = gene) %>% 
      distinct() %>% 
      filter(gene %in% names(cluster_membership)) %$% 
      split(gene, term)
    
    
    joined_genes_terms <- 
      cluster_membership %>% 
      enframe("gene", "cluster") %>% 
      full_join(term_membership_moderated %>%
                  enframe("term", "gene") %>% 
                  unnest(cols = gene), 
                by = "gene") %>% 
      mutate(term = factor(term),
             cluster = factor(cluster))
    
    
    fischer_data1 <- 
      joined_genes_terms %>%
      filter(!is.na(term)) %>% 
      group_by(cluster, term) %>% 
      count() %>% 
      ungroup()
    
    fischer_data2 <- 
      joined_genes_terms %>%
      filter(!is.na(term)) %>% 
      group_by(term) %>% 
      summarise(total_term = n_distinct(gene),
                total_nonterm = n_genes - total_term) %>% 
      ungroup()
    
    fischer_data3 <- 
      joined_genes_terms %>% 
      group_by(cluster) %>% 
      summarise(total_in_cluster = n_distinct(gene),
                total_not_in_cluster = n_genes - total_in_cluster) %>% 
      ungroup()
    
    
    fisher_data <-
      fischer_data1 %>% 
      left_join(fischer_data2,
                by = "term") %>% 
      left_join(fischer_data3,
                by = "cluster") %>% 
      mutate(incluster_nonterm = total_in_cluster - n,
             notcluster_term = total_term - n,  
             notcluster_nonterm = total_nonterm - notcluster_term)
    
    
    fischer_res <- 
      fisher_data %>% 
      group_by(cluster, term) %>% 
      do({
        g_data <<- .
        
        contingency_matrix <-
          matrix(c(g_data$n, 
                   g_data$incluster_nonterm,
                   g_data$notcluster_term,  
                   g_data$notcluster_nonterm), 
                 nrow = 2,
                 dimnames =
                   list(c("Term", "Nonterm"),
                        c("In cluster", "Not in cluster")))
        
        contingency_matrix %>% 
          fisher.test(alternative = alternative) %>% 
          tidy() %>%
          mutate(term_overlap = paste(contingency_matrix[1], 
                                      contingency_matrix[1] + contingency_matrix[3],
                                      sep = "/"),
                 cluster_membership = paste(contingency_matrix[1], 
                                            contingency_matrix[1] + contingency_matrix[2],
                                            sep = "/"))
      }) %>% 
      ungroup() %>%
      mutate(p.value = ifelse(p.value == 0, .Machine$double.xmin, p.value),
             adj_p = p.adjust(p.value, method = "BH")) %>% 
      rename(odds_ratio = estimate) %>%
      arrange(adj_p)
    
    return(fischer_res)
  }

```

## Cluster levels

```{r}
cluster_levels <- 
  function(x, y, z, distance_metric = "euclidean", clustering_method = "ward.D2", fill = 0) {
    
    if(length(x) < 2) {
      unique(x)
    } else {
      tibble(x, y, z) %>% 
        spread(y, z, fill = 0) %>% 
        column_to_rownames("x") %>% 
        dist(method = distance_metric) %>% 
        hclust(method = clustering_method) %$%
        labels[order] 
    }
    
  }
```

## Cluster enrichement barplots

```{r}

plot_data <- 
  all_res %>% 
  filter(!grepl("original", test_type)) %>% 
  mutate(Description = ifelse(str_length(Description) > 50,
                              paste0(substr(Description, 0, 50), "..."),
                              Description)) %>% 
  group_by(dataset_id, cluster, test_type) %>% 
  top_n(20, -p.adjust) %>% 
  ungroup() %>% 
  mutate(`-log10(p-value)` = -log10(p.adjust)) %>% 
  select(test_type, dataset_id, cluster, ID, Description, 
         `Fraction of cluster` = GeneFrac, 
         `-log10(p-value)`) %>%
  gather(value_type, value, 
         `Fraction of cluster`, 
         `-log10(p-value)`)
  

plot_settings <- 
  plot_data %>% 
  select(dataset_id, cluster) %>% 
  distinct() %>% 
  arrange(dataset_id, as.numeric(cluster)) %>% 
  mutate(i = row_number())
  
plots <- 
  lapply(plot_settings$i,
         function(i) {
           plot_dataset <- plot_settings$dataset_id[i]
           plot_cluster <- plot_settings$cluster[i]
           
           plot_data %>% 
             filter(cluster == plot_cluster,
                    dataset_id == plot_dataset) %>% 
             mutate(Description = factor(Description, 
                                         filter(., value_type == "-log10(p-value)") %>% 
                                           arrange(-value) %>% 
                                           pull(Description) %>% 
                                           unique())) %>% 
             
             ggplot(aes(value, Description)) +
             geom_col(fill = "orangered") +
             # geom_point() +
             facet_grid(test_type ~ value_type, scales = "free", space = "free_y") +
             scale_x_continuous(expand = expansion(c(0,0.05))) +
             theme_bw() +
             theme(strip.text.y = element_text(angle = 0)) +
             labs(y = "Term") +
             ggtitle(paste0(plot_dataset, " cluster ", plot_cluster))
         })

pdf(paste0("results/cluster enrichment plots.pdf"),
    width = 10, height = 14)
plots
dev.off()

```

## Compare clusterings

```{r}


comparison_settings <- 
  clustering_data %>% 
  select(dataset_id, cluster) %>% 
  distinct() %>% 
  unite(cluster_id1, dataset_id, cluster) %>% 
  expand_grid(cluster_id2 = .$cluster_id1) %>% 
  filter(cluster_id1 < cluster_id2) %>% 
  separate(cluster_id1, into = c("dataset_id1", "cluster1"), sep = "_") %>% 
  separate(cluster_id2, into = c("dataset_id2", "cluster2"), sep = "_") %>% 
  filter(dataset_id1 != dataset_id2)


clustering_overlap <- 
  clustering_data %>% 
  select(dataset_id1 = dataset_id) %>% 
  distinct() %>% 
  expand_grid(dataset_id2 = .$dataset_id1) %>% 
  filter(dataset_id1 < dataset_id2) %>% 
  group_by_all() %>% 
  do({
    g_data <<- .
    
    clustering_data %>% 
      select(dataset_id, gene, cluster) %>% 
      filter(dataset_id %in% c(g_data$dataset_id1, g_data$dataset_id2)) %>% 
      spread(dataset_id, cluster) %>% 
      select(gene, 
             cluster1 = g_data$dataset_id1,
             cluster2 = g_data$dataset_id2) %>% 
      group_by(cluster1) %>% 
      mutate(n1 = n_distinct(gene)) %>% 
      group_by(cluster2) %>% 
      mutate(n2 = n_distinct(gene)) %>% 
      group_by(cluster1, cluster2, n1, n2) %>% 
      summarise(ncomb = n_distinct(gene), .groups = "keep") %>% 
      ungroup() %>% 
      mutate(frac = ncomb / ifelse(n1 < n2, 
                                   n1, 
                                   n2)) %>% 
      arrange(-frac)
  }) %>% 
  ungroup()


clustering_overlap %>% 
  filter(!(is.na(cluster1) | is.na(cluster2))) %>% 
  arrange(-frac) %>% 
  write_csv("results/cluster_overlap_data.csv")


clustering_overlap %>% 
  ggplot(aes(frac)) +
  geom_density()


clustering_overlap_data <- 
  clustering_data %>% 
  select(-resolution) %>% 
  unite(cluster_id, dataset_id, cluster) %>% 
  mutate(i = 1) %>% 
  spread(cluster_id, i, fill = 0) %>% 
  column_to_rownames("gene")


clustering_overlap_hclust <- 
  clustering_overlap_data %>% 
  t() %>% 
  dist() %>% 
  hclust(method = "ward.D2")

# tibble(cluster_id = clustering_overlap_hclust$labels,
#        order = clustering_overlap_hclust$order) %>% 
#   arrange(order) %>% 
#   mutate(cluster_id = factor(cluster_id, cluster_id)) %>% 
#   separate(cluster_id, into = c("dataset_id", "cluster"), sep = "_", remove = F) %>% 
#   ggplot(aes(1, cluster_id, fill = dataset_id)) +
#   geom_tile() +
#   coord_fixed()

clustering_overlap %>% 
  filter(frac > 0.2) %>% 
  filter(!(is.na(cluster1) | is.na(cluster2))) %>% 
  unite(cluster_id1, dataset_id1, cluster1) %>% 
  unite(cluster_id2, dataset_id2, cluster2) %>% 
  select(cluster_id1, cluster_id2, frac) %>%
  bind_rows(select(., 
                   cluster_id1 = cluster_id2,
                   cluster_id2 = cluster_id1, 
                   frac))  %>%
  spread(cluster_id1, frac, fill = 0) %>% 
  column_to_rownames("cluster_id2") %>% 
  {
    dat <- .
    diag(dat) <- 1
    dat
  }  %>% 
  pheatmap::pheatmap(clustering_method = "ward.D2",
                     color = colorRampPalette(c("white", "orangered"))(20), 
                     fontsize = 6)


clustering_overlap_pca <- 
  clustering_overlap_data %>% 
  # scale(center = F, scale = colSums(.)) %>% 
  t() %>% 
  calculate_pca(100)
  
clustering_overlap_pca$stats %>% 
  ggplot(aes(PC, R2cum)) +
  geom_line() +
  geom_point()

cluster_sizes <- 
  clustering_data %>% 
  mutate(cluster = as.character(cluster)) %>% 
  group_by(dataset_id, cluster) %>% 
  count() %>% 
  ungroup()

clustering_overlap_pca$scores %>% 
  as_tibble(rownames = "cluster_id") %>% 
  separate(cluster_id, into = c("dataset_id", "cluster"), sep = "_", remove = F) %>% 
  left_join(cluster_sizes) %>% 
  ggplot(aes(PC1, PC2, color = dataset_id, label = cluster, size = n)) +
  geom_point() +
  geom_text(color = "black",
            size = 2) +
  theme_bw() +
  coord_fixed()

```


