---
title: "HPA gene clustering"
author: "Max J Karlsson & María Bueno Alvez"
date: "08/06/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Setup
```{r setup, include=FALSE}
library(vroom)
library(tidyverse)
library(magrittr)
library(pbapply)
# library(NOISeq)
library(uwot)
library(igraph)
library(edgebundle)
# library(MASS)
library(viridis)
library(pcaMethods)
library(factoextra)
library(Seurat)
library(uwot)
library(plotly)
library(htmlwidgets)
library(aricode)
library(clValid)
library(infotheo)
library(ggheatmap)
library(multidplyr)
library(clue)
library(ggalluvial)



source("scripts/functions_preprocessing.R")
source("scripts/functions_dimensionalityreduction.R")
source("scripts/functions_clustering.R")
source("scripts/functions_evaluation.R")
source("scripts/cluster rasterizer.R")
source("scripts/generate_gene_clusters.R")
source("scripts/functions_utility.R")
source("scripts/functions_classification.R")

```

## Read data
```{r}

all_data <- 
  readRDS("data/processed/all_cluster_data_mockup2.rds")

```

## Read metadata
```{r}

all_data_meta <- 
  readRDS("data/processed/all_cluster_data_metadata_mockup2.rds")

dataset_metadata <- all_data_meta$dataset_metadata

all_data_meta <- all_data_meta$all_data_meta
  
all_genes_in_data <-
  all_data %>% 
  map(. %>% 
        select(1)) %>% 
  bind_rows() %>% 
  distinct() %>% 
  pull(ensg_id)

```

## TEMP: Generate tissue classification

```{r}
# TO DO: Change this classification for v21

consensus_hierarchy <- read_csv("data/meta/final_consensus_hierarcy.csv")

classification_tissue <- 
  all_data$tissue %>% 
  gather(consensus_tissue, ntpm, -1) %>% 
  hpa_gene_classification(expression_col = "ntpm", 
                          tissue_col = "consensus_tissue", 
                          gene_col = "ensg_id",
                          enr_fold = 4, 
                          max_group_n = 5, 
                          det_lim = 1)

expressed_genes_tissue <- 
  expressed_genes %>% 
  filter(dataset_id == "tissue") %>% 
  pull(ensg_id)

classification_tissue_long <- 
  classification_tissue %>%
  filter(gene %in% expressed_genes_tissue) %>%
  separate_rows(enriched_tissues, sep = ';') %>%
  select(gene,enriched_tissues) %>%
  remove_missing()

```


## Read database data

```{r}

# GO annotation
GOterms <- 
  vroom("data/meta/Ensembl103 GO terms.txt") %>% 
  select(ensg_id = 1, 
         gene_name = 2, 
         GO_accession = 3, 
         GO_domain = 4, 
         GO_term_name = 5, 
         GO_term_evidence = 6) %>% 
  filter(ensg_id %in% all_genes_in_data)


```


## Color data 

```{r}
colors <-
  read_delim("../HPA-normalization2021/data/meta/HPA_colors.txt", delim = "\t")

# colors <-
# read_csv("data/meta/Blood cell colors.csv")

# color_data <-
#   data %>%
#   group_by(celltype, ensg_id) %>% 
#   summarise(tmm = mean(tmm, na.rm = T)) %>%
#   left_join(colors,
#             by = c("celltype")) %>%
#     
#   group_by(ensg_id, color = lineage_color) %>% 
#   summarise(tmm = max(tmm, na.rm = T)) %>% 
#   
#   group_by(ensg_id) %>%
#   mutate(tmm = tmm / sum(tmm, na.rm = T)) %>%
#   ungroup()
# 
# ensg_exp_palette <-
#   color_data %>%
#   bind_cols(col2rgb(.$color) %>% 
#               t() %>% 
#               as_tibble()) %>%
#   mutate(red = red * tmm,
#          green = green * tmm,
#          blue = blue * tmm) %>%
#   group_by(ensg_id) %>% 
#   summarise(red = sum(red),
#             green = sum(green),
#             blue = sum(blue)) %>%
#   filter(complete.cases(.)) %>%
#   mutate(color = rgb(red, green, blue, maxColorValue = 255))

dataset_palette <- c("tissue" = "#0083C4", "singlecell" = "#6AA692", "brain" = "#FFDD00", "celline" = "#97CF16", 
                     "celline_consensus" = "#97CF16", "blood" = "#CF161A", "blood_consensus" = "#CF161A")
   
```


## Run ID

```{r}

# This run ID is used to check whether the run is already performed. 
# In that case files will be reloaded instead of rerunning calculations.
run_id <- 
  "new_aggregation_100seeds"
run_file <- 
  paste0("data/processed/clusters_", run_id, ".rds")


run_id_savename <- 
  . %>% 
  paste0("results/", run_id, " ", .)
```



# Clustering
```{r}

gene_max_exp <- 
  all_data %>% 
  map(. %>%
        column_to_rownames("ensg_id") %>% 
        apply(MARGIN = 1,
              max) %>% 
        enframe("ensg_id", "max_exp")) %>% 
  bind_rows(.id = "dataset_id")

expressed_genes <- 
  gene_max_exp %>% 
  filter(max_exp >= 1) %>% 
  select(1, 2)

# Prepare cluster for parallelization
cluster <- new_cluster(n = 6)
cluster_library(cluster, c("dplyr",
                           "tidyverse",
                           "factoextra",
                           "pcaMethods",
                           "Seurat",
                           "magrittr",
                           "cluster",
                           "clValid")) 
cluster_copy(cluster, c("HPA_gene_clustering", 
                        "generate_savefile", 
                        "scale_data",
                        "calculate_pca",
                        "calculate_distance",
                        "cluster_genes",
                        "run_id", 
                        "run_file", 
                        "run_id_savename"))


if(file.exists(run_file)) {
  all_data_cluster <- 
    readRDS(run_file)
} else {
  all_data_cluster <- 
    all_data %>% 
    map(. %>% 
          gather(sample_id, tmm, -1)) %>% 
    bind_rows(.id = "dataset_id") %>% 
    inner_join(expressed_genes) %>% 
    group_by(dataset_id) %>%      
    partition(cluster) %>% 
    do({
      g_data <<- .
      
      cluster_res <-
        HPA_gene_clustering(g_data,
                            col_value = "tmm",
                            col_gene = "ensg_id",
                            col_sample = "sample_id",
                            scaling = "zscore",
                            distance_metric = "spearman",
                            clustering_method = "louvain",
                            seed = 1:100,
                            resolution = seq.int(1, 8, by = 0.5),
                            npcs = 100,
                            log_transform = F,
                            run_id = paste(run_id,
                                           unique(g_data$dataset_id),
                                           sep = "_"))

      cluster_res$cluster_data
    }) %>%
    collect() %>% 
    ungroup()
  
  saveRDS(all_data_cluster, 
          run_file)
}

pal<-
  colorRampPalette(ggthemes::tableau_color_pal(palette = "Classic Cyclic")(13))(n_distinct(all_data_cluster$cluster))

cluster_palette50 <- 
  colorRampPalette(ggthemes::tableau_color_pal(palette = "Classic Cyclic")(13))(50)

cluster_palette75 <- 
  colorRampPalette(ggthemes::tableau_color_pal(palette = "Classic Cyclic")(13))(75)

cluster_palette100 <- 
  colorRampPalette(ggthemes::tableau_color_pal(palette = "Classic Cyclic")(13))(100)


cluster_palette_tbl <- 
  cluster_palette %>% 
  enframe("cluster", "color") %>% 
  mutate(cluster = as.character(cluster - 1))

all_data_graphs <- 
  list.files("data/processed/", full.names = T) %>% 
  enframe("i", "filename") %>% 
  filter(grepl(run_id, filename)) %>% 
  filter(grepl("neighbors", filename)) %>% 
  mutate(dataset_id = gsub(paste0(".*", run_id, "_|_neighbors_.*$"), "", filename)) %$% 
  set_names(filename, dataset_id) %>%
  lapply(function(filename_) readRDS(filename_))

# Distance files are very big, so just read when used:
all_data_dists_filenames <- 
  list.files("data/processed/", full.names = T) %>% 
  enframe("i", "filename") %>% 
  filter(grepl(run_id, filename)) %>% 
  filter(grepl("_dist_", filename)) %>% 
  mutate(dataset_id = gsub(paste0(".*", run_id, "_|_dist_.*$"), "", filename)) %$% 
  set_names(filename, dataset_id) 

```

## Clustering association

```{r}

# 
# gene_association_data <- 
#   data_cluster$cluster_data %>% 
#   filter(resolution == 1)
# 
# gene_association <-
#   gene_association_data %>% 
#   group_by(gene) %>% 
#   do({
#     g_data <<- .
#     
#     g_data %>% 
#       left_join(gene_association_data,
#                 by = c("resolution", "seed", "value"),
#                 suffix = c("1", "2")) %>% 
#       group_by(gene2) %>% 
#       count() %>% 
#       ungroup()
#   })
# 
# 
# gene_association %>% ungroup() %>% 
#   mutate(frac = n / 10) %>%
#   group_by(frac) %>% 
#   count() %>%
#   ggplot(aes(frac, n)) +
#   geom_col()


```


# Evaluation

## ARI
```{r Retrieve partitions for ARI}

ARI_file <- 
  paste0("data/processed/", run_id, "_ARI_res.rds")

# ARI_comparisons <- 
#   all_data_cluster %>% 
#   select(dataset_id, resolution, seed) %>% 
#   distinct() %>% 
#   unite(resolution_seed, resolution, seed) %>% 
#   group_by(dataset_id) %>% 
#   do({
#     tidyr::crossing(resolution_seed1 = .$resolution_seed,
#                     resolution_seed2 = .$resolution_seed) %>% 
#       as_tibble()
#   }) %>% 
#   separate(resolution_seed1,
#            into = c("resolution1", "seed1"), sep = "_") %>% 
#   separate(resolution_seed2,
#            into = c("resolution2", "seed2"), sep = "_") %>% 
#   
#   ungroup() %>% 
#   mutate_at(-1, .funs = as.numeric)
# 
# 
# if(file.exists(ARI_file)) {
#   ARI_res <- 
#     readRDS(ARI_file)
# } else {
#   ARI_res <- 
#     ARI_comparisons %>% 
#     group_by(dataset_id) %>% 
#     do({
#       g_data <<- .
#       
#       g_data_cluster <- 
#         all_data_cluster %>% 
#         filter(dataset_id == unique(g_data$dataset_id))
#       
#       
#       g_data %>% 
#         left_join(g_data_cluster,
#                   by = c("resolution1" = "resolution",
#                          "seed1" = "seed")) %>% 
#         left_join(g_data_cluster,
#                   by = c("resolution2" = "resolution",
#                          "seed2" = "seed",
#                          "gene"),
#                   suffix = c("1", "2")) %>% 
#         group_by(resolution1, seed1, resolution2, seed2) %>% 
#         summarise(ARI = ARI(cluster1,
#                             cluster2))
#       
#     }) %>% 
#     ungroup()
#     
#   
#   saveRDS(ARI_res, 
#           ARI_file)
# }
# 
# ARI_res %>% 
#   filter(resolution1 == resolution2,
#          seed1 < seed2) %>% 
#   ggplot(aes(resolution1, ARI, color = dataset_id)) +
#   geom_point() +
#   geom_smooth() +
#  # scale_x_log10() +
#   facet_grid(dataset_id~.) +
#   theme_bw() +
#   ggtitle("Replicate clustering stability")
# 
# 
# ARI_res %>% 
#   filter(dataset_id == "tissue") %>% 
#   # filter(resolution1 == 1, resolution2 == 2) %>% 
#   unite(resolution_seed1, resolution1, seed1) %>% 
#   unite(resolution_seed2, resolution2, seed2) %>% 
#   select(-1) %>% 
#   spread(resolution_seed2, ARI) %>% 
#   column_to_rownames("resolution_seed1") %>% 
#   pheatmap()


# Alternative
df_to_vector <- function(cluster_df) {
  v <- cluster_df$cluster
  names(v) <- cluster_df$gene
  return(v)
}

if(file.exists(ARI_file)) {
  ARI_res <- 
    readRDS(ARI_file)
} else {
  ARI_res <- 
    all_data_cluster %>% 
    group_by(dataset_id, resolution) %>% 
    do({
      dat <<- .
      
      combn(1:max(dat$seed), m=2) %>% 
        t() %>% 
        as_tibble() %>% 
        set_names(c("seed1", "seed2")) %>% 
        group_by_all() %>% 
        mutate(ARI = ARI(dat %>% filter(seed == seed1) %>% select(gene,cluster) %>%  df_to_vector(), 
                         dat %>%  filter(seed == seed2) %>% select(gene,cluster) %>%  df_to_vector()))
      
    }) %>% 
    ungroup()
  
  saveRDS(ARI_res, 
          ARI_file)
}

ARI_res %>% 
  ggplot(aes(resolution, ARI, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~dataset_id, ncol = 6) +
  theme_bw() +
  ggtitle("Replicate clustering stability")

```

## Connectivity & Average Silhouette 
```{r Evaluation (connectivity, dunn, asw, BHI, BP_enriched, time)}

connectivity_file <- 
  paste0("data/processed/", run_id, "_connectivity.rds")
# silhouette_file <- 
#   paste0("data/processed/", run_id, "_silhouette.rds")

# Prepare cluster for parallelization
cluster <- new_cluster(n = 6)

cluster_library(cluster, c("dplyr",
                           "tidyverse",
                           "clValid",
                           "cluster"))

cluster_copy(cluster, c("connectivity_file",
                        "all_data_dists_filenames",
                        "all_data_cluster"))

Sys.time()

if(file.exists(connectivity_file)) {
  all_data_cluster_connectivity <- 
    readRDS(connectivity_file)
} else {
  
  
  all_data_cluster_connectivity <- 
    all_data_cluster %>% 
    group_by(dataset_id) %>% 
    partition(cluster) %>% 
    do({
      g_data <<- .
      g_dataset_id <- 
        unique(g_data$dataset_id)
      
      cat(paste("Running dataset:", g_dataset_id, "\n"))
      
      
      connectivity_dataset_file <-
        connectivity_file %>% 
        gsub(".rds$", "", .) %>% 
        paste0("_", g_dataset_id, ".rds") 
      
      
      if(file.exists(connectivity_dataset_file)) {
        cat(paste("  Files already present.\n"))
      
        g_res <- 
          readRDS(connectivity_dataset_file)
      } else {
        
        cat(paste("  Calculating metrics.\n"))
        
        
        temp_distan <- 
          readRDS(all_data_dists_filenames[g_dataset_id])
        
        g_res <- 
          g_data %>% 
          group_by(resolution, seed) %>% 
          summarise(connectivity = connectivity(distance = temp_distan, 
                                                cluster = cluster),
                    avg_silhouette_width = silhouette(dist = temp_distan, 
                                                      x = cluster) %>% 
                      summary() %>% 
                      {.$avg.width})
        # dunn = dunn(distance = temp_distan, 
        #             cluster = value))
        
        saveRDS(g_res, connectivity_dataset_file)
        
        
      }
      
      g_res
    }) %>% 
    collect() %>% 
    ungroup()
  
  
  saveRDS(all_data_cluster_connectivity, 
          connectivity_file)
}

Sys.time()


# if(file.exists(silhouette_file)) {
#   all_data_cluster_silhouette <- 
#     readRDS(silhouette_file)
# } else {
#   
#   all_data_cluster_silhouette <- 
#     all_data_cluster %>% 
#     group_by(dataset_id) %>% 
#     do({
#       g_data <<- .
#       
#       temp_distan <- 
#         readRDS(all_data_dists_filenames[unique(g_data$dataset_id)])
#       
#       
#       g_data %>% 
#         filter(resolution == 1, 
#                seed == 1) %>% 
#         group_by(resolution, seed) %>% 
#         summarise(silhouette = silhouette(dist = temp_distan, 
#                                           x = .$value) %>% 
#                     summary() %>% 
#                     {.$avg.width})
#           
#         })
#       
#       
#       a
#       
#       silhouette = silhouette(dist = d, 
#                               x = value)
#     })
#   
# 
#   
#   saveRDS(all_data_cluster_connectivity, 
#           silhouette_file)
# }


all_data_cluster_connectivity %>%
  gather(metric, value, connectivity, avg_silhouette_width) %>% 
  ggplot(aes(resolution, value, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
 # scale_x_log10() +
  facet_grid(metric ~ dataset_id, scales = "free_y") +
  theme_bw() +
  ggtitle("Connectivity and Silhouette index")



all_data_cluster_connectivity %>% 
  left_join(all_data_cluster %>% 
              group_by(dataset_id,resolution, seed) %>% 
              mutate(k = max(cluster) + 1) %>% 
              select(-gene,-cluster) %>% 
              distinct) %>% 
  gather(metric, value, connectivity, avg_silhouette_width) %>% 
  ggplot(aes(k, value, color = as.factor(resolution))) +
  geom_point() +
  geom_smooth() +
  facet_grid(metric ~ dataset_id, scales = "free_y") +
  theme_bw() +
  ggtitle("Connectivity and Silhouette index")

```

##MI
```{r Evaluation based on Mutual Information}

MI_file <- ## Connectivity & Average Silhouette
  paste0("data/processed/", run_id, "_MI.rds")
MI_random_data_file <- 
  paste0("data/processed/", run_id, "_MI_rand_data.tsv")


MI_GOterms <- 
  GOterms %>%
  filter(!is.na(GO_accession),
         !is.na(GO_domain)) %>%
  distinct() %>% 
  group_by(GO_accession) %>% 
  mutate(n_genes = n_distinct(ensg_id)) %>% 
  ungroup() %>% 
  filter(n_genes <= 500, 
          n_genes >= 10) # %>% 
  # filter(GO_domain == "biological_process") %>% 
  # select(ensg_id, GO_accession, n_genes)


n_random_clustering <- 
  10


if(file.exists(MI_file)) {
  all_data_cluster_MI <- 
    readRDS(MI_file)
} else {
  
  
  all_data_cluster_MI <- 
    all_data_cluster %>% 
    group_by(dataset_id, resolution, seed) %>% 
    do({
      g_data <<- .
      
      cat(paste0(unique(g_data$dataset_id), "\n",
                 "Resolution ", unique(g_data$resolution), "\n",
                 "Seed ", unique(g_data$seed)), "\n")
      
      g_GOterms <- 
        GOterms %>% 
        filter(ensg_id %in% g_data$gene)
      
      if(!file.exists(MI_random_data_file)) {
        tibble(dataset_id = "test",
               resolution = 1,
               seed = 1,
               replicate = 1,
               run_id = "test",
               MI = NA) %>% 
        write_tsv(MI_random_data_file)
        
      }
      rand_data_old <-
        read_tsv(MI_random_data_file,
                 col_types = cols(
                   dataset_id = col_character(),
                   resolution = col_double(),
                   seed = col_double(),
                   replicate = col_double(),
                   run_id = col_character(),
                   MI = col_double()
                 ))
      
      g_runs <- 
        g_data %>% 
        select(1:3) %>% 
        distinct() %>% 
        expand_grid(replicate = 1:n_random_clustering) %>% 
        mutate(run_id = run_id) 
      
      g_runs_torun <- 
        g_runs %>% 
        left_join(rand_data_old,
                  by = c("dataset_id",
                         "resolution",
                         "seed",
                         "replicate",
                         "run_id")) %>% 
        filter(is.na(MI)) 
      
      cat(paste0("  ", n_random_clustering - nrow(g_runs_torun), 
                 " random clustering MI results precalculated.\n",
                 "  Running ", nrow(g_runs_torun)), "\n")
      
      if(nrow(g_runs_torun) != 0) {
        
        rand_MI_data <- 
          g_runs_torun %>% 
          group_by(dataset_id,
                   resolution,
                   seed,
                   replicate,
                   run_id) %>% 
          do({
            set.seed(.$replicate)
            g_data %>% 
              select(gene, 
                     cluster) %>% 
              mutate(cluster = sample(cluster))
          }) %>% 
          ungroup() %>% 
          left_join(g_GOterms,
                    by = c("gene" = "ensg_id")) %>% 
          group_by(dataset_id,
                   resolution,
                   seed,
                   replicate,
                   run_id) %>% 
          summarise(MI = mutinformation(cluster, GO_accession)) 
        
        rand_MI_data %>% 
          write_tsv(MI_random_data_file, append = T)
      } 
      
      
      
      
      
      rand_data <-
        read_tsv(MI_random_data_file, 
                 col_types = cols(
                   dataset_id = col_character(),
                   resolution = col_double(),
                   seed = col_double(),
                   replicate = col_double(),
                   run_id = col_character(),
                   MI = col_double()
                 )) %>% 
        inner_join(g_runs, 
                   by = c("dataset_id",
                          "resolution",
                          "seed",
                          "replicate",
                          "run_id"))
      
      
      g_data %>% 
        left_join(g_GOterms,
                  by = c("gene" = "ensg_id")) %>% 
        summarise(MI = mutinformation(cluster, GO_accession)) %>% 
        bind_cols(rand_data %>% 
                    ungroup() %>% 
                    summarise(mean = mean(MI), 
                              sd = sd(MI))) %>% 
        mutate(zscore = (MI - mean)/sd)
      
      
    }) %>% 
    ungroup()
  
  saveRDS(all_data_cluster_MI, 
          MI_file)
}

all_data_cluster_MI %>%
  ggplot(aes(resolution, zscore, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
  scale_x_log10() +
  facet_grid( ~ dataset_id, scales = "free_y") +
  theme_bw() +
  ggtitle("Mutual information")



# Alternative

MI_file <- ## Connectivity & Average Silhouette
  paste0("data/processed/", run_id, "_MI.rds")
MI_random_data_file <- 
  paste0("data/processed/", run_id, "_MI_rand_data.tsv")

n_random_clustering <- 
  50

# Prepare cluster for parallelization
cluster <- new_cluster(n = 6)

cluster_library(cluster, c("dplyr",
                           "purrr",
                           "tidyverse",
                           "infotheo"))

cluster_copy(cluster, c("MI_random_data_file",
                        "n_random_clustering",
                        "all_data_cluster",
                        "MI_GOterms"))

# 1. Create random_MI file

if(file.exists(MI_random_data_file)) {
  rand_MI_data <- 
    read_tsv(MI_random_data_file)
} else {
  random_clustering <- 
    map(c(1:n_random_clustering), function(x)
      all_data_cluster %>% 
        filter(seed == 1) %>% 
        group_by(dataset_id, resolution) %>% 
        do ({
          dat <<- .
          dat %>% 
            mutate(cluster = sample(cluster))
        }) %>% 
        mutate(run = x)) %>% 
    purrr::reduce(full_join)
  
  rand_MI_data <- 
    random_clustering %>% 
    group_by(dataset_id, resolution, seed, run) %>% 
 #   partition(cluster) %>% 
    do({
      g_data <<- .
      
      g_GOterms <- 
        MI_GOterms %>% 
        filter(ensg_id %in% g_data$gene) %>% 
        filter(GO_domain  == "biological_process")
      
      g_data %>%
        ungroup() %>% 
        left_join(g_GOterms,
                  by = c("gene" = "ensg_id")) %>%
        summarise(MI = mutinformation(cluster, GO_accession)) 
    }) %>% 
    collect() %>% 
  #  ungroup() %>% 
    group_by(dataset_id, resolution, seed) %>% 
    summarise(mean = mean(MI), 
              sd = sd(MI))%>% 
     select(-seed)
  
  write_tsv(rand_MI_data, MI_random_data_file)
}

# 2. Calculate MI in parallel for all clusterings 
cluster <- new_cluster(n = 6)

cluster_library(cluster, c("dplyr",
                           "tidyverse",
                           "infotheo"))

cluster_copy(cluster, c("MI_file",
                        "all_data_cluster",
                        "rand_MI_data",
                        "MI_GOterms"))

if(file.exists(MI_file)) {
  all_data_cluster_MI <- 
    readRDS(MI_file)
} else {
   
   all_data_cluster_MI <-
     all_data_cluster %>% 
    group_by(dataset_id, resolution, seed) %>% 
     partition(cluster) %>% 
    do({
      g_data <<- .
      
      g_GOterms <- 
        MI_GOterms %>% 
        filter(ensg_id %in% g_data$gene) %>% 
        filter(GO_domain  == "biological_processes")
      
      g_data %>%
        ungroup() %>% 
        left_join(g_GOterms,
                  by = c("gene" = "ensg_id")) %>%
        summarise(MI = mutinformation(cluster, GO_accession))  
      }) %>% 
     collect() %>% 
    ungroup() %>%
        left_join(rand_MI_data) %>% 
       mutate(zscore = (MI - mean)/sd) 
  
  saveRDS(all_data_cluster_MI, 
          MI_file)
}

all_data_cluster_MI %>%
  ggplot(aes(resolution, zscore, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
  facet_grid( ~ dataset_id, scales = "free_y") +
  theme_bw() +
  ggtitle("Mutual information")

all_data_cluster_MI %>% 
  left_join(all_data_cluster %>% 
              group_by(dataset_id,resolution, seed) %>% 
              mutate(k = max(cluster) + 1) %>% 
              select(-gene,-cluster) %>% 
              distinct) %>% 
  ggplot(aes(k, zscore, color = k)) +
  geom_point() +
  geom_smooth() +
  facet_grid(~dataset_id, scales = "free_y") +
  theme_bw() +
  ggtitle("Mutual information")


##########
MI_GOterms <- 
  GOterms %>%
  filter(!is.na(GO_accession),
         !is.na(GO_domain)) %>%
  distinct() %>% 
  group_by(GO_accession) %>% 
  mutate(n_genes = n_distinct(ensg_id)) %>% 
  ungroup() %>% 
  filter(n_genes <= 500, 
          n_genes >= 10)

calculate_MI <-
  function(clustering, GO, nrep = 10, random_simplify = F) {
    overlap_GO <-
      GO %>%
      filter(ensg_id %in% clustering$gene)
    overlap_clustering <-
      clustering %>%
      filter(gene %in% overlap_GO$ensg_id)
    # t <- Sys.time()
    random_clustering_MI <-
      overlap_GO %>%
      group_by(GO_domain) %>%
      do({
        domain_GOterms <- .
        if(random_simplify) {
          domain_GOterms <-
            domain_GOterms %>%
            group_by(ensg_id) %>%
            summarise(GO_accession = sample(GO_accession, 1)) %>%
            ungroup()
        }
        expand_grid(run = 1:nrep) %>%
          group_by(run) %>%
          do({
            overlap_clustering %>%
              mutate(cluster = sample(cluster)) %>%
              inner_join(domain_GOterms,
                         by = c("gene" = "ensg_id")) %>%
              summarise(MI = mutinformation(cluster, GO_accession))
          })
      }) %>%
      ungroup()
    # Sys.time() - t
    random_clustering_MI_specs <-
      random_clustering_MI %>%
      group_by(GO_domain) %>%
      summarise(mean_MI = mean(MI),
                sd_MI = sd(MI))
    overlap_clustering_MI <-
      overlap_GO %>%
      group_by(GO_domain) %>%
      do({
        domain_GOterms <- .
        if(random_simplify) {
          domain_GOterms <-
            domain_GOterms %>%
            group_by(ensg_id) %>%
            summarise(GO_accession = sample(GO_accession, 1)) %>%
            ungroup()
        }
        overlap_clustering %>%
          inner_join(domain_GOterms,
                     by = c("gene" = "ensg_id")) %>%
          summarise(MI = mutinformation(cluster, GO_accession))
      }) %>%
      ungroup() %>%
      left_join(random_clustering_MI_specs,
                by = "GO_domain") %>%
      mutate(z_score = (MI - mean_MI) / sd_MI)
    return(overlap_clustering_MI)
  }

t <- Sys.time()
worker_cluster <- new_cluster(n = 6)
cluster_library(worker_cluster, c("dplyr",
                                  "tidyverse",
                                  "infotheo"))
cluster_copy(worker_cluster, c("MI_GOterms",
                               "calculate_MI"))
MI_res_iter_cosensus <-
  clustering_data_consensus %>%
  group_by(dataset_id,resolution) %>%
  partition(worker_cluster) %>%
  do({
    g_data <- .
    select(g_data, gene, cluster) %>%
      calculate_MI(., MI_GOterms, nrep = 1000)
  }) %>%
  ungroup() %>%
  collect()
rm(worker_cluster)
gc()
Sys.time() - t


saveRDS(MI_res_iter_cosensus, "data/processed/MI_res_iter_consensus.rds")

MI_res_iter_cosensus %>%
  ggplot(aes(resolution, z_score, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
  facet_grid(GO_domain ~ dataset_id, scales = "free_y") +
  theme_bw() +
  ggtitle("Mutual information")
```


## Summarize evaluation

```{r}

# Metrics ~ k
ARI_res %>% 
  #mutate(seed = NA) %>% 
  gather(metric, value, ARI) %>% 
  select(-seed2) %>% 
  rename(seed = seed1) %>% 
  full_join(all_data_cluster_connectivity %>% 
              left_join(all_data_cluster_MI %>%  select(-MI, -mean, -sd)) %>%
              gather(metric, value, connectivity, avg_silhouette_width, zscore)) %>% 
    left_join(all_data_cluster %>% 
              group_by(dataset_id,resolution, seed) %>% 
              mutate(k = max(cluster) + 1) %>% 
              select(-gene,-cluster) %>% 
              distinct) %>% 
  ggplot(aes(k, value, color = dataset_id)) +
  geom_point() +
  geom_smooth() +  
  facet_grid(metric ~ dataset_id, scales = "free_y") +
  theme_minimal() +
  ggtitle("Clustering evaluation")

#Number of clusters per resolution
all_data_cluster %>% 
  group_by(dataset_id, resolution, seed) %>%  
  summarize(k = max(cluster)) %>% 
  ggplot(aes(x = resolution, y = k, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
  theme_bw() +
  facet_wrap(~dataset_id, ncol = 6)

# Summary
all_data_cluster %>% 
  group_by(dataset_id, resolution, seed) %>%  
  summarize(k = max(cluster)) %>% 
  left_join(ARI_res %>% select(-seed1,-seed2)) %>% 
  left_join(all_data_cluster_MI %>% select(-MI,-mean,-sd)) %>% 
  left_join(all_data_cluster_MI_consensus %>% 
              select(-MI,-mean,-sd) %>%  
              rename(zscore_consensus = zscore)) %>% 
  gather(key="measure", value = "value", -dataset_id, -resolution, -seed) %>% 
  ggplot(aes(x = resolution, y = value, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
  theme_bw() +
  facet_grid(measure~dataset_id, scales = "free_y")


all_consensus %>% 
  group_by(dataset_id, resolution) %>%  
  summarize(k = max(cluster)) %>% 
  left_join(ARI_res %>% select(-seed1,-seed2)) %>% 
  left_join(all_data_cluster_MI_BP %>% select(-MI,-mean,-sd) %>% 
              rename(zscore_BP = zscore)) %>% 
  left_join(all_data_cluster_MI_MF %>% select(-MI,-mean,-sd) %>% 
              rename(zscore_MF = zscore)) %>% 
  left_join(all_data_cluster_MI_CC %>% select(-MI,-mean,-sd) %>% 
              rename(zscore_CC = zscore)) %>% 
  gather(key="measure", value = "value", -dataset_id, -resolution) %>% 
  ggplot(aes(x = resolution, y = value, color = dataset_id)) +
  geom_point(size = 0.5) +
  geom_smooth() +
  theme_bw() +
  facet_grid(measure~dataset_id, scales = "free_y") +
  scale_color_manual(values = dataset_palette) 
ggsave("eval_MI_BP_MF_CC_allseeds_HPAcolors.pdf",
       width = 10, height = 10)

```


# Choose a final clustering

```{r}

# # This is just a placeholder. Later we will want to choose clustering based on performance
# final_cluster <-
#   all_data_cluster %>% 
#   filter(resolution == 1,
#          seed == 1) %>% 
#   arrange(cluster) %>% 
#   mutate(cluster = factor(cluster, unique(cluster))) %>% 
#   arrange(dataset_id, gene)



to_cluster <- function(cluster_df) {
  v <- cluster_df$cluster
  names(v) <- cluster_df$gene
  return(v)
}

# Consensus clustering for all clustering solutions

cons_file <- 
  paste0("data/processed/", run_id, "_consensus.rds")

if(file.exists(cons_file)) {
  all_consensus <- 
    readRDS(cons_file)
} else {
 all_consensus <- 
  all_data_cluster %>% 
  group_by(dataset_id, resolution) %>% 
  do ({
    dat <- .
    k <-  dat %>% 
      mutate(cluster = cluster + 1) %>% 
      group_by(seed) %>%  
      select(-gene) %>% 
      unique() %>% 
      group_by(dataset_id, resolution, seed) %>%  
      add_count(seed) %>% 
      select(resolution, seed, n) %>%  
      unique() %>% 
      pull(n) %>%
      median() %>% 
      ceiling()
    
    clusters <- dat %>% 
      mutate(cluster = cluster + 1) %>% 
      unite(resolution_seed, dataset_id, resolution, seed)  %>% 
      group_split(resolution_seed) %>% 
      map(to_cluster) %>%  
      map(~as.cl_partition(.x))
      
    cons_clustering <- cl_consensus(clusters, method = "SE")
    
    final_clustering <- data.frame(dataset_id = unique(dat$dataset_id),
                                   resolution = unique(dat$resolution), 
                                   gene = names(cl_class_ids(cons_clustering)), 
                                   cluster = as.numeric(cl_class_ids(cons_clustering)))
  })
 saveRDS(all_consensus, 
          cons_file)
}

consensus_match_file <- 
  paste0("data/processed/", run_id, "_consensus_match.rds")

if(file.exists(consensus_match_file)) {
  all_perc <- 
    readRDS(consensus_match_file)
} else {
  all_perc <- 
  all_data_cluster %>% 
  group_by(dataset_id, resolution) %>% 
  do ({
    dat <- .
    clusters <- dat %>% 
      unite(resolution_seed, dataset_id, resolution, seed)  %>% 
      group_split(resolution_seed) %>% 
      map(to_cluster) %>%  
      map(~as.cl_partition(.x))
    
    cons_clustering <- cl_consensus(clusters, method = "SE")
        cons_matrix <- 
      cons_clustering$.Data[,] %>% 
      as_tibble(rownames = "gene")
    
    perfect_match <- 
      cons_matrix %>% 
      gather(key = "cluster", value, -1) %>% 
      filter(value == 1) %>% 
      pull(gene) %>% 
      unique() %>% 
      length() 
    
    perc <- perfect_match / length(cons_matrix$gene)
    
    match_80 <- 
      cons_matrix %>% 
      gather(key = "cluster", value, -1) %>% 
      filter(value >= 0.8) %>% 
      pull(gene) %>% 
      unique() %>% 
      length() 
    
    perc_80 <- match_80 / length(cons_matrix$gene)
    
    final_clustering <- data.frame(dataset_id = unique(dat$dataset_id),
                                   resolution = unique(dat$resolution), 
                                   perfect_match = perc,
                                   match_80 = perc_80)
  })
  saveRDS(all_perc, 
          consensus_match_file)
}

p1 <- all_perc %>% 
  ggplot(aes(x = resolution, y = perfect_match, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~dataset_id, ncol = 6) +
  theme_bw()

p2 <- all_perc %>% 
  ggplot(aes(x = resolution, y = match_80, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~dataset_id, ncol = 6) +
  theme_bw()

p1 / p2

all_perc %>% 
  gather(key = "type", value = "value", -dataset_id, -resolution) %>% 
  ggplot(aes(x = resolution, y = value, color = type)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~dataset_id, ncol = 6) +
  theme_bw()

```

## Intra cluster semantic similarity
```{r}
library(GOSemSim)
d <- godata(OrgDb = "org.Hs.eg.db", keytype = "ENSEMBL", ont = "BP", computeIC = FALSE)


# Prepare cluster for parallelization
cluster_semsim <- new_cluster(n = 6)
cluster_library(cluster_semsim, c("dplyr",
                           "tidyverse",
                           "GOSemSim"))
cluster_copy(cluster_semsim, c("all_consensus",
                               "d"))

Sys.time()
results_semsim <-
  all_consensus %>% 
  group_by(dataset_id, resolution, cluster) %>% 
  partition(cluster_semsim) %>% 
  do ({
    dat <- .
    genes <- dat$gene
    mat <- mgeneSim(genes, semData=d, measure="Wang", combine="BMA", verbose=FALSE)
    avg_sim <- mat[upper.tri(mat)] %>%  mean()
    data.frame(dataset_id = unique(dat$dataset_id), 
               resolution = unique(dat$resolution),
               cluster = unique(dat$cluster),
               sim = avg_sim)
  }) %>% 
 collect()

saveRDS(results_semsim, "data/processed/results_semsim.rds")
Sys.time()

results_semsim <- readRDS("data/processed/results_semsim.rds")

results_semsim %>% 
  spread(key = dataset_id, value = sim)

results_semsim %>% 
  group_by(dataset_id,resolution) %>% 
  mutate( avg = mean(sim)) %>% 
  ggplot(aes(resolution, avg, color = dataset_id)) +
  geom_point() +
  geom_smooth()+
  facet_wrap(~dataset_id, ncol = 6) +
  theme_bw()


saveRDS(results_semsim, 
          "data/processed/results_semsim.rds")

```



## Sensitivity & Specificiy

```{r}

prop <- 
  all_consensus %>% 
    ungroup() %>% 
    left_join(MI_GOterms %>% 
                filter(GO_domain == "biological_process") %>% 
                select(ensg_id, GO_accession), 
              by = c("gene" = "ensg_id")) %>% 
  group_by(dataset_id, resolution, GO_accession) %>%     
  add_count(GO_accession, name = "tot_n") %>% 
  ungroup() %>% 
  group_by(dataset_id, resolution, cluster) %>% 
    do({
      dat <- .
      ngenes <- length(unique(dat$gene))
      dat %>%  mutate (ngenes = ngenes)
    }) %>% 
    group_by(dataset_id, resolution, cluster, GO_accession) %>% 
    add_count(GO_accession) %>%     
    ungroup() %>%  
    mutate(sensitivity_prop = n/ngenes,
           selectivity_prop = n/tot_n) %>% 
    group_by(dataset_id, resolution, cluster) %>% 
    summarise(mean(sensitivity_prop), mean(selectivity_prop)) 

prop %>% 
  ungroup() %>%  
  group_by(dataset_id, resolution) %>% 
  summarise(sen = mean(`mean(sensitivity_prop)`), 
            sel = mean(`mean(selectivity_prop)`)) %>% 
  na.omit() %>% 
  gather(key = "metrics", value = "value", -dataset_id, -resolution) %>% 
  ggplot(aes(x = resolution, y = value, color = dataset_id)) +
  geom_point() +
  geom_smooth() +
  facet_grid(metrics~dataset_id) +
  theme_bw()



# Number of GO temrs per gene
MI_GOterms  %>% 
  filter(GO_domain == "biological_process") %>% 
  select(ensg_id, GO_accession, GO_term_name) %>% 
  distinct() %>% 
  group_by(ensg_id) %>% 
  add_count(ensg_id) %>% 
  ungroup() %>% 
  # select(-GO_accession) %>% 
  # distinct() %>% 
  arrange(-n) 


prop <- 
  all_consensus %>% 
  ungroup() %>% 
  filter(dataset_id == "blood") %>% 
  #mutate(cluster = sample(cluster)) %>% 
  left_join(MI_GOterms %>% 
              filter(GO_domain == "biological_process") %>% 
              select(ensg_id, GO_accession), 
            by = c("gene" = "ensg_id")) %>% 
  na.omit() %>% 
  group_by(dataset_id, resolution, GO_accession) %>%     
  add_count(GO_accession, name = "tot_n") %>% 
  ungroup() %>% 
  group_by(dataset_id, resolution, cluster) %>% 
    do({
      dat <- .
      ngenes <- length(unique(dat$gene))
      dat %>%  mutate (ngenes = ngenes)
    }) %>% 
    group_by(dataset_id, resolution, cluster, GO_accession) %>% 
    add_count(GO_accession) %>%     
    ungroup() %>%  
    mutate(sensitivity_prop = n/ngenes,
           selectivity_prop = n/tot_n) %>% 
    group_by(dataset_id, resolution, cluster) %>% 
    summarise(mean(sensitivity_prop), mean(selectivity_prop)) 

sensitivity <- 
  all_consensus %>% 
  ungroup() %>% 
  filter(dataset_id == "blood") %>% 
  #mutate(cluster = sample(cluster)) %>% 
  left_join(MI_GOterms %>% 
              filter(GO_domain == "biological_process") %>% 
              select(ensg_id, GO_accession), 
            by = c("gene" = "ensg_id")) %>% 
  na.omit() %>% 
  distinct() %>% 
  group_by(dataset_id, resolution, cluster) %>% 
    do({
      dat <- .
      ngenes <- length(unique(dat$gene))
      dat %>%  mutate (ngenes = ngenes)
    }) %>% 
    group_by(dataset_id, resolution, cluster, GO_accession) %>% 
    add_count(GO_accession) %>%     
    ungroup() %>%  
  select(cluster, GO_accession, ngenes, n) %>% 
  distinct() %>% 
    mutate(sensitivity_prop = n/ngenes) %>% 
    group_by(dataset_id, resolution, cluster) %>% 
    summarise(mean(sensitivity_prop), mean(selectivity_prop)) 

sensitivity %>% 
  ungroup() %>% 
  group_by(cluster) %>% 
  summarize(sum(sensitivity_prop))
  top_n(sensitivity_prop, n =10) %>% 
  summarize(sensitivity_prop = mean(sensitivity_prop)) %>% 
  ggplot(aes(x = cluster, y = sensitivity_prop)) +
  geom_point() +
  geom_smooth()

```



# Cluster dispersion
```{r}
umaps_scaled <- 
  all_graph_umaps %>% 
  filter(dataset_id %in% c("blood", "brain", "singlecell", "tissue", "celline_consensus")) %>% 
  gather(UMAP, UMAP_value, UMAP_1, UMAP_2) %>% 
  group_by(dataset_id) %>% 
  mutate(UMAP_value = scales::rescale(UMAP_value, c(-1, 1))) %>% 
  group_by(dataset_id, UMAP) %>% 
  mutate(UMAP_value = UMAP_value - mean(range(UMAP_value))) %>% 
  ungroup() %>%
  spread(UMAP, UMAP_value)

umaps_scaled_clusterings <- 
  umaps_scaled %>% 
  left_join(all_consensus %>% 
  filter(dataset_id %in% c("blood", "brain", "singlecell", "tissue", "celline_consensus")))

umap_dist <- 
  umaps_scaled_clusterings %>% 
  group_by(dataset_id,resolution,cluster) %>% 
  do({
    dat <<- .
    
    dat %>% 
      select(gene,UMAP_1,UMAP_2) %>%
      column_to_rownames("gene") %>% 
      dist() %>%
      as.matrix() %>% 
      as_tibble(rownames = "gene1") %>% 
      gather(gene2,dist,-1) %>% 
      filter(gene1 < gene2) %>% 
      summarise(dist = mean(dist), 
                n_genes = n_distinct(dat$gene))
  })




plot_data <- 
  umap_dist %>% 
  group_by(dataset_id,resolution) %>% 
  mutate(total_genes = sum(n_genes)) %>% 
  ungroup() %>% 
  mutate(weighted_dist = dist * n_genes / total_genes)
  
plot_data %>% 
  group_by(dataset_id,resolution) %>% 
  summarise(mean_dist = mean(dist, na.rm = T), 
            max_dist = max(dist, na.rm = T),
            intracluster_dist = sum(weighted_dist, na.rm = T)) %>% 
  gather(metric, value, mean_dist, max_dist, intracluster_dist) %>% 
  ggplot(aes(x = resolution, y = value)) +
  geom_point() +
  geom_smooth() +
  facet_grid(metric ~ dataset_id, scale = "free_y")

plot_data %>% 
  group_by(dataset_id,resolution) %>% 
  summarise(mean_dist = mean(dist, na.rm = T), 
            max_dist = max(dist, na.rm = T),
            intracluster_dist = sum(weighted_dist, na.rm = T)) %>% 
  ggplot(aes(mean_dist, intracluster_dist, color = dataset_id)) +
  geom_point() +
  geom_abline() +
  coord_fixed()

# plot_data %>% 
#   ggplot(aes(dist, weighted_dist)) + 
#   geom_point() +
#   facet_wrap(~dataset_id)

################


res_consensus_2 <- 
  umaps_scaled_clusterings %>% 
  group_by(dataset_id, resolution) %>% 
  do({
    dat <- . 
    cluster_hulls <- generate_cluster_hulls(V1 = dat$UMAP_1,
                                            V2 = dat$UMAP_2, 
                                            element_id = dat$gene,
                                            cluster_membership = factor(dat$cluster), 
                                            bandwidth_fct = 5,
                                            n = 100)
    
    cluster_hulls$data %>% 
      select(cluster,sub_cluster,n_sub_genes,sub_type) %>% 
      distinct() %>% 
      group_by(cluster) %>% 
      mutate(n_total = sum(n_sub_genes)) %>% 
      ungroup() %>% 
      filter(sub_type == "primary") %>% 
      mutate(prop = n_sub_genes / n_total) %>%
      mutate(avg_prop = mean(prop))
  })
#10.16

res_consensus_2 %>% 
  select(dataset_id, resolution, avg_prop) %>% 
  ggplot(aes(x = resolution, y = avg_prop)) +
  geom_point() +
  geom_smooth() +
  facet_grid(~dataset_id)

```


# UMAP

## Generate UMAP
```{r}
UMAP_file <- 
  paste0("data/processed/", run_id, "_UMAPs.rds")

if(file.exists(UMAP_file)) {
  all_graph_umaps <- 
    readRDS(UMAP_file)
} else {
  
  all_graph_umaps <- 
    all_data_graphs[-5]%>% # cell line graph - ERROR
    pblapply(function(graph_) {
      graph_$snn %>% 
        RunUMAP(umap.method = "umap-learn") %>% 
        {.@cell.embeddings} %>% 
        as_tibble(rownames = "gene") 
    }) %>% 
    bind_rows(.id = "dataset_id")
  
  saveRDS(all_graph_umaps, UMAP_file)
}

final_cluster <- all_consensus %>% 
  filter(dataset_id != "celline") #resolution == 4, seed == 1
  

all_graph_umap_data <- 
  all_graph_umaps %>% 
  left_join(final_cluster) 

all_graph_umaps <- 
    all_data_graphs[[5]]$snn %>% 
        RunUMAP(umap.method = "umap-learn") %>% 
        {.@cell.embeddings} %>% 
        as_tibble(rownames = "gene") 


all_graph_umap_data %>% 
  ggplot(aes(UMAP_1, UMAP_2, color = as.factor(cluster))) +
  geom_point(show.legend = F,
             size = 0.1) +
  facet_grid(~dataset_id) +
  coord_fixed() +
  theme_bw() + 
  scale_color_manual(values = cluster_palette75) +
ggsave(run_id_savename("UMAP.pdf"),
       width = 8, height = 5)

# Scale and center UMAP coordinates:
all_graph_umap_data_scaled <- 
  all_graph_umap_data %>% 
  gather(UMAP, UMAP_value, UMAP_1, UMAP_2) %>% 
  group_by(dataset_id) %>% 
  mutate(UMAP_value = scales::rescale(UMAP_value, c(-1, 1))) %>% 
  group_by(dataset_id, UMAP) %>% 
  mutate(UMAP_value = UMAP_value - mean(range(UMAP_value))) %>% 
  ungroup() %>%
  spread(UMAP, UMAP_value)

all_graph_umap_data_scaled %>% 
  ggplot(aes(UMAP_1, UMAP_2, color = as.factor(cluster))) +
  geom_point(show.legend = F,
             size = 0.1) +
  facet_grid(resolution~dataset_id) +
  coord_fixed() +
  theme_bw() + 
  scale_color_manual(values = cluster_palette100)
ggsave(run_id_savename("UMAP scaled.pdf"),
       width = 8, height = 5)


all_graph_umap_data_scaled %>% 
  ggplot(aes(UMAP_1, UMAP_2)) +
  geom_hex(aes(fill = stat(log10(count))),
           bins = 100) +
  facet_grid(~dataset_id) +
  coord_fixed() +
  theme_bw() + 
  scale_fill_viridis()


all_graph_umap_data_scaled %>% 
  ggplot(aes(UMAP_1, UMAP_2)) +
  geom_density_2d_filled(aes(fill = ifelse(unclass(..level..) < 5, 
                                           NA, 
                                           ..level..)), 
                         bins = 100, n = 100, 
                         h = 0.004) +
  facet_grid(~dataset_id) +
  coord_fixed() +
  theme_void() +
  scale_fill_viridis(name = "%", 
                     na.value = NA)

all_graph_umap_data_scaled %>% 
  ggplot(aes(UMAP_1, UMAP_2)) +
  geom_density_2d_filled(aes(fill = ifelse(unclass(..level..) < 5, 
                                           NA, 
                                           ..level..)), 
                         bins = 100, n = 100, 
                         h = 0.004 * 20) +
  facet_grid(~dataset_id) +
  coord_fixed() +
  theme_void() +
  scale_fill_viridis(name = "%", 
                     na.value = NA)

```


## Generate hulls

```{r}

UMAP_hull_file <- 
  paste0("data/processed/", run_id, "_UMAP_hulls.rds")

if(file.exists(UMAP_hull_file)) {
  all_graph_umap_hulls <- 
    readRDS(UMAP_hull_file)
} else {
  
  all_graph_umap_hulls <- 
    all_graph_umap_data_scaled %>% 
    select(dataset_id) %>% 
    distinct() %>% 
    # slice(6) %>% 
    # slice(1, 4, 6) %>% 
    pull(dataset_id) %>%
    set_names(., .) %>% 
    pblapply(function(dataset_id_) {
      all_graph_umap_data_scaled %>% 
        filter(dataset_id == dataset_id_) %$%
        generate_cluster_hulls(UMAP_1,
                               UMAP_2, 
                               element_id = gene,
                               cluster_membership = factor(cluster), 
                               bandwidth_fct = 5,
                               n = 1000)
    })
  
  saveRDS(all_graph_umap_hulls, UMAP_hull_file)
}

all_graph_umap_hull_data <- 
  all_graph_umap_hulls %>% 
  map(. %$% 
        hulls) %>% 
  bind_rows(.id = "dataset_id")


rm(all_graph_umap_hulls)
gc()
# all_graph_umap_hull_data %>% 
#   filter(density_type == "primary secondary") %>% 
#   select(1, 2, 3, 4, 5) %>% 
#   distinct()

all_graph_umap_hull_data %>% 
  filter(density_type == "primary secondary") %>%
  ggplot(aes(X,Y, group = paste(cluster, sub_cluster, landmass), fill = cluster)) +
  geom_polygon(show.legend = F,
               color = "black") +
  theme_bw() +
  facet_grid(~dataset_id) +
  coord_fixed()
ggsave(run_id_savename("UMAP scaled hulls.pdf"),
       width = 8, height = 5)


# Save
# all_graph_umap_hull_data %>% 
#   filter(density_type == "primary secondary") %>% 
#   unite(polygon_id, cluster, sub_cluster, landmass, remove = F) %>% 
#   select(polygon_id, cluster, UMAP1 = X, UMAP2 = Y) %>% 
#   write_csv(run_id_savename("UMAP draft polygon coordinates.csv"))
#   
# 
# cluster_hulls$data %>% 
#   select(ensg_id = element_id, 
#          UMAP1 = V1, 
#          UMAP2 = V2,
#          cluster) %>% 
#   write_csv(run_id_savename("UMAP draft gene coordinates.csv"))

```


## plots
```{r}

graph_umap_data %>% 
  ggplot(aes(V1, V2, color = factor(cluster))) +
  geom_point(size = 0.1) +
  scale_color_manual(values = cluster_palette) +
  theme_bw() +
  coord_fixed()

example_genes <- 
  data_cluster$cluster_data %>% 
  filter(resolution == 2,
         cluster == 14,
         seed == 1) %>%
  head(30) %>% 
  pull(gene)

data_cluster$neighbors$nn[example_genes, example_genes] %>% 
  {. + t(.)} %>%
  as.matrix() %>% 
  pheatmap(cluster_rows = F,
           cluster_cols = F,
           color = c("white", "gray", "black"))


graph_umap_edges <- 
  data_cluster$neighbors$snn %>%
  elongate_sparsemat() %>% 
  left_join(graph_umap_data,
            by= c("id1" = "gene")) %>% 
  left_join(graph_umap_data,
            by= c("id2" = "gene", "resolution", "seed"),
            suffix = c("1", "2")) %>%
  mutate(dist = sqrt(((V12 - V11) ^ 2) + 
                       ((V22 - V21) ^ 2))) %>%
    filter(id1 < id2)
  
# 
# graph_umap_edges %>% 
#   select(matches("V\\d\\d")) %>% 
#   as.data.frame() 
graph_umap_edges_filtered <- 
  graph_umap_edges %>% 
  # filter(value1 != value2) %>%
  filter(x > 0.25,
         dist > 0.01,
         cluster1 != cluster2)

graph_umap_edges_filtered %>%
  
  
  ggplot(aes(x = V11, xend = V12,
             y = V21, yend = V22)) +
  geom_segment() +
  geom_point(data = graph_umap_data,
             aes(V1, V2, color = factor(cluster)),
             size = 0.1,
             inherit.aes = F) +
  scale_color_manual(values = cluster_palette) +
  theme_bw()


g <- 
  graph_umap_edges_filtered %>% 
                      select(1,2) %>% 
  as.matrix() %>%
  graph_from_edgelist(directed = F)



xy <-
  V(g)$name %>% 
  enframe("i", "gene") %>% 
  left_join(graph_umap_data) %>%
  select(V1, V2) %>%
  as.matrix()


# fbundle <- 
#   edge_bundle_force(g,
#                     xy,
#                     compatibility_threshold = 0.1)

fbundle <- 
  edge_bundle_hammer(g,
                     xy, 
                     decay = 1,
                     bw = 0.075) %>%
  as_tibble() %>% 
  # filter(group == 1) %>% 
  left_join(graph_umap_edges_filtered %>% 
              mutate(group = row_number()) %>% 
              select(group, id1, id2, cluster1, cluster2)) %>% 
  left_join(cluster_palette_tbl,
            by = c("cluster1" = "cluster")) %>%
  left_join(cluster_palette_tbl,
            by = c("cluster2" = "cluster"),
            suffix = c("1", "2")) %>%
  mutate(color = mix_colors(color1, color2, mix = index))

fbundle %>%
  
  ggplot() +
  geom_point(data = graph_umap_data,
             aes(V1, V2, color = factor(cluster)),
             size = 0.1,
             inherit.aes = F) +
  geom_path(aes(x,y,group = group),
            show.legend = FALSE,
            alpha = 0.3)+
  
  
  scale_color_manual(values = cluster_palette) +
  theme_bw() +
  coord_fixed()

fbundle %>%
  
  ggplot() +
  geom_point(data = graph_umap_data %>% 
               mutate(cluster = as.character(cluster)) %>%
               left_join(cluster_palette_tbl),
             aes(V1, V2, color = color),
             size = 0.1,
             inherit.aes = F) +
  geom_path(aes(x,y,
                group = group, 
                color = color),
            show.legend = FALSE)+
  
  scale_color_identity() +
  # scale_color_manual(values = cluster_palette) +
  theme_bw() +
  coord_fixed()

fbundle %>%
  mutate(bundle = paste(cluster1, cluster2)) %>%
  filter(bundle %in% sample(unique(bundle), 20)) %>%
  # filter(value1 == 12, 
  #        value2 == 28) %>% 
  ggplot(aes(x, y)) +
  geom_path(aes(group = group, 
                color = color),
            size = 2,
            show.legend = FALSE)+
  
  scale_color_identity() +
  facet_wrap(~bundle) +
  theme_bw()

  

# Calculate distance
distance_data <- 
  data_cluster$pca_data$scores %>%
  calculate_distance(distance_metric = "pearson") %>%
  {
    genes <- rownames(.)
    as.matrix(.) %>%
      set_colnames(genes) %>%
      set_rownames(genes)
  } %>%
  as.dist()

# Calculate neighbor graph
neighbors <-
  FindNeighbors(
    distance_data,
    k.param = 200,
    compute.SNN = TRUE,
    prune.SNN = 1/15,
    nn.method = "annoy", 
    annoy.metric = "euclidean"
  )

set.seed(42)
graph_umap2 <- 
  neighbors$snn %>%
  {. + t(.)} %>%
  umap(n_neighbors = 15)

graph_umap2_data <- 
  graph_umap2 %>% 
  as_tibble() %>% 
  mutate(gene = rownames(data_cluster$neighbors$nn)) %>% 
  left_join(data_cluster$cluster_data %>% 
              filter(resolution == 2, seed == 1)) 

graph_umap2_data %>% 
  ggplot(aes(V1, V2, color = factor(cluster))) +
  geom_point(size = 0.1) +
  scale_color_manual(values = cluster_palette) +
  theme_bw()


cluster_hulls <- 
  graph_umap2_data %$%
  generate_cluster_hulls(V1,
                         V2, 
                         element_id = gene,
                         cluster_membership = factor(cluster), 
                         n = 500)

cluster_hulls$data %>% 
  ggplot(aes(V1, V2, color = sub_type)) +
  geom_point(size = 0.1) + 
  theme_bw()




p1 <-
  cluster_hulls$data %>% 
  ggplot(aes(V1, V2, color = cluster)) +
  geom_point(size = 0.1, 
             show.legend = F) +
  theme_void() +
  coord_fixed() +
  scale_color_manual(values = cluster_palette) +
  ggtitle("All points")

p2 <-
  cluster_hulls$data %>% 
  filter(sub_type != "outlier") %>%
  ggplot(aes(V1, V2, color = cluster)) +
  geom_point(size = 0.1, 
             show.legend = F) +
  theme_void() +
  coord_fixed() +
  scale_color_manual(values = cluster_palette) +
  ggtitle("No outliers")

p3 <- 
  cluster_hulls$hulls %>% 
  filter(density_type == "primary secondary") %>%
  ggplot(aes(X, Y, fill = cluster, group = paste(cluster, sub_cluster, landmass))) +
  geom_polygon(color = "black", 
               show.legend = F,
               size = 0.1) +
  geom_text(data = . %>% 
              filter(sub_cluster == 1) %>% 
              group_by(cluster) %>%
              summarise(X = mean(X), 
                        Y = mean(Y)),
            aes(X, Y, label = cluster),
            inherit.aes = F) +
  theme_void() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.spacing = unit(0, "mm")) +
  coord_fixed() +
  scale_fill_manual(values = cluster_palette) +
  ggtitle("Complete hull")


p4 <- 
  cluster_hulls$hulls %>% 
  filter(density_type == "primary") %>%
  ggplot(aes(X, Y, fill = cluster, group = paste(cluster, sub_cluster, landmass))) +
  geom_polygon(color = "black", 
               show.legend = F,
               size = 0.1) +
  theme_void() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.spacing = unit(0, "mm")) +
  coord_fixed() +
  scale_fill_manual(values = cluster_palette) +
  ggtitle("Mainland hull")


p1 | p2 | p3 | p4


cluster_hulls$data %>% 
  # filter(sub_type != "outlier") %>%
  left_join(ensg_exp_palette,
            by = c("element_id" = "ensg_id")) %>% 
  ggplot(aes(V1, V2, color = color)) +
  geom_point(size = 0.1, 
             show.legend = F) +
  theme_void() +
  coord_fixed() +
  scale_color_identity() +
  ggtitle("No outliers")


plot_data <- 
  cluster_hulls$data %>% 
  filter(sub_type != "outlier") %>%
  left_join(ensg_exp_palette,
            by = c("element_id" = "ensg_id")) 

plot_range <- 
  range(c(plot_data$V1, plot_data$V2))

plot_n <- 300

pixel_width <- diff(plot_range) / plot_n

plot_coords <- 
  seq(plot_range[1],
      plot_range[2],
      length.out = plot_n) %>% 
  crossing(V1 = ., 
           V2 = .) %>%
  mutate(V1_start = V1 - pixel_width / 2,
         V1_end = V1 + pixel_width / 2,
         V2_start = V2 - pixel_width / 2,
         V2_end = V2 + pixel_width / 2)
  

plot_pixels <- 
  plot_coords %>% 
  group_by_all() %>% 
  do({
    pixel <<- .
    plot_data %>% 
      filter(V1 >= pixel$V1_start,
             V1 < pixel$V1_end,
             V2 >= pixel$V2_start,
             V2 < pixel$V2_end) %>% 
      summarise(red = mean(red),
                green = mean(green),
                blue = mean(blue)) 
  }) %>% 
  ungroup() %>% 
  mutate(red = ifelse(is.finite(red),
                      red,
                      255),
         green = ifelse(is.finite(green),
                        green,
                        255),
         blue = ifelse(is.finite(blue),
                       blue,
                       255))
  
plot_pixels %>% 
  mutate(color = rgb(red, green, blue, maxColorValue = 255)) %>%
  ggplot(aes(V1, V2, fill = color)) +
  geom_tile() +
  theme_void() +
  coord_fixed() +
  scale_fill_identity()

plot_span <- pixel_width/4
red_function <- 
  loess(red ~ V1:V2, data = plot_pixels, span = plot_span)
blue_function <- 
  loess(blue ~ V1:V2, data = plot_pixels, span = plot_span)
green_function <- 
  loess(green ~ V1:V2, data = plot_pixels, span = plot_span)

plot_predict <- 
  seq(plot_range[1],
      plot_range[2],
      length.out = 500) %>% 
  crossing(V1 = ., 
           V2 = .)

plot_canvas <- 
  plot_predict %>% 
  mutate(red = predict(red_function, newdata = .),
         blue = predict(blue_function, newdata = .),
         green = predict(green_function, newdata = .)) %>% 
  mutate_at(c("red", "green", "blue"), function(x) {ifelse(x > 255, 255, x)}) %>%
  mutate_at(c("red", "green", "blue"), function(x) {ifelse(x < 0, 0, x)}) %>%
  filter(complete.cases(.)) %>%
  mutate(color = rgb(red, green, blue, maxColorValue = 255))

plot_canvas %>%
  ggplot(aes(V1, V2, fill = color)) +
  geom_tile() +
  theme_void() +
  coord_fixed() +
  scale_fill_identity()



# ---- tissue contr plots -----


tissue_contr_data <-
  data %>% 
  group_by(ensg_id) %>% 
  mutate(tmm = tmm / sum(tmm)) %>% 
  ungroup()

plots <- 
  lapply(unique(sort(tissue_contr_data$celltype)),
         function(tis_) {
           cluster_hulls$data %>% 
             left_join(tissue_contr_data %>% 
                         filter(celltype == tis_),
                       by = c("element_id" = "ensg_id")) %>%
             ggplot(aes(V1, V2, color = tmm)) +
             geom_point(size = 0.1, 
                        show.legend = F) +
             theme_void() +
             coord_fixed() +
             scale_color_gradient(low = "lightgray", high = "red") +
             ggtitle(tis_)
         })



wrap_plots(plots)

```


### color space

```{r}

shared_colorspace_facet_plot <- 
  function(data, V1, V2, point_id, facet_id, ref_id) {
    plot_data <- 
      data %>% 
      rename(V1 = V1, 
             V2 = V2,
             point_id = point_id, 
             facet_id = facet_id) 
    
    
    ref_palette <- 
      plot_data %>% 
      filter(facet_id == ref_id) %>% 
      mutate(r = scales::rescale(V1, c(0,1)),
             g = scales::rescale(V2, c(0,1)),
             b = 0.5,
             color = rgb(r, g, b)) %>% 
      select(point_id, color)
    
    plot_data %>% 
      left_join(ref_palette,
                by = "point_id") %>% 
      mutate(color = ifelse(is.na(color), 
                            "darkgray", 
                            color))
  
    }

plot_data <- 
  all_graph_umaps %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  group_by(dataset_id) %>% 
  do({
    shared_colorspace_facet_plot(data = all_graph_umap_data_scaled,
                             V1 = "UMAP_1",
                             V2 = "UMAP_2",
                             point_id = "gene",
                             facet_id = "dataset_id",
                             ref_id = .$dataset_id)
  }) %>% 
  ungroup()
  

plot_data %>% 
  ggplot(aes(V1, V2, color = color)) +
  geom_point(size = 0.1) +
  theme_bw() +
  coord_fixed() +
  facet_grid(dataset_id~facet_id) +
  scale_color_identity()


plot_data %>% 
  group_by(dataset_id, 
           facet_id) %>% 
  
  do({
    g_data <- .
    g_data %$% 
      makeHexData(x = V1, 
                  y = V2, 
                  z = color, 
                  bins = 50, color_mean)
  }) %>% 
  ggplot(aes(x = x, y = y, fill = z, group = paste(dataset_id, facet_id))) +
  geom_hex(stat = "identity") +
  facet_grid(dataset_id~facet_id) +
  theme_bw() +
  coord_fixed() +
  scale_fill_identity()

```

## 3D

```{r}
# ----- 3D UMAP -----
graph_umap_3d <- 
  data_cluster$neighbors$nn %>%
  {. + t(.)} %>%
  umap(n_components = 3)

umap_3d %>% 
  as_tibble() %>% 
  mutate(gene = rownames(data_cluster$pca_data$scores)) %>% 
  left_join(data_cluster$cluster_data %>% 
              filter(resolution == 2, seed == 1)) %>%
  plot_ly(x = ~V1, 
          y = ~V2,
          z = ~V3,
          color = ~cluster,
          colors = cluster_palette, 
          size = 0.1)

# umap_3d %>% 
#   as_tibble() %>% 
#   mutate(gene = rownames(data_cluster$pca_data$scores)) %>% 
#   left_join(data_cluster$cluster_data %>% 
#               filter(resolution == 2, seed == 1)) %>%
#   plot_ly(x = ~V1, 
#           y = ~V2,
#           z = ~V3,
#           color = ~gene,
#           colors = set_names(ensg_exp_palette$color,
#                              ensg_exp_palette$ensg_id), 
#           size = 0.1,
#           showscale=FALSE) %>% 
#   add_trace(showlegend = FALSE)
#   add_trace(showscale = FALSE)

# graph_umap_3d <- 
#     data_cluster$neighbors$nn %>%
#     {. + t(.)} %>%
#     umap(n_components = 3, n_neighbors = 15)
  
p <- 
  graph_umap_3d %>% 
  as_tibble() %>% 
  mutate(gene = rownames(data_cluster$pca_data$scores)) %>% 
  left_join(data_cluster$cluster_data %>% 
              filter(resolution == 2, seed == 1)) %>%
  mutate(cluster = factor(cluster)) %>%
  plot_ly(x = ~V1, 
          y = ~V2,
          z = ~V3,
          color = ~cluster,
          colors = set_names(cluster_palette,
                             0:(length(cluster_palette) - 1)), 
          size = 0.1)

p

saveWidgetFix(p, "results/3d_umap.html", selfcontained = T, libdir = "lib")

tissue_contr_data <-
  data %>% 
  group_by(ensg_id) %>% 
  mutate(tmm = tmm / sum(tmm))



gene_info92 <- read_tsv("data/meta/geninfo_92.tsv")

data %>%
  inner_join(data_cluster$cluster_data %>% 
               filter(resolution == 2, seed == 1) %>% 
               filter(cluster == 43),
             by = c("ensg_id" = "gene")) %>%
  left_join(gene_info92) %>%
  select(celltype, tmm, gene_name) %>%
  spread(celltype, tmm) %>% 
  column_to_rownames("gene_name") %>% 
  {log10(. + 1)} %>%
  pheatmap()

data_cluster$cluster_data %>% 
  filter(resolution == 2, seed == 1) %>% 
  filter(cluster == 43) %>%
  left_join(gene_info92,
            by = c("gene" = "ensg_id")) %>%
  pull(gene_name) %>% 
  paste(collapse = "\n") %>% 
  cat()

tissue_contr_data$celltype %>% unique
graph_umap_3d %>% 
  as_tibble() %>% 
  mutate(gene = rownames(data_cluster$pca_data$scores)) %>% 
  left_join(data_cluster$cluster_data %>% 
              filter(resolution == 0.6, seed == 1)) %>%
  left_join(tissue_contr_data %>% 
              filter(celltype == "testis"),
            by = c("gene" = "ensg_id")) %>%
  plot_ly(x = ~V1, 
          y = ~V2,
          z = ~V3,
          name = ~cluster,
          color = ~tmm,
          # colors = cluster_palette, 
          size = 0.1)


```

# Annotation report

## Settings

```{r}

# The annotation report will be separately made for each cluster and dataset.
# This table contains a unique ID for each combination of dataset and cluster. 
# If we write code starting from this table, we can control all output easily. 
annotation_report_settings <- 
  final_cluster %>%
  filter(dataset_id %in% c("blood_consensus", "tissue")) %>% 
  select(dataset_id, cluster) %>% 
  distinct() %>% 
  arrange(dataset_id, cluster) %>% 
  mutate(settings_id = row_number()) 



dataset_metadata %>% 
  write_csv("results/Annotation reports/dataset metadata.csv")

```

## Save clustering

```{r}

annotation_report_clustering <- 
  annotation_report_settings %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  left_join(final_cluster) %>% 
  select(dataset_id, gene, cluster)


for(dataset_ in unique(annotation_report_clustering$dataset_id)) {
  annotation_report_clustering %>% 
    filter(dataset_id == dataset_) %>% 
    select(-dataset_id) %>% 
    write_csv(paste0("results/Annotation reports/clustering results/annotation_clustering ", dataset_, ".csv"))
}
```


## Save UMAPs

```{r}

annotation_report_UMAP <- 
  annotation_report_settings %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  left_join(all_graph_umap_data) %>% 
  select(dataset_id, gene, UMAP_1, UMAP_2)

for(dataset_ in unique(annotation_report_UMAP$dataset_id)) {
  annotation_report_UMAP %>% 
    filter(dataset_id == dataset_) %>% 
    select(-dataset_id) %>% 
    write_csv(paste0("results/Annotation reports/UMAP/annotation_UMAP_coords ", dataset_, ".csv"))
}

```


## General clustering metrics

```{r}

annotation_report_general <- 
  annotation_report_settings %>% 
  select(dataset_id) %>% 
  distinct() %>% 
  left_join(final_cluster) %>% 
  group_by(dataset_id, cluster) %>% 
  summarise(n_genes = n_distinct(gene)) %>% 
  ungroup()

write_csv(annotation_report_general,
          "results/Annotation reports/clustering info.csv")

```


## Expression per cluster

```{r Tissue expression per cluster (pheatmaps)}

# Annotation heatmaps are generated using ggheatmap, as ggplot are easier to handle than pheatmaps

annotation_report_heatmaps <- 
  unique(annotation_report_settings$dataset_id) %>% 
  set_names(., .) %>% 
  pblapply(function(dataset_) {
    
    cat(paste0(dataset_, "\n"))
    plot_data <- 
      all_data[[dataset_]] %>% 
      inner_join(expressed_genes %>% 
                   filter(dataset_id == dataset_) %>% 
                   select(ensg_id),
                 by = "ensg_id") %>% 
      left_join(final_cluster %>% 
                  filter(dataset_id == dataset_) %>% 
                  select(gene, cluster),
                by = c("ensg_id" = "gene")) %>% 
      select(ensg_id, cluster, everything())
    
    plot_settings <- 
      annotation_report_settings %>% 
      filter(dataset_id == dataset_)
    
    plot_settings$cluster %>%
      as.character() %>% 
      set_names(., .) %>% 
      lapply(function(cluster_) {
        cat(paste0("  ", cluster_, "\n"))
        plot_data %>%
          filter(cluster == cluster_) %>% 
          select(-cluster) %>% 
          column_to_rownames("ensg_id") %>%
          apply(MARGIN = 1,
                function(x) x / max(x, na.rm = T)) %>% 
          t() %>% 
          ggheatmap(hclust_method = "ward.D2", 
                    cluster_rows = T, 
                    cluster_cols = T, 
                    color = viridis(20, direction = -1, option = "B"),
                    text_angle_cols = 90, 
                    text_just_cols = c(1, 0.5), 
                    text_show_rows = F, 
                    legendName = "Relative\nexpression")
          # pheatmap(clustering_method = "ward.D2", 
          #          color = 
          #          show_rownames = F,border_color = NA, fontsize = 8)
      })
    
  })

for(dataset_ in names(annotation_report_heatmaps)) {
  
  for(cluster_ in names(annotation_report_heatmaps[[dataset_]])) {
    savename <- 
      paste("results/Annotation reports/expression heatmaps/annotation_exp_heatmap", dataset_, cluster_)
    
    ggsave(annotation_report_heatmaps[[dataset_]][[cluster_]], 
           file = paste0(savename, ".png"), 
           width = 6, height = 5, dpi = 300)
    
    ggsave(annotation_report_heatmaps[[dataset_]][[cluster_]], 
           file = paste0(savename, ".pdf"), 
           width = 6, height = 5)
    
  }
}



```

## Fischer test

```{r}

# This section is very rudimentary right now - needs further development
# Max TO DO: Fix the bug in the Fischer test that gives wrong number of genes overlapping
# TO DO perform fischer test to various classification results including cell line, brain, single cell, 

annotation_report_fischer_tissue <-
  annotation_report_settings %>% 
  select(dataset_id) %>% 
  distinct %>% 
  group_by(dataset_id) %>% 
  do({
    g_dataset <- .$dataset_id
    
    cluster_membership <- 
      final_cluster %>% 
      filter(dataset_id == g_dataset) %$% 
      set_names(cluster,
                gene)
    
    term_membership <- 
      classification_tissue_long %$%
      split(gene, enriched_tissues)

    gene_fischer_test(cluster_membership = cluster_membership, 
                      term_membership = term_membership)
  }) %>% 
  ungroup()

for(dataset_ in unique(annotation_report_fischer_tissue$dataset_id)) {
  annotation_report_fischer_tissue %>% 
    filter(dataset_id == dataset_) %>% 
    select(-dataset_id) %>% 
    write_csv(paste0("results/Annotation reports/fischer test results/annotation_fischertissue ", dataset_, ".csv"))
}

```


## María's functions not yet implemented

```{r Agreement with previous classification}

clus <- 
  all_partitions %>%
  filter(id == "zscore pearson medoids 70") %>%
  select(gene, value)

clas <- 
  gene_info %>%
  select(gene = ensg_id, specificity_category)

# UMAP by classification
umaps[[1]] %>%
  rename(gene = features)%>%
  left_join(clus) %>%
  left_join(clas) %>%
  ggplot(aes(V1,V2, color = specificity_category )) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = gene_category_pal) +
  coord_fixed()


# UMAP by tissue
umaps[[1]] %>%
  rename(gene = features)%>%
  left_join(clus) %>%
  left_join(tissue_info, by = c("gene" = "ensg_id")) %>%
  ggplot(aes(V1,V2, color = enhanced_tissues)) +
  geom_point(show.legend = F) +
  theme_minimal() +
  scale_color_manual(values = tissue_colors) +
  coord_fixed()



# Tissue contributaion UMAPs

#saveRDS(clust_louvain, "results/Human Clustering.RDS")
# load("data/processed/data_scaled.Rdata")
# load("data/processed/pcas.Rdata")

clus_umap <- pcas[[1]]$scores %>%
  set_rownames(gene_names) %>%
  as.data.frame() %>%
  select((1:30)) %>%
  rownames_to_column("enssscg_id") %>%
  umap_calc(row_names = "enssscg_id", n_neigh = 15, n_comp = 2)

clus_umap %>% umap_plot(color_by = "density")

#ggsave("results/human UMAP density.pdf", width = 5, height = 5)

clus_umap %>%  
  left_join(clust_louvain$cluster, 
            by = c("features" = "gene")) %>%
  ggplot(aes(V1,V2, color = as.factor(value))) +
  geom_point(show.legend = F) +
  theme_minimal() +
  # scale_color_manual(values = pal) +
  coord_fixed()

ggsave("results/human UMAP clusters.pdf", width = 5, height = 5)


color_data <- 
  tmm_data %>%
  left_join(colors,
            by = c("celltype" = "tissue")) %>%
  group_by(ensg_id, color = consensus_tissue_color) %>% 
  summarise(tpm = max(tpm)) %>% 
  
  group_by(ensg_id) %>%
  mutate(tpm = tpm / sum(tpm, na.rm = T)) %>%
  ungroup()

ensg_exp_palette <-
  color_data %>%
  bind_cols(col2rgb(.$color) %>% 
              t() %>% 
              as_tibble()) %>%
  mutate(red = red * tpm,
         green = green * tpm,
         blue = blue * tpm) %>%
  group_by(ensg_id) %>% 
  summarise(red = sum(red),
            green = sum(green),
            blue = sum(blue)) %>%
  mutate(color = rgb(red, green, blue, maxColorValue = 255))


clus_umap %>%
  left_join(ensg_exp_palette,
            by = c("features" = "ensg_id")) %>%
  ggplot(aes(V1,V2, color = color)) +
  geom_point(show.legend = F,
             size = 0.2) +
  theme_minimal() +
  scale_color_identity() +
  coord_fixed()

#ggsave("results/human UMAP tissue contribution.pdf", width = 5, height = 5)
```

```{r HT - Fischer function}
load("data/processed/all_partitions.Rdata")
### Fisher score
clustering <- 
  all_partitions %>%
  filter(id == "zscore pearson medoids 70") %>%
  select(gene, value) %>%
  left_join(ortholog_info, by = c("gene" = "enssscg_id")) %>%
  select(gene,value) %>%
  na.omit() 

clust_list <- set_names(clustering$value,
                        clustering$gene)

tissue_list <- 
  tissue_info %$%
  split(ensg_id, enhanced_tissues)


FT_medoids_zscore_pearson_70 <- 
  gene_fischer_test(cluster_membership = clust_list, term_membership = tissue_list)


fisher_res <- FT_medoids_zscore_pearson_70 %>% rename(enhanced_tissues = term)


query_data <-
  
  fisher_res %>% 
  mutate(odds_ratio = ifelse(adj_p < 0.05,
                             odds_ratio, 
                             0)) %>%
  select(gene_name = cluster, tissue = enhanced_tissues, exp = odds_ratio)


query_clustering <-
  query_data %>%
  cluster_query(clustering = "ward.D2", distance_method = "binary")


plot_data <-
  query_data %>%
  mutate(gene_name = factor(gene_name, query_clustering$gene$labels[query_clustering$gene$order]),
         tissue = factor(tissue, query_clustering$sample$labels[query_clustering$sample$order]))

gene_segments <-
  dendro_data(query_clustering$gene)$segments %>%
  as_tibble() %>%
  rename(x = y,
         y = x,
         xend = yend,
         yend = xend) %>%
  mutate(segment_id = row_number()) %>%
  gather(x_type, x_coord, x, xend) %>%
  mutate(x_coord = - scales::rescale(x_coord,
                                     c(0, 5)) + 0.5) %>%
  spread(x_type, x_coord)

sample_segments <-
  dendro_data(query_clustering$sample)$segments %>%
  as_tibble() %>%
  mutate(segment_id = row_number()) %>%
  gather(y_type, y_coord, y, yend) %>%
  mutate(y_coord = scales::rescale(y_coord,
                                   c(0, 5)) +
           n_distinct(query_data$gene_name) + 0.5) %>%
  spread(y_type, y_coord)


plot_data %>%
  left_join(fisher_res, by = c("gene_name"="cluster", "tissue"="enhanced_tissues")) %>%
  filter(exp > 0) %>%
  bind_rows(data.frame(gene_name = as.factor(35))) %>% 
  ggplot(aes(tissue, gene_name,size = exp, fill=tissue)) +
  geom_point( shape = 21) +
  geom_segment(data = gene_segments,
               aes(x, y, xend = xend, yend = yend),
               inherit.aes = F) +
  geom_segment(data = sample_segments,
               aes(x, y, xend = xend, yend = yend),
               inherit.aes = F) +
  coord_fixed() +
  scale_y_discrete(expand = expansion(0),
                   position = "right") +
  theme(panel.background = element_blank(),
        panel.spacing = unit(0, "mm"),
        plot.margin = unit(c(0,0,0,0), "mm"),
        panel.grid =element_line(color="grey85"),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.3, hjust = 1),
        legend.position = "right") +
  scale_fill_manual(values = tissue_colors)



```

```{r Annotation reports}
# Clustering information

load("data/processed/all_partitions.Rdata")

final_clustering <-
  all_partitions %>%
  filter(id=="zscore pearson medoids 70")

n_clusters <- final_clustering %>% pull(value) %>% unique() %>% length()

# Summarized tissue expression data
data_tissue <-
  tmm_data %>%
  mutate(tissue = gsub('.{2}$', '', sample_ID)) %>%
  group_by(tissue,enssscg_id) %>%
  mutate(tissue_exp = mean(tmm)) %>%
  select(enssscg_id, tissue, tissue_exp) %>%
  unique()

scaled_data <-
  data_tissue %>%
  group_by(enssscg_id) %>%
  mutate(scaled_value = tissue_exp / max(tissue_exp)) %>%
  ungroup()


# Run GO enrichments
GO_enrichments <- map(c("BP","CC","MF"),
                      ~enrich(clustering = final_clustering, ontology_type = .x, orthologs = ortholog_info))

#save(GO_enrichments, file = "data/processed/GO_enrichments.Rdata")
#load("data/processed/GO_enrichments.Rdata")


# Annotations
tissues <- scaled_data %>%  pull(tissue) %>%  unique()

tissue_ann <-
  colors %>%  select(tissue,consensus_tissue_name) %>% 
  unique() %>% 
  filter(tissue %in% tissues) %>% 
  column_to_rownames("tissue")

e <- tissue_ann %>% 
  rownames_to_column("tissue") %>% 
  left_join(enframe(tissue_colors), by = c("consensus_tissue_name" = "name")) %>% 
  select(-tissue) %>% 
  unique()

ann_column <- e %>%  pull(value)
names(ann_column) <- e %>% pull(consensus_tissue_name)

row_annotation<-
  gene_info %>% 
  select(ensg_id, specificity_category) %>% 
  filter(ensg_id %in% gene_names)

categories <- row_annotation %>% pull(specificity_category) %>% unique()
ann_row <- gene_category_pal[categories]


all_ann <- list(consensus_tissue_name = ann_column, specificity_category = ann_row)


# Indidicual annotation reports -> selected clusters: 22,43,61,65,67
r22 <- annotation_report(43)
pdf("22.pdf", width = 7, height = 9)
r22
dev.off()

# All annotation reports

reports_v5 <- map(c(1:n_clusters), ~annotation_report(.x))

pdf("reports_v5.pdf", width = 7, height = 9)
reports_v5
dev.off()



# Specificity classifiction legend
load("data/processed/umaps_euc_corr.Rdata")
gene_info %>% 
  select(ensg_id,specificity_category) %>% 
  left_join(umaps[[1]], by = c("ensg_id"="features")) %>% 
  ggplot(aes(x = V1, y = V2, color = specificity_category)) +
  geom_point() +
  scale_color_manual(values = gene_category_pal) +
  theme_simple

# Create file for annotation
cluster_info <-
  final_clustering %>% 
  group_by(value) %>% 
  count %>% 
  rename(cluster = value,
         nr_genes = n) %>% 
  left_join(fisher_res %>% 
              filter(adj_p<0.05) %>% 
              select(cluster,enhanced_tissues) %>% 
              ungroup() %>% 
              group_by(cluster) %>%
              summarize(enhanced_tissues = toString(enhanced_tissues)) %>%
              ungroup %>% 
              mutate(cluster = as.numeric(cluster)))

GO_info <- bind_rows(GO_terms_BP,GO_terms_CC,GO_terms_MF)


gene_description <- final_clustering %>% 
  select(gene, cluster=value) %>% 
  left_join(ortholog_info %>% select(gene=enssscg_id,human_id = ensg_id)) %>% 
  left_join(geninfo %>% 
              select(ensg_id,gene_name,gene_description), by = c("human_id" = "ensg_id")) %>% 
  left_join(gene_info %>% 
              select(ensg_id,specificity_category,distribution_category,enhanced_tissues), by= c("gene" = "ensg_id")) %>% 
  arrange(cluster) 




all <- list(cluster_info, GO_info, gene_description)
names(all) <- c("cluster_info","GO_info","gene_info")
write_xlsx(all, path = "annotation.xlsx")

write.xlsx(cluster_info, file = "annotation.xlsx",
           sheetName = "cluster_info", append = FALSE)
write.xlsx(GO_info, file = "annotation.xlsx", 
           sheetName="GO_info", append=TRUE)
write.xlsx(gene_description, file = "annotation.xlsx",
           sheetName="gene_info", append=TRUE)
```

```{r GOSemSim}
library(GOSemSim)

# Save GO data from orgs.Hs.eg.db
GO_data <- godata("org.Hs.eg.db", ont = "BP", computeIC=FALSE)

# Filtering too broad and too specific terms
GOs <- GO_data@geneAnno%>% group_by(GO) %>% count() %>% filter(n<500) %>% filter (n>3) %>% pull(GO) 

GO_data@geneAnno <-
  GO_data@geneAnno %>% 
  filter(GO %in% GOs)

clustering <- 
  all_partitions %>%
  filter(id == "zscore pearson medoids 70") %>%
  select(gene, value) %>%
  left_join(ortholog_info, by = c("gene" = "enssscg_id")) %>%
  select(entrez,value) %>%
  na.omit() 

num_clusters <- length(unique(clustering$value)) 


cluster_list <- c()
num_clus <- length(unique(clustering$value))
num_genes <- length(clustering$value)
for (i in c(0:num_clus)) {
  genes <- clustering %>%
    filter(value==i) %>%
    pull(entrez) %>%
    as.character()
  cluster_list[[as.character(i)]] <- genes
}


# Calculate semantic similarity matrix
sim_matrix <- mclusterSim(clusters= cluster_list, semData = GO_data, measure = "Wang")

sim_matrix_max_pearson_medoids_50 %>% 
  pheatmap(border_color = NA, 
           clustering_method = "ward.D2", 
           color = colorRampPalette(c("white","#00846b"))(50))


# Filter uncharacterized clusters from the matrix (no or few GO terms present)
unchar <- cluster_annotation %>% 
  filter(Region == "Uncharacterized") %>% 
  pull(cluster)

char <- setdiff(c(1:70),unchar)

sim_matrix_zscore_pearson_medoids_70[char,char]%>% 
  pheatmap(color= viridis(20))
```

```{r GO Semantic similarity UMAP}
load("data/processed/GO_enrichments.Rdata")
all_go_terms <- 
  bind_rows(GO_enrichments[[1]]$enriched_terms,
            GO_enrichments[[2]]$enriched_terms,
            GO_enrichments[[3]]$enriched_terms)
GO <- 
  GO_enrichments[[1]]$enriched_terms %>% # Only BP
  filter(qvalue < 0.05) %>% 
  select(cluster = Cluster,
         GOID = ID)

GO_sparse <-
  GO %>%
  mutate(value = 1) %>% 
  pivot_wider(names_from = GOID, 
              values_from = value, values_fill = 0)

GO_pca <- 
  GO_sparse %>% 
  pca_calc(npcs = 200)

set.seed(42)

GO_umap <- 
  GO_pca$scores %>%
  as_tibble() %>%
  mutate(cluster = GO_sparse_2$cluster) %>% 
  select(c(1:50,"cluster")) %>% 
  umap_calc(n_neigh = 5, row_names = "cluster")

toplot <- cluster_annotation %>%
  mutate(cluster = as.character(cluster)) %>% 
  select(cluster, annotation = Max) %>% 
  left_join(GO_umap, by = c("cluster" = "features"))


toplot %>% 
  ggplot(aes(V1,V2,color=cluster)) +
  geom_point(show.legend = F, size =4) + 
  theme_minimal() +
  geom_text_repel(label = toplot$annotation, show.legend = F, size = 5) +
  scale_color_manual(values = cluster_palette)+
  theme_simple

```

```{r UMAP tissue contribution}

color_data <- 
  data_tissue %>% 
  left_join(colors) %>%
  group_by(enssscg_id, organ_name, color = organ_color) %>% 
  summarise(tissue_exp = max(tissue_exp)) %>% 
  group_by(enssscg_id) %>%
  mutate(tissue_exp = tissue_exp / max(tissue_exp, na.rm = T)) %>%
  ungroup() %>% 
  filter(tissue_exp > 0.1) %>% 
  group_by(enssscg_id) %>%
  mutate(tissue_exp = tissue_exp / sum(tissue_exp, na.rm = T)) %>%
  ungroup() 

ensg_exp_palette <-
  color_data %>%
  bind_cols(col2rgb(.$color) %>% 
              t() %>% 
              as_tibble()) %>%
  mutate(red = red * tissue_exp,
         green = green * tissue_exp,
         blue = blue * tissue_exp) %>%
  group_by(enssscg_id) %>% 
  summarise(red = sum(red),
            green = sum(green),
            blue = sum(blue)) %>%
  mutate(color = rgb(red, green, blue, maxColorValue = 255))

# Tissue contribution plot
umap_zscores_correlation %>%
  left_join(ensg_exp_palette,
            by = c("features" = "enssscg_id")) %>%
  ggplot(aes(V1,V2, color = color)) +
  geom_point(show.legend = F,
             size = 0.2) +
  theme_minimal() +
  scale_color_identity() +
  coord_fixed()

# organ_name legend
color_groups <- tissue_info %>% 
  left_join(colors %>%  select(consensus_tissue_name, organ_name,organ_color),
            by = c("enhanced_tissues" = "consensus_tissue_name"))

organ_colors_df <- color_groups %>%  select(organ_name, organ_color) %>% 
  distinct()

organ_colors <- organ_colors_df$organ_color  
names(organ_colors) <- organ_colors_df$organ_name

umap_zscores_correlation %>% 
  mutate(enssscg_id = gene_names) %>%
  left_join(color_groups, by = c("enssscg_id" = "ensg_id")) %>%
  ggplot(aes(V1,V2, color = organ_name)) +
  geom_point(show.legend = , size=3) +    
  theme_minimal() +
  scale_color_manual(values = organ_colors) +
  coord_fixed()

```

```{r UMAP summarized organ group expression heatmap}

dat <-
  color_data %>% 
  left_join(final_clustering, by = c("enssscg_id" = "gene")) %>% 
  group_by(value, organ_name) %>%
  summarise(tissue_exp = mean(tissue_exp)) %>%
  group_by(value) %>%
  mutate(tissue_exp = tissue_exp / sum(tissue_exp)) %>% 
  mutate(value = as.factor(value))

# Annotation
dat$value <- factor(dat$value, levels = levels(plot_data$gene_name))
annotation_cluster <- data.frame(value = factor(dat$value, levels = levels(plot_data$gene_name)))
annotation_cluster$value <- as.factor(annotation_cluster$value)
ann_colors <- colorRampPalette(tableau_color_pal(palette = "Classic Cyclic")(13))(70)
names(ann_colors) = as.factor(c(1:70))
organs <- color_data %>%  select(organ_name, color) %>%  distinct() 
ann_organs <- organs$color
names(ann_organs) <- organs$organ_name
annotation_organ <- data.frame(organ_name = organs$organ_name)
rownames(annotation_organ) <- annotation_organ$organ_name
ann = list(value = ann_colors, organ_name = ann_organs)

# Expression heatmap
dat %>% 
  spread(organ_name, tissue_exp) %>% 
  column_to_rownames("value") %>% 
  t() %>% 
  
  pheatmap(color = viridis(20, direction = -1, option = "B"),
           clustering_method = "ward.D2",
           border_color = NA,
           cluster_cols = F,
           cluster_rows = T,
           annotation_col = annotation_cluster,
           annotation_row = annotation_organ,
           annotation_colors = ann)

plot_data %>% 
  left_join(dat, by = c("gene_name" = "value"))%>% 
  select(-tissue,-exp) %>% 
  group_by(gene_name) %>% 
  spread(organ_name, tissue_exp) %>% 
  column_to_rownames("gene_name") %>% 
  t() %>% 
  pheatmap(color = viridis(20, direction = -1, option = "B"),
           clustering_method = "ward.D2",
           border_color = NA)


color_data %>% 
  left_join(final_clustering, by = c("enssscg_id" = "gene")) %>% 
  group_by(value, organ_name) %>%
  summarise(tissue_exp = mean(tissue_exp)) %>%
  group_by(value) %>%
  mutate(tissue_exp = tissue_exp / sum(tissue_exp)) %>% 
  spread(organ_name, tissue_exp) %>% 
  column_to_rownames("value") %>% 
  t() %>% 
  
  pheatmap(color = viridis(20, direction = -1, option = "B"),
           clustering_method = "ward.D2",
           border_color = NA,
           cluster_cols = F,
           annotation_col = annotation_cluster,
           annotation_colors = ann)

```

```{r UMAP with annotation}

# Cluster centers based on average coordinates
cluster_centers<- 
  umap_zscores_correlation %>%
  left_join(final_clustering, by = c("features" = "gene")) %>% 
  group_by(value) %>%
  mutate(V1_mean = mean(V1),
         V2_mean = mean(V2)) %>% 
  select(value,V1_mean,V2_mean) %>% 
  left_join(cluster_annotation %>% select(cluster,Max), by = c("value" = "cluster")) %>% 
  distinct()


# UMAP colored by tissue
umaps[[4]] %>%
  left_join(ensg_exp_palette,
            by = c("features" = "enssscg_id")) %>%
  ggplot(aes(V1,V2, color = color)) +
  geom_point(show.legend = T,
             size = 1, alpha=0.5,  stroke = 0, shape=16) +
  theme_simple_2 +
  scale_color_identity() +
  coord_fixed()


# UMAP colored by cluster
cluster_centers %>% 
  ggplot(aes(V1_mean,V2_mean, label=Max, color = as.factor(value))) + # color=as.factor(value),
  geom_point(size = 4, show.legend = F)+ #, color = "grey29"
  geom_point(data = cluster_centers, 
             inherit.aes = F, aes(V1_mean,V2_mean, label=Max, color = as.factor(value))) + #, color = "gray90"
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  geom_text() +
  simple_theme 


```



